### 2021-03-13 specklelives.tex
Initial Overleaf Import

\documentclass[11pt,preprint]{aastex}
\usepackage{amsmath,amsfonts,amssymb}
\usepackage{mathrsfs}
\DeclareMathAlphabet{\mathpzc}{OT1}{pzc}{m}{it}

\usepackage[]{graphicx, epstopdf}
\graphicspath{{Figures/}}

\usepackage{indentfirst}
\usepackage{lscape}
\usepackage{afterpage}
\usepackage{rotating}
\let\captionbox\undefined
%\usepackage{caption}

%\captionsetup{width=6.3in,font=footnotesize}
\usepackage{wrapfig}
\usepackage{multicol}
\usepackage{textcomp}
\newcommand{\textapprox}{\raisebox{0.5ex}{\texttildelow}}

\usepackage{multirow}

\usepackage{xcolor}
\usepackage{ulem}

\usepackage{soul}

\setlength{\textwidth}{6.5in}
\setlength{\hoffset}{0in}
\setlength{\oddsidemargin}{0in}
\setlength{\evensidemargin}{0in}

\setlength{\textheight}{9.in}
\setlength{\voffset}{0in}
\setlength{\topmargin}{0in}
\setlength{\headsep}{0.25in}
\setlength{\headheight}{0in}

%Some handy-dandy commands
%***Distance and Length***
\newcommand{\meters}{\mbox{m}}
\newcommand{\cm}{\mbox{cm}}
\newcommand{\km}{\mbox{km}}
\newcommand{\dpc}{d_{pc}}
\newcommand{\au}{\mbox{AU}}
\newcommand{\microns}{\mu\mbox{m}}
\newcommand{\rsun}{R_{\sun}}
\newcommand{\rjup}{R_J}
\newcommand{\rearth}{R_{\earth}}

%***Times***
\newcommand{\hours}{\mbox{ hrs}}
\newcommand{\seconds}{\mbox{ s}}
\newcommand{\years}{\mbox{ yrs}}

%**Mass**
\newcommand{\kg}{\mbox{kg}}
\newcommand{\msun}{M_{\sun}}
\newcommand{\mearth}{M_{\earth}}
\newcommand{\mjup}{M_J}

\newcommand{\lambdasci}{\lambda_\mathrm{sci}}
\newcommand{\lambdawfs}{\lambda_\mathrm{wfs}}
\newcommand{\tautd}{\tau_\mathrm{td}}
\newcommand{\tauwfs}{\tau_\mathrm{wfs}}
\newcommand{\dopt}{d_\mathrm{opt}}
\newcommand{\mmax}{m_\mathrm{max}}
\newcommand{\nmax}{n_\mathrm{max}}
\newcommand{\Ji}{\mathrm{Ji}}
\newcommand{\pone}{ {\scalebox{0.6}{+1}}}
\newcommand{\mone}{ {\scalebox{0.6}{-1}} }

\newcommand{\karman}{K\'{a}rm\'{a}n }

\newcommand{\argmax}{\operatornamewithlimits{argmax}}
\newcommand{\argmin}{\operatornamewithlimits{argmin}}

% Alter some LaTeX defaults for better treatment of figures:
    % See p.105 of "TeX Unbound" for suggested values.
    % See pp. 199-200 of Lamport's "LaTeX" book for details.
    %   General parameters, for ALL pages:
    \renewcommand{\topfraction}{0.9}    % max fraction of floats at top
    \renewcommand{\bottomfraction}{0.8} % max fraction of floats at bottom
    %   Parameters for TEXT pages (not float pages):
    \setcounter{topnumber}{2}
    \setcounter{bottomnumber}{2}
    \setcounter{totalnumber}{4}     % 2 may work better
    \setcounter{dbltopnumber}{2}    % for 2-column pages
    \renewcommand{\dbltopfraction}{0.9} % fit big float above 2-col. text
    \renewcommand{\textfraction}{0.07}  % allow minimal text w. figs
    %   Parameters for FLOAT pages (not text pages):
    \renewcommand{\floatpagefraction}{0.7}      % require fuller float pages
        % N.B.: floatpagefraction MUST be less than topfraction !!
    \renewcommand{\dblfloatpagefraction}{0.7}   % require fuller float pages

\begin{document}

\title{The Mysterious Lives Of Speckles. I. Residual atmospheric speckle lifetimes behind AO-fed coronagraphs}
\author{Jared Males}

\begin{abstract}
High contrast imaging observations are fundamentally limited by the spatially and temporally correlated noise source called speckles.  Suppression and removal of speckles are the key goals of wavefront control and adaptive optics (AO), coronagraphy, and a host of post-processing techniques.   The formation of speckles as spatial components of the instantaneous point-spread function (PSF) is well understood. However, while many empirical and simulation studies have been carried out, there has been relatively little effort focused on a theoretical understanding of the time-domain behavior of the intensity behind a coronagraphic wavefront control system.  Here we present a method for calculating the temporal power spectral density (PSD) of the intensity, and then show how the PSD can be used to calculate the statistical speckle lifetime.  Considering a frozen-flow turbulence model, we use these techniques to analyze the speckle lifetimes behind a MagAO-X-like AO system and extremely large telescope (ELT) scale systems.   We find that standard AO control shortens atmospheric speckle lifetime from $\sim$130 ms to $\sim$50 ms, and predictive control will further shorten the lifetime to $\sim$20 ms on MagAO-X.  Speckle-limited integration time is directly proportional to these lifetimes.  While the lifetime varies with telescope diameter, windspeed, and turbulence strength, there are no simple scaling laws among these quantities.   Finally, we describe how the intensity PSD can be used to explore the fundamental limits of post-processing in high contrast imaging.  These results motivate continued effort in the area of wavefront sensor and telemetry based reconstruction and speckle sensing for PSF subtraction, which promise to eliminate the long temporal correlations which currently limit high contrast imaging. Last updated \today.
\end{abstract}

\section{Introduction}

Characterization of extrasolar planets with resolved imaging is limited by the halo of starlight scattered onto the image plane by optical aberrations.  In addition to the photon noise caused by this halo, the scattered light is coherent and produces ``speckles'', which can be thought of as copies of the instrument point spread function (PSF).  This adds structured noise, with both spatial and temporal correlations, which is almost always the limiting noise source in a coronagraphic observation.    Both in space and on the ground, imperfections within the telescope and instrument optics cause speckles.   On the Earth's surface, the aberrations include those imposed by turbulence in the Earth's atmosphere.

The index of refraction of air depends on temperature, and turbulence causes the air temperature to stochastically vary across the telescope aperture.  As the stellar wavefront traverses these regions of different index of refraction, parts of the wavefront travel different optical path lengths.  Due mainly to wind, the atmospheric temperature structure is constantly changing.  The technique of adaptive optics (AO) is used to correct the resulting time-variable optical aberrations in real time.  Even with so-called ``extreme'' AO systems (ExAO, Guyon), this correction is not perfect.  The AO-corrected wavefront, though significantly improved, still causes a halo of speckles significantly brighter than the instrumental speckles in a modern high performance coronagraph system using wavefront control.

In order to analyze the capabilities of ground-based high-contrast imaging, we must understand the spatial and temporal evolution of the residual atmospheric speckles behind an ExAO coronagraph.  Speckle formation in the spatial-domain, as structures within the coronagraphic PSF, is well understood (xxxxx).  In the time-domain, we can imagine two ideal regimes for these structures.  In the first regime, the speckles are perfectly static.  Were this case, it would be a simple matter of measuring the PSF structure and then subtracting it from all images.  While this would still leave the photon noise from the scattered light, we could perfectly remove the spatially correlated noise.  In the second imaginary regime, the speckles are a perfect temporally white noise source: every image taken has a different speckle structure.  Similar to photon noise, it would now be a simple matter of taking enough images to average the speckles to a smooth halo.

Here we are concerned with the actual temporal behavior of atmospheric speckles, which vary on short time-scales making standard PSF-subtraction essentially impossible, but exhibit non-negligible temporal correlations which cause them to average much more slowly than photon noise.  The statistical speckle lifetime, that is the time-scale which controls the averaging of speckle noise, is the quantity that sets the fundamental sensitivity limit of ground-based high-contrast imaging.

There have been several definitions of speckle lifetime presented in the literature.  Early treatments of speckle dynamics were motivated by speckle interferometry, which relies on exposures short enough to ``freeze the atmosphere''.  The optimum exposure time is therefore dependent on the coherence time of the intensity in the focal plane.  A common definition for this coherence time was based on a $1/e$ reduction in the autocorrelation of the intensity, which was experimentally found to be on the order of a few ms \citep{1978ApOpt..17.3779S,1990JMOp...37.1247D}.    \citet{1982JOpt...13..263R} derived the following expression for the autocorrelation e-folding time in multi-layer Kolmogorov atmosphere:
\begin{equation}
\tau_{boil} = 0.36 \frac{r_0}{\sqrt{\sigma^2_\mathcal{V}}}
\label{eqn:e-fold}
\end{equation}
called there the ``speckle boiling time'', where $r_0$ is Fried's parameter and $\sigma^2_\mathcal{V}$ is the variance with respect to height of the layer wind velocities $\mathcal{V}$  above the telescope.  At good astronomical sites such as MKO and LCO, values of $\tau_{boil}$ should range from $\sim$5 to $\sim$20 ms. \citet{1999PASP..111..587R} employed Equation (\ref{eqn:e-fold}) for the speckle lifetime and showed that speckles dominate photon noise by factors of 100 or more in high contrast imaging.

The autocorrelation e-folding time is a somewhat arbitrary choice.  If we instead consider the evolution of the variance in the estimate of the mean given uncorrelated trials, we can define the speckle lifetime according to
\begin{equation}
\sigma_{mean}^2 = \frac{\sigma_o^2}{t/\tau}
\label{eqn:varmean_def_intro}
\end{equation}
Here $\tau$ is given by the integral of the autocorrelation \citep{2006ApJ...637..541F}, and in this expression determines the number of uncorrelated realizations of the speckle pattern included in the mean.  Note that this assumes $t >> \tau$.  \citet{1986JOSAA...3.1001A} employed this definition, and argued for longer speckle lifetimes, replacing $0.36$ by $~$$1.14$ in Equation (\ref{eqn:e-fold}) based on arguments similar to \citet{1982JOpt...13..263R}.  Using AO corrected images, \citet{2006ApJ...637..541F} empirically found significantly longer values of $\tau$, giving an overall median of 175 ms at Lick observatory.  Of note, the results of \citet{2006ApJ...637..541F} hinted that inside the region of the image plane controlled by the AO system, the speckle lifetime was significantly shorter.

\citet{2005SPIE.5903..170M} considered the time averaging of residual atmospheric speckles behind a coronagraph, directly analyzing the reduction in variance as a function of exposure time.  They argued that the speckle lifetime should not be affected by the properties of the AO system nor by the seeing parameter $r_0$, and instead should be determined by the wind-crossing time of a telescope with diameter $D$.  Based on a simulation, \citet{2005SPIE.5903..170M} found that 
\begin{equation}
\tau \approx 0.3 \frac{D}{\mathcal{V}}
\end{equation}
for use in Equation (\ref{eqn:varmean_def_intro}).  This result is quite different from the previous predictions, but coarsely in agreement with the experiments by \citet{2006ApJ...637..541F} (it is challenging to make more definitive experimental comparisons due to the uncertainty in atmospheric parameters).

%Note for comparison that the Greenwood time-constant is
%\begin{equation}
%\tau_0 = 0.31 \frac{r_0}{\bar{v}}
%\end{equation}
%where $\bar{v}$ is the 5/3-moment weighted mean velocity (cf. Hardy sec 3.3.6).  $\tau_0$ usually ranges from $\sim$2 to $\sim$8 ms, and is a characteristic time-scale 
%which sets the update rate requirement for AO systems.

More recent experimental studies have used on-sky data with modern ExAO systems.  Using focal plane data from SPHERE, \citet{2016SPIE.9909E..4ZM} found evidence for speckle decorrelation on several time-scales, but were unable to identify an atmospheric component given the various instrument-induced time-scales and the relatively low cadence used. In a study using SCExAO at much higher frame-rates, \citet{2018PASP..130j4502G} analyzed the short time-scale evolution of the intensity using the variance of difference images (essentially a form of autocorrelation).  Here short time-scale (a few ms) speckle evolution was attributed to the atmosphere, and the authors argued that the AO system modulated the speckles on longer time-scales.  \citet{2017JATIS...3b5001S} present a study of speckle lifetimes at the LBT, where they found short-lived speckles attributed to the atmosphere and show some evidence for a difference in lifetime inside and outside the AO control radius.

In the present study, we develop a semi-analytic method for predicting the residual speckle lifetime according to Equation (\ref{eqn:varmean_def_intro}).  This model explicitly accounts for closed-loop AO control, including advanced predictive control laws.  We verify the results of the model by comparison to end-to-end simulations of an ideal closed-loop AO system.  We here concentrate on the speckles due to the atmosphere under the frozen flow hypothesis, and do not consider more complicated turbulence evolution (e.g. boiling) nor do we consider instrumental speckles. We begin with a review of the theory of speckle noise in coronagraph images, followed by a derivation of the statistical speckle lifetime based on the power spectral density (PSD) of the intensity.  We then extend the methods of \citet{2018JATIS...4a9001M} to calculate the intensity PSD.  The model predictions for contrast and speckle lifetimes are compared, and we use the model to predict the behavior of specle lifetimes under various conditions.  Finally, we use the intensity PSDs to consider the benefits of telemetry-based post-processing.

\section{Noise Dynamics in Coronagraphic Focal Planes}

The quantity of interest in this analysis is the intensity in the focal plane as a function of time, $I(\vec{r}, t)$, where $\vec{r}$ is the position.  Useful units for this analysis are photons/sec/$(\lambda/D)^2$.  Following \citet{2007ApJ...669..642S} we consider three contributions to the intensity, such that
\begin{equation}
I(\vec{r},t) = I_{*}(t)\left[I_c(\vec{r},t) + I_{as}(\vec{r},t) + I_{qs}(\vec{r},t)\right]
\end{equation}
where $I_{*}$ is the peak intensity of the star's PSF without a coronagraph, itself a function of time due to Strehl ratio variation.  $I_c$ is the fractional contribution from residual diffraction from the coronagraph, as well as purely static aberrations.  $I_{as}$ is the fractional contribution from residual atmospheric speckles, and $I_{qs}$ is the fractional contribution from quasi-static (QS) speckles.  The instantaneous raw contrast is
\begin{equation}
C(\vec{r},t) = \frac{I(\vec{r},t)}{I_*(t)} = I_c(\vec{r},t) + I_{as}(\vec{r},t) + I_{qs}(\vec{r},t).
\end{equation}

The total variance, considering both speckles and and photon noise, is \citep{2007ApJ...669..642S} 
\begin{equation}
\sigma^2_{fp}  = I_* \Delta t [\underbrace{I_c + I_{as} + I_{qs}}_{\mbox{photon noise}} + I_* [ \underbrace{\tau_{as}\left(I_{as}^2 + 2[I_cI_{as} + I_{as}I_{qs}]  \right)}_{\mbox{atm. speckles}} +  \underbrace{\tau_{qs}\left( I_{qs}^2 + 2 I_c I_{qs}\right)}_{\mbox{QS speckles}}]]
\label{eqn:variance}
\end{equation}
where $\tau_{as}$ is the lifetime of the residual atmospheric speckles, and $\tau_{qs}$ is the lifetime of the QS speckles. $\Delta t$ is the exposure time.  For simplicity we have suppressed the $\vec{r}$ dependence.

We have likewise dropped the $t$ dependence, but we note that this is somewhat subtle.  We could assume that $\Delta t$ is sufficiently short that the $I$s can be treated as constant, and then state the long exposure variance as a sum of the variance in these short exposures.  Alternatively, we can treat the $I$s in Equation (\ref{eqn:variance}) as mean values, and then $\sigma^2_{fp}$ is the mean of the variance.  For large total exposure time these will converge.

Our present focus is on the atmospheric residual $I_{as}$.  From here on we will assume a perfect coronagraph and instrument, such that $I_c = 0$ and $I_{qs}$ = 0.  These (optimistic) conditions eliminate the speckle pinning \citep{2001ApJ...558L..71B} terms in Equation (\ref{eqn:variance}).  The variance is then
\begin{equation}
\sigma^2_{fp}  = I_* \Delta t \left(I_{as} + I_* \tau_{as} I_{as}^2 \right)
\label{eqn:variance_as}
\end{equation}
If we attempt to observe a planet, with a contrast with respect to the star of $C_p$, then the peak signal-to-noise ratio of the planet detection will be
\begin{equation}
\mbox{S/N} = \frac{C_p \sqrt{I_* \Delta t}}{ \sqrt{I_{as} (1 + I_*I_{as} \tau_{as})}} 
\end{equation}
For argument's sake, assume a star having $I_* = 10^8$ photons/sec/$(\lambda/D)^2$ in the peak (roughly an 8th mag star on a 10 m telescope in the near-IR), and $I_{as} = 1\times10^{-4}$ (see \citet{2018JATIS...4a9001M}).  Next we (for now) assume the \citet{2005SPIE.5903..170M} crossing-time speckle lifetime for a 10 m telescope with 10 m/s wind, giving $\tau_{as}$ = 0.3 s.  We now have $I_*I_{as} \tau_{as} = 3000$.  We see that under such conditions  $I_*I_{as} \tau_{as} >> 1$, which would hold even for significantly smaller $\tau_{as}$.  This leads us to
\begin{equation}
\mbox{S/N} \approx \frac{C_p \sqrt{\Delta t}}{ I_{as} \sqrt{\tau_{as}}}. 
\end{equation}
Such an observation is therefore speckle noise limited, and the S/N improves with square-root of the number of realizations of the speckle pattern as $\sqrt{\Delta t/ \tau_{as}}$.  The ratio of the time to reach a given S/N when speckle limited, $\Delta t_{sl}$, to the time needed when photon noise limited, $\Delta t_{pn}$ is
\begin{equation}
\frac{\Delta t_{sl}}{\Delta t_{pn}} = I_*I_{as} \tau_{as} >> 1.
\label{eqn:relexptime}
\end{equation}

Understanding the speckle lifetime is therefore crucial to understanding the limits of high contrast imaging.  To that end, we next show how to calculate the statistical correlation time in a stochastic process from its PSD, and then show how to apply it to post-coronagraph speckles.

\subsection{Speckle Lifetime From A PSD}
\label{sec:psd_lifetime}
We will denote the temporal PSD as $\mathcal{T}(f)$ where $f$ is frequency.  We will restrict ourselves to one-sided PSDs, that is $f \ge 0$.  The PSD is normalized such that the variance of measurements from a process governed by the PSD is
\begin{equation}
\sigma_o^2 = \int_{0}^{f_s/2} \mathcal{T}(f) df.
\label{eqn:process_var}
\end{equation}
where $f_s$ is the sampling frequency defined by the sampling rate $\Delta t$ according to $f_s = 1/\Delta t$.

Through the Wiener-Khinchin theorem, the PSD is equivalent to the autocorrelation (they are a Fourier transform pair).  This means that the PSD describes how the measurements derived from the process are correlated.

\begin{figure}[h]
\centering
\includegraphics[width=3.25in]{onLegPol.pdf}
\caption{The orthonormal Legendre polynomials, which are used as a basis for analyzing the intensity time-series.  We show the first six. \label{fig:onlegpol}}
\end{figure}

Assume that we obtain a sample $I(t)$ of length $T$ consisting of measurements uniformly spaced by $\Delta t$.  For notational simplicity we define the time coordinate such that $-T/2 \le t \le T/2$, with no loss of generality.  Next, we express this time-series as a linear expansion using normalized Legendre polynomials as a signal processing basis:
\begin{equation}
P(\xi) = \sqrt{\frac{2n+1}{2}}\, \mathscr{P}_n(\xi)
\label{eqn:basis}
\end{equation}
where $\mathscr{P}_n(\xi)$ are the Legendre polynomials defined in the usual manner \citep[cf.][]{Olver:2010:NHMF} on the domain $-1 \le \xi \le 1$. The first six of these functions are shown in Figure \ref{fig:onlegpol}.  With our normalization we have the orthogonality condition
\begin{equation}
\int_{-1}^{1}P_n(\xi)P_{n'}(\xi) d\xi = \delta_{nn'}
\label{eqn:orthonormal}
\end{equation}
where $\delta_{nn'}$ is the Kronecker delta.

Now changing variables with
\begin{equation}
\xi = \frac{2}{T}t
\end{equation}
we have
\begin{equation}
I(t) = I\left(\frac{T}{2}\xi\right) = \sum_n a_n P_n(\xi)
\label{eqn:basisexpansion}
\end{equation}
with coefficients
\begin{equation}
a_n = \int_{-1}^{1} I\left(\frac{T}{2}\xi\right)P_n(\xi) d\xi.
\label{eqn:expansecoeff}
\end{equation}
The first orthonormal Legendre polynomial is $P_0(\xi) = 1/\sqrt{2}$, which means that $a_0/\sqrt{2}$ is the mean value of $I(t)$.  Thus the variance of the mean is given by
\begin{equation}
\sigma_\mathrm{mean}^2 = \frac{\left\langle \left|a_0\right|^2 \right\rangle}{2}
\label{eqn:varmean_def}
\end{equation}
where $\langle \cdot \rangle$ denotes the ensemble average.  \citet{1976JOSA...66..207N} derived the coefficient covariances in the Zernike polynomial expansion of Kolmogorov phase screens.  We can apply the same procedure to our Legendre expansion of an arbitrary one dimensional PSD.  We give the step-by-step derivation in Appendix \ref{app:tau_psd}, and here give the result that:
\begin{equation}
\sigma_\mathrm{mean}^2 = \frac{\left\langle \left|a_0\right|^2 \right\rangle}{2} = \frac{1}{2T}  \int_0^{\infty} \frac{ J_{\frac{1}{2}}^2(2\pi \kappa)}{\kappa} \mathcal{T}\left( \frac{2}{T} \kappa \right) d\kappa
\label{eqn:varmean}
\end{equation}
where $J_n$ denotes the cylindrical Bessel functions of the first kind.  This expression gives the variance of the mean in a process governed by $\mathcal{T}(f)$.

Now consider the case of a white noise process, which has a constant PSD
\begin{equation}
\mathcal{T}(f) = \beta.
\end{equation}
Evaluating Equation (\ref{eqn:varmean}), the variance of the mean after a time $T$ is 
\begin{equation}
\sigma_\mathrm{mean}^2 = \frac{\beta}{2T}.
\label{eqn:vm_wn}
\end{equation}
Now from Equation (\ref{eqn:process_var}) the total variance, which gives the variance of a single measurement, is
\begin{equation}
\sigma_o^2  = \frac{\beta}{2\Delta t}.
\end{equation}
Rearranging and substituting for $\beta$ in Equation (\ref{eqn:vm_wn}) yields
\begin{equation}
\sigma_\mathrm{mean}^2 =  \frac{\sigma_o^2 \Delta t}{T} = \frac{\sigma_o^2}{N}
\end{equation}
where $N = T/\Delta t$ is the number of measurements.  This is exactly the result we expect for the variance of the mean in a white noise process with variance $\sigma^2_o$.

To generalize this result we define $\tau$ as the correlation length of the process over the period $T$.  For an arbitrary PSD
\begin{equation}
\tau = \lim_{T\to\infty}\frac{ \displaystyle\int_0^{\infty} \frac{ J_{\frac{1}{2}}^2(2\pi \kappa)}{\kappa} \mathcal{T}\left( \frac{2}{T} \kappa \right) d\kappa}{ 2 \displaystyle\int_{0}^{f_s/2} \mathcal{T}(f) df}.
\label{eqn:psd_lifetime}
\end{equation}
As long as $t >> \tau$, the variance of the mean decreases with time according to
\begin{equation}
\sigma_\mathrm{mean}^2 =  \frac{\sigma_o^2}{t/\tau}
\end{equation}

The lifetime is mainly set by the lowest frequency components of the PSD.  The limit in the numerator of Equation (\ref{eqn:psd_lifetime}) means that $f = \frac{2}{T}\kappa \rightarrow 0$, so we can approximate the lifetime as $\tau \approx \mathcal{T}(0)/(2\sigma_o^2)$.  That is, the lifetime is set by the proportion of power in the 0-frequency component of the PSD.  As we will show, both real-time AO control and post-processing act to reduce power in the low frequencies of the PSD, so we therefore expect these processes to reduce the speckle lifetime.

Similar conclusions have been found through other approaches.  The statistical lifetime can be found from the integral of the autocovariance \citep{1986JOSAA...3.1001A}, which gives the 0-frequency component of the PSD (this is essentially how the derivation in Appendix \ref{app:tau_psd} proceeds).  The lifetimes so derived have the same time-averaging interpretation we use here \citep{2006ApJ...637..541F}.  \citet{2006OExpr..14.7499P} used the moving average filter to analyze the time evolution of variance in atmospheric speckles \citep[see also][]{2005SPIE.5903..170M}.  This likewise selects the lowest frequency components as integration time grows, and so produces equivalent results.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Speckle Intensity PSDs
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Speckle Intensity PSDs}

\subsection{Intensity as a Function of Time}
Here we very briefly review the derivation of the intensity time-series from \citet{2018JATIS...4a9001M}, with the goal to introduce the notation needed for this analysis.  See that paper for a detailed treatment.

% The position vector $\vec{q}$ in the pupil plane is defined by the unit vectors $(\hat{\mathpzc{u}},\hat{\mathpzc{v}})$, such that
% \begin{equation}
% \textstyle\vec{q} = u \hat{\mathpzc{u}} + v \hat{\mathpzc{v}} \\
% \label{eqn:q_def}
% \end{equation}
% and the vector spatial-frequency is 
% \begin{equation}
% \vec{k} = k_u \hat{\mathpzc{u}} + k_v \hat{\mathpzc{v}}. 
% \end{equation}
% We discretize the wavefront control problem according to
% \begin{equation}
% \Delta k = \frac{1}{D}.
% \end{equation}
% where $D$ is the aperture diameter.  This yields
% \begin{equation}
% \textstyle\vec{k}_{mn} = \textstyle\frac{m}{D} \hat{\mathpzc{u}} + \textstyle\frac{n}{D} \hat{\mathpzc{v}}
% \label{eqn:kq_def}
% \end{equation}
% where $m$ and $n$ are integer indices.

We use a real-valued Fourier basis
\begin{equation}
M_{mn}^p(\vec{q}) = \cos(2\pi \vec{k}_{mn} \cdot \vec{q}) + p  \sin(2\pi \vec{k}_{mn} \cdot \vec{q})
\label{eqn:modified_fourier}
\end{equation}
where $p=\pm 1$, $\vec{q}$ is the position vector in the pupil plance and the vector spatial-frequency is 
\begin{equation}
\textstyle\vec{k}_{mn} = \textstyle\frac{m}{D} \hat{\mathpzc{u}} + \textstyle\frac{n}{D} \hat{\mathpzc{v}}
\label{eqn:kq_def}
\end{equation}
where $m$ and $n$ are integer indices, $D$ is the aperture diameter, and $(\hat{\mathpzc{u}},\hat{\mathpzc{v}})$ are unit vectors.  The modal basis described by Equation (\ref{eqn:modified_fourier}) is normalized, and has the advantage that on the unobscured circular aperture both $p$ modes at a given $mn$ have the same symmetric spatial and temporal PSD.  This is not true for pure sines and cosines.  We note that this basis is the one derived for the unobscured circular aperture in a 2D Lomb-Scargle analysis \citep{2020arXiv200110200S}, and, as such, phase-shifted Fourier bases with the same properties exist for more complicated apertures.  In this modal basis the phase in the pupil plane is
\begin{equation}
\Phi(\vec{q},t) = \frac{2\pi}{\lambda} \sum\limits_{mn} \left[ h_{mn}^{+}(t) M_{mn}^{+}(\vec{q}) + h_{mn}^{-}(t) M_{mn}^{-}(\vec{q})\right]
\label{eqn:phi_expansion}
\end{equation}
where $h_{mn}^{+}$ and $h_{mn}^{-}$ are the coefficients of the modes in the same units as wavelength, and the phase $\Phi$ has units of radians.

\citet{2018JATIS...4a9001M} then showed how the residual intensity due to the phase $\Phi$ behind a perfect coronagraph is described by
\begin{equation}
I_\Phi(\vec{r},t) \approx \left(\frac{2\pi}{\lambda}\right)^2 2 \sum_{mn} \left\{  \left[ (h_{mn}^{+})^2 + (h_{mn}^{-})^2 \right] \left[ \Ji^2(\pi D k_{mn}^+) +\Ji^2(\pi D k_{mn}^-)\right] + \mbox{cross-terms} \right\}
\label{eqn:cross_terms}
\end{equation}
where $\Ji$ is the Jinc function
\begin{equation}
\Ji(x) = \frac{\mathrm{J}_1(x)}{x}.
\end{equation}
The cross-terms are a series of terms including products of $h_{mn}^+h_{mn}^-$ and $h_{mn}^ph_{m'n'}^{p'}$ ($mn \neq m'n'$) with $\Ji(x)\Ji(x')$. As long as the aperture truncation is accounted for when the long-exposure statistics are calculated,  these small cross-terms vanish and the long exposure intensity is 
\begin{equation}
\left< I_\Phi(\vec{r}_{mn}) \right> =  \left(\frac{2\pi}{\lambda}\right)^2\left< |h_{mn}|^2 \right> .
\label{eqn:contrast_h2}
\end{equation}
where the discretized position 
\begin{equation}
\vec{r}_{mn} = \vec{k}_{mn} \lambda
\end{equation}
describes the relationship between the spatial-frequency in the pupil plane and projected-angular position in the focal plane.

\subsection{The Intensity PSD}

The quantity we seek is the temporal PSD of the intensity time-series described by Equation (\ref{eqn:cross_terms}).  This is the ensemble average of the modulus-squared of its Fourier transform:
\begin{equation}
\mathcal{T}_{I_{mn}}(f;\vec{r}) = \left\langle \left| \mathcal{F} \left\{ I_\Phi(\vec{r},t) \right\} \right|^2 \right\rangle.
\end{equation}
where $\mathcal{F}\{\cdot\}$ denotes the Fourier transform.

The average Fourier transform of a time-series formed from the products of uncorrelated 0-mean processes will be 0, so we can employ the same logic as above and write
\begin{equation}
\mathcal{T}_{I_{mn}}(f;\vec{r}_{mn}) \approx \left(\frac{2\pi}{\lambda}\right)^2  \left\langle \mathcal{F}\left\{  (h_{mn}^{+})^2 + (h_{mn}^{-})^2  \right\} \right\rangle  .
\label{eqn:psd_h2}
\end{equation}
We thus arrive at the need to determine the PSDs of $(h_{mn}^{+})^2$ and $(h_{mn}^{-})^2$.  Next we review how the PSDs of $h_{mn}^{+}$ and $h_{mn}^{-}$ can be calculated.

\subsection{AO Corrected Post-Coronagraph PSDs}

We express the PSDs of the $h_{mn}^p$ in closed loop ($\mathcal{T}_{\mathrm{cl},mn}(f;g)$) as functions of the open-loop PSD ($\mathcal{T}_{\mathrm{ol},mn}(f)$), the error transfer function (ETF), the measurement noise PSD ($\mathcal{T}_{\mathrm{ph},mn}(f)$), and the noise transfer function (NTF): 
\begin{equation}
\mathcal{T}_{\mathrm{cl},mn}(f;g) = \mathcal{T}_{\mathrm{ol},mn}(f) \left| \mbox{ETF}_{\mathrm{cl}}(f;g) \right|^2 + \mathcal{T}_{\mathrm{ph},mn}(f) \left| \mbox{NTF}_{\mathrm{cl}}(f;g) \right|^2 
\label{eqn:cl_psd}
\end{equation}
where the control loop gain $g$ is optimized to minimize the residual variance in the closed-loop PSD \citep{1999aoa..book.....R_ch6,2016ApOpt..55..323P}. \citet{2018JATIS...4a9001M} derived open-loop PSDs of Fourier modes in frozen-flow turbulence.  In simplified form, the wavefront sensor (WFS) noise is given by
\begin{equation}
\sigma_{ph}^2 = \frac{\beta_{p,mn}^2}{F_\gamma \tau_\mathrm{wfs}}
\label{eqn:photon_noise}
\end{equation}
where $\beta_{p,mn}$ is the WFS sensitivity to photon noise \citep{2005ApJ...629..592G}, $F_\gamma$ is the total photon rate (photons/sec) and $\tau_\mathrm{wfs}$ is the WFS integration time.  This gives a white-noise PSD
\begin{equation}
\mathcal{T}_{\mathrm{ph},mn}(f) = 2\frac{\beta_{p,mn}^2}{F_\gamma }.
\end{equation}

We consider ETFs and NTFs for control laws based on linear filters of the form
\begin{equation}
\widetilde{h} (t_i) = \sum_{j=1}^J a_j \widetilde{h}(t_{i-j}) + g\sum_{l=0}^L b_l \Delta h (t_{i-l})  
\label{eqn:linfilt}
\end{equation}
Here the $\widetilde{h}_i$ are the estimates of the mode coefficient going back in time, and  the $\Delta h_i$ are the measurements of the closed-loop residual amplitude of the mode.  The filter combines the $J$ previous estimates and $L$ previous measurements to form the optimum estimate for the current time, which is then applied as a DM command.

With $J=1$, $a_1 = 1$, $L=0$, and $b_0 = 1$ Equation (\ref{eqn:linfilt}) represents the simple integrator (SI), the simplest control law.  \citet{2018JATIS...4a9001M} showed how adding more coefficients, using the well-known Linear Prediction (LP) technique, results in significant performance improvements in terms of post-coronagraph contrast.  Most, if not all, controllers can be cast in the form of Equation (\ref{eqn:linfilt}) \citep{2007JOSAA..24.2645P}.

% \begin{equation}
% H_{wfs}(s) = H_{dm}(s) = \frac{1 - e^{-sT}}{sT}
% \label{eqn:H_wfs}
% \end{equation}
% where $T = 1/f_s$, $f_s$ being the loop sampling frequency, and $s = i 2\pi f$
% 
% \begin{equation}
% H_{\tau}(s) = e^{-s\tau}
% \label{eqn:H_tau}
% \end{equation}
% 
% \begin{equation}
% H_{con}(z;g) = \frac{g \sum_{l=0}^L b_l z^{-l}}{1+\sum_{j=1}^J a_l z^{-j} }
% \label{eqn:H_con}
% \end{equation}
% and we can map from the $z$-domain to the $s$-domain by the substitution $z \rightarrow e^{sT}$.
% 
% \begin{equation}
% \mbox{ETF}_{ol}(s;g) = H_{wfs}(s) H_{con}(s;g) H_{\tau}(s) H_{dm}(s)
% \label{eqn:etf_ol}
% \end{equation}
% From this it follows that the closed-loop ETF is
% \begin{equation}
% \mbox{ETF}_{cl}(s;g) = \left(1 + H_{wfs}(s) H_{con}(s;g) H_{\tau}(s) H_{dm}(s) \right)^{-1}
% \end{equation}
% The closed-loop noise transfer function (NTF) is
% \begin{equation}
% \mbox{NTF}_{cl}(s;g) = H_{dm}(s) H_{\tau}(s)  H_{con}(s;g)\:\mbox{ETF}_{cl}(s ; g)
% \end{equation}

\subsection{The Relationship Between $h_{mn}^+$ and $h_{mn}^-$ in Frozen Flow Turbulence}

An alternate way to describe the time-evolution of the turbulent wavefront is (see \citet{2005ApJ...629..592G} and Appendix B of \citet{2018JATIS...4a9001M})
\begin{equation}
\Phi_{mn}(\vec{q},t) = \frac{2\pi}{\lambda} h_{mn}^\dagger(t) \cos \left( 2\pi \vec{k}_{mn} \cdot \vec{q} + \phi_{mn}(t) \right)
\end{equation}
These relate to the coefficients of the basis defined by Equation (\ref{eqn:modified_fourier}) as
\begin{eqnarray}
h_{mn}^+ &=& \frac{1}{2} h_{mn}^\dagger(t) \left[ \cos(\phi_{mn}(t)) + \sin(\phi_{mn}(t))\right] \\
h_{mn}^- &=& \frac{1}{2} h_{mn}^\dagger(t) \left[ \cos(\phi_{mn}(t)) - \sin(\phi_{mn}(t))\right] \nonumber
\end{eqnarray}
Under the Taylor frozen flow hypothesis the amplitude $h_{mn}^\dagger$ of each spatial frequency is constant, which means that $h_{mn}^-(\phi_{mn}(t))$ = $h_{mn}^+(\phi_{mn}(t) + \frac{\pi}{2})$.  This phase shift yields
\begin{equation}
\mathcal{F}\left\{h_{mn}^-\right\} = \mathcal{F}\left\{h_{mn}^+\right\} e^{i\frac{\pi}{2}}
\end{equation}
which we employ next when calculating the PSDs of the intensity time-series.

\subsection{A Recipe for Intensity PSDs}
We are now ready to calculate $\mathcal{T}_{I_{mn}}(f)$.  Unfortunately, we have found no purely analytic way to calculate $\left| \mathcal{F}\left\{ h^2 \right\} \right|^2$ given only $\left| \mathcal{F}\left\{ h \right\} \right|^2$.  Instead, we employ a straightforward Monte Carlo procedure:
\begin{enumerate}
\item Generate an open-loop time-series for each $h_{mn}^p$ \label{step:olh}:
\begin{enumerate}
\item Generate a Gaussian white-noise time-series of the desired length
\item Fourier transform the white-noise time-series
\item Multiply the Fourier transform by the square-root of the open-loop PSD.  We now have $\mathcal{F}\left\{h_{mn}^p\right\}$ \citep[cf.][]{kasdin_falpha}.
\item Apply the inverse Fourier transform to produce $h_{mn}^p(t)$, and then normalize to unit variance.
\item Repeat for each mode.
\end{enumerate}
\item Correlate the modal time-series by multiplying the vector of coefficients at each time-step by the decomposition of the covariance matrix.  See \citet{2018JATIS...4a9001M} for a treatment of the covariance matrix calculation for the Fourier basis.
\item For each spatial frequency $mn$, replace $h_{mn}^-(t)$ with the phase shifted $h_{mn}^+(t)$ as described above.
\item Generate a WFS measurement noise time-series for each mode, $\eta_{mn}^+$ and $\eta_{mn}^-$, and Fourier transform them.  These are white-noise PSDs, with no phase relationship.
\item Generate the closed-loop Fourier transforms for each mode \label{step:clh}
\begin{eqnarray}
\mathcal{F}\{h_{mn,cl}^+\} = \mathcal{F}\{h_{mn}^+\} \mathrm{ETF}(f) + \mathcal{F}\{\eta_{mn}^+\} \mathrm{NTF}(f) \\
\mathcal{F}\{h_{mn,cl}^-\} = \mathcal{F}\{h_{mn}^-\} \mathrm{ETF}(f) + \mathcal{F}\{\eta_{mn}^-\} \mathrm{NTF}(f) \nonumber
\end{eqnarray}
\item Calculate the closed-loop modal coefficient time-series with the inverse Fourier transform.  We now have $h_{mn,cl}^+(t)$ and $h_{mn,cl}^-(t)$ 
\item Now we can calculate the focal plane intensity as a function of time
\begin{equation}
I_{mn}(t) \approx \frac{2\pi}{\lambda}\left[ \left(h_{mn,cl}^+\right)^2 + \left(h_{mn,cl}^-\right)^2 \right] 
\end{equation}
\item Finally, we calculate $|\mathcal{F}\{I_{mn}(t)\}|^2$.  For best results this is done with a window function applied.
\item Now the entire procedure is repeated N times, averaging the result to form the estimate of $\mathcal{T}_{I_{mn}}(f)$.
\end{enumerate}

In the following sections we compare the results of these calculations to end-to-end numerical simulations and analyze the results in terms of speckle lifetimes.

\afterpage{\clearpage}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Results
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Results}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Simulations
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Simulations}

We conducted a series of end-to-end numerical simulations of an idealized MagAO-X-like instrument to validate our calculations.  Our goal is not to develop a realistic hardware simulation, rather to ensure that the the calculations produce reasonable results.

We simulated a circular unobscured aperture of 6.5 m.   An idealized deformable mirror was used, which exactly produced commanded shapes as projection of modes.  The DM had 2400 degrees of freedom, corresponding to a 48x48 actuator illuminated pupil.  The wavefront sensor (WFS) worked by simply projecting modes onto the phase screen.  The basis set used for control was the modified Fourier basis (Equation (\ref{eqn:modified_fourier})) but orthornomalized with the stabilized Gramm-Schmidt procedure.  This was necessary for the loop to be stable under the LP controller.  Photon noise was added as Gaussian noise uniformly distributed across the WFS image.  The noise was scaled to correspond to the unmodulated pyramid sensor  \cite[$\beta_p = \sqrt{2}$,][]{2005ApJ...629..592G}.

Wavefront propagation was monochromatic using the Fraunhofer approximation, and the science and sensing wavelength were both 800 nm.  The WFS was spatially filtered in the Fourier domain to minimize aliasing.  The perfect 1st order coronagraph was simulated.  The wavefront was spatially filtered before the coronagraph as well, to minimize Gibbs ringing on the dark hole edge due to the digitized pupil.

The simulated system operated at 2000 Hz with a 1.5 frame delay in addition to the 1 frame total sample and hold.  To achieve the 1/2 frame, the WFS used the average turbulence between two time-steps.  Turbulence was generated in 7 layers as fixed oversized phase screens, translated according to the layer windspeed.   Turbulence parameters corresponded to the median LCO model used in \citet{2018JATIS...4a9001M}, where $r_0 = 0.16$ m, $L_0 = 25$ m, and layer averaged windspeed was 18.7 m/s.

The control laws, both SI and LP with 21 coefficients, were optimized on the open-loop PSDs measured by running the simulation with the control loop off.

\subsection{Raw Contrast}
Figure \ref{fig:contrast_2Dcomp} shows post-coronagraph contrast maps, comparing the results of calculations using \citet{2018JATIS...4a9001M} to the simulation outputs.  We show results for an $\infty$ mag guidestar to show the effects of the input turbulence alone, and for an $8$th mag guidestar.  The agreement is very good, though some small differences are visible in the LP (bottom row). Figure \ref{fig:contrast_radprof} shows the median radial profile for each of the maps in Figure \ref{fig:contrast_2Dcomp}.

\begin{figure}
\hspace{-0.3in}
\includegraphics[width=3.54in]{contrast0mag.pdf}
\includegraphics[width=3.54in]{contrast8mag.pdf}
\caption{Comparison of long-exposure raw contrast from control theory predictions and end-to-end simulation for a 6.5 m telescope. Left: without photon noise, considering the input turbulence only.  Right: with photon noise on an 8th mag star assuming the sensitivity of an unmodulated pyramid WFS. Note that the colorbar is adjusted to best show detail.  Agreement between the calculations and simulations is very good.\label{fig:contrast_2Dcomp}}
\end{figure}

\begin{figure}
\hspace{-0.3in}
\includegraphics[width=6.5in]{contrastRP.pdf}
\caption{Median radial profiles of the contrast maps shown in Figure \ref{fig:contrast_2Dcomp}.  Left: without photon noise, considering the input turbulence only.  Right: with photon noise on an 8th mag star assuming the sensitivity of an unmodulated pyramid WFS. Agreement between the calculations and simulations is very good.\label{fig:contrast_radprof}}
\end{figure}

\subsection{Dynamics}
Figure \ref{fig:opd_psdcomp} evaluates the dynamics of a single spatial frequency component  $m,n = 10,14$.  Note that this required re-fitting the Fourier coefficient after the simulations completed, since the control basis was different.

In the top row, the black lines compare the calculated (dot-dashed) and simulated (solid) open-loop (OL) PSD.  The blue lines compare the output of the calculations to the simulations for the SI, and red lines show the LP.  The bottom row shows the ETF for each controller.  On the $\infty$ mag guidestar, the OL PSD shows excellent agreement at low frequencies, with a slight divergence in slope beginning at a $\sim$100 Hz. The SI likewise shows nearly perfect agreement, with the exception of a lower overshoot peak in the simulation.  The LP calculations are qualitatively a good match to the simulations, showing the same general shape, but the simulation shows better rejection at lower temporal frequencies.  This better low-frequency performance is true on the 8th mag star for both SI and LP.

The disagreements evident in the figure are likely due to the differences between the idealized continuous-time semi-analytic calculations and the discretized simulations.  Additionally, the use of an orthogonalized basis in the simulation can be expected to produce different noise rejection characteristics on the 8th mag star.   Overall the comparison of outputs presented here give us confidence that the semi-analytic model is valid, and can be used to analyze the closed-loop contrast performance and wavefront dynamics in a coronagraphic AO system.

\begin{figure}
\hspace{-0.3in}
\includegraphics[width=6.5in]{psdcomp_10_14.pdf}
\caption{Comparison of predicted and simulated PSDs for Fourier mode (m=10,n=14).  Due to the many subtle differences between the idealized semi-analytic model and the discretized end-to-end simulation, we do not expect perfect agreement.  See text for discussion.    \label{fig:opd_psdcomp}}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Speckle Lifetimes
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Speckle Lifetimes}

In Figure \ref{fig:sppsdcomp} we compare the intensity PSDs in the post-coronagraph focal plane.  In intensity, the agreement between calculations and simulations is even better than in phase.  There is a departure at high frequencies, where the calculations predict stronger peaks than occur in simulations. The agreement at low-frequencies, which are most important for determining speckle lifetimes, is very good.

\begin{figure}
\hspace{-0.3in}
\includegraphics[width=6.5in]{spPsdComp_lp_10_14.pdf}
\caption{Comparison of predicted and simulated PSDs for speckle intensity at position (m=10,n=14).  The agreement between the calculations and the simulations is very good at low frequencies, which determine the speckle lifetime.  At higher frequencies the calculations predict stronger peaks than occur in the simulations  \label{fig:sppsdcomp}}
\end{figure}

In Figure \ref{fig:binvarcomp} we show the evolution of the variance of the mean with time.  The lines show the result of applying Equation (\ref{eqn:varmean}). The points show the direct calculation of the variance in the simulations, as a function of increasing averaging length.  Importantly, the solid lines are in nearly perfect agreement with the points, showing the validity of Equation (\ref{eqn:varmean}).  The dot-dashed lines are slightly offset for the LP, which is due to the difference in simulated vs. predicted contrast (equivalent to the overall variance level), but the shape of the predicted evolution with time for both SI and LP controllers is in good agreement with the simulations.

\begin{figure}
\hspace{-0.3in}
\includegraphics[width=6.5in]{binVarComp_lp_10_14.pdf}
\caption{Comparison of predicted and simulated variance evolution for speckle intensity at position (m=10,n=14). \label{fig:binvarcomp}}
\end{figure}

Finnally in Figure \ref{fig:lifetime_2Dcomp} we present maps of speckle lifetime, and in Figure \ref{fig:lifetime_radprof} the median radial profile of speckle lifetime.  Overall these results show that Equation (\ref{eqn:psd_lifetime}) correctly describes the statistical speckle lifetime for a given intensity PSD, which in turn can be predicted by our semi-analytic model of an ExAO coronagraph.

\begin{figure}
\hspace{-0.3in}
\includegraphics[width=3.54in]{lifetimes0mag.pdf}
\includegraphics[width=3.54in]{lifetimes8mag.pdf}
\caption{Comparison of control theory predictions for speckle lifetime to end-to-end simulation for a 6.5 m telescope.  Left: without photon noise, considering the input turbulence only.  Right: with photon noise on an 8th mag star assuming the sensitivity of an unmodulated pyramid WFS. Note that the colorbar is adjusted in each panel to best show detail. For the integrator agreement is very good.  For the predictor on the infinitely bright star, the agreement is good overall but the calculations over-predict speckle lifetimes long a diagonal assocated with the wind vectors.  \label{fig:lifetime_2Dcomp}}
\end{figure}

\begin{figure}
\hspace{-0.3in}
\includegraphics[width=6.5in]{lifetimesRP.pdf}
\caption{Median radial profiles of the lifetime maps shown in Figure \ref{fig:lifetime_2Dcomp}.  Left: without photon noise, considering the input turbulence only.  Right: with photon noise on an 8th mag star assuming the sensitivity of an unmodulated pyramid WFS. The calculations are somewhat conservative at smaller separations in the bright star case.  \label{fig:lifetime_radprof}}
\end{figure}

\clearpage

\subsection{Trends in Speckle Lifetime}
\label{sec:trends}

We now investigate the behavior of speckle lifetime over range of parameters.  The first comparison we make is between control laws.  Here we simply calculate a histogram of the lifetimes over the 48x48 $\lambda/D$ dark hole for the SI and LP controller.  We also apply the Monte Carlo procedure to the OL case.  The three resulting histograms are shown in Figure \ref{fig:slHist}.  It is clear that AO control shortens the speckle lifetime, and that predictive control further reduces speckle lifetime compared to the integrator.

\begin{figure}[h]
\centering
\includegraphics[width=3.25in]{slHist.pdf}
\caption{Distribution of speckle lifetimes in the dark-hole for open-loop (OL), simple integrator (SI), and linear predictor (LP) controllers. Each histogram contains 2400 measurements.  AO control shows a strong impact on speckle lifetime, with predictive control improving further over that of the simple integrator. 
\label{fig:slHist}
}
\end{figure}

Next, we analyze the change in speckle lifetime with telescope diameter, windspeed, and seeing.  Figure \ref{fig:scalings} shows the behavior of speckle lifetime at a constant absolute spatial frequency.  This means that for a given spatial frequency, the shape of the input PSD is the same regardless of $D$, but with an integrated variance different by a factor of $1/D^2$.  The speckles resulting from that spatial frequency occur at different projected separations according to $\lambda/D$.

On the 6.5 m aperture using the SI controller, windspeed has a strong effect, but on the larger ELT scale apertures wind has no significant effect.  We note, that given the range of plausible mean windspeeds, the ELTs are in a different regime of $\mathcal{V} < D$.  That is the crossing time $D/\mathcal{V}$ > 1 s.  Were a mean windspeed significantly higher than 30 m/s to occur, the speckle lifetimes on the ELT will be shorter.

For the LP controller, increasing windspeed tends to increase speckle lifetime.  This is a result of increasing power concentration in the PSD peaks at $\vec{\mathcal{V}} \cdot \vec{k}$ Hz, which causes the LP controller to optimize there at the expense of worse low-frequency rejection, thus causing the lifetime to increase.

Diameter has a weak effect, with lifetimes dropping as $D$ increases.  When combined as ``crossing time'' $D/V$ there is some evidence for the $\tau \propto D/V$ relation predicted by \citet{2005SPIE.5903..170M} on smaller apertures, but this does not hold for larger telescopes and does not apply to predictive control.

Finally, our calculations predict that improving seeing ($r_0$ growing larger) lowers speckle lifetime.  This is because increasing $r_0$ lowers the overall power at each spatial frequency, which in turn causes the control system to perform better at lower temporal frequencies.  This better low-frequency rejection lowers the speckle lifetime.

\begin{figure}
\includegraphics[width=5.5in]{scalings.pdf}
\caption{Changes in speckle lifetime with turbulence parameters.  Here absolute spatial frequency is held constant.  The $(m,n)=(10,14)$ mode on the 6.5 m telescope is compared to the $(m,n)=(60,84)$ mode on the 39 m aperture.  This means the input temporal PSD is identical in shape, though the integrated variance changes $\propto D^{-2}$.  Results shown here for turbulence only, without WFS noise. There are few general trends for $D$ and $V$. For the integrator controller there is evidence of a $D/V$ scaling as found by \citet{2005SPIE.5903..170M} up to around $D/V$$\sim$1, however this breaks down on larger apertures and higher winds, and is not evident in predictive control.  Improvement in $r_0$ (larger value, better seeing) shows a general reduction in speckle lifetime in all cases.
\label{fig:scalings}
}
\end{figure}

We next held relative spatial frequency constant with $D$, such that the resultant speckle occurs at the same relative projected separation in $\lambda/D$ units.  This means that the input temporal PSD has a different shape as well as different power.  While more difficult to interpret in terms of atmospheric parameters, this compares the performance of coronagraphic AO systems given the improvement in spatial resolution with $D$.

Here, higher windspeed generally shortens speckle lifetime.  Under SI control, speckle lifetime increases with $D$, but shows little change under LP control.  As $D/V$, we see that there is a scaling up to $D/\mathcal{V} \sim 2$ for the SI, but this breaks down at longer crossing times (slower winds). For the LP controller speckle lifetime is essentially flat vs. $D/\mathcal{V}$.

As in the absolute spatial frequency case, improving seeing lowers speckle lifetime in this relative spatial frequency comparison.

\begin{figure}
\includegraphics[width=5.5in]{scalings_rel.pdf}
\caption{Changes in speckle lifetime with turbulence parameters.  Here relative spatial frequency is held constant.  The $(m,n)=(10,14)$ mode on the 6.5 m telescope is compared to the $(m,n)=(10,14)$ mode on the 39 m aperture.  This means the input temporal PSD is different in shape.  Results shown here for turbulence only, without WFS noise. In this comparison there is stronger evidence for the integrator controll showing a $D/V$ scaling as found by \citet{2005SPIE.5903..170M} up to around $D/V$$\sim$2. Predictive control shows little dependence on $D$ or $V$.  
\label{fig:scalings_rel}
}
\end{figure}
%\clearpage

\section{The Potential of Reconstruction Post-Processing}
In Section \ref{sec:psd_lifetime} we developed the expansion of the intensity time-series in terms of orthonormal Legendre polynomials $P_n$.  Suppose we have a technique for measuring the Legendre mode coefficients $a_n$, such as from WFS telemetry reconstruction.  We do not specify a specific algorithm, but assuming a suitable one exists we use it determine and then subtract the first $N$ modes from the intensity time-series 
\begin{equation}
I_{mn,pp}(t) = I_{mn}(t) - \sum_{n=0}^{n=N-1} a_n P_n. 
\end{equation}
The outcome will be fewer long-term correlations as the low order modes are sucessively subtracted, which will in turn lower the speckle lifetime.  We can describe this outcome statistically as a high-pass filter (HPF).  The PSD of a mode is
\begin{equation}
\mathcal{T}_n(f) = \mathcal{T}(f) \left|Q_n(f)\right|^2
\end{equation}
where ${Q}_n$ is the Fourier transform of $P_n$, which we derive in Appendix \ref{app:tau_psd}, giving
\begin{equation}
 \left|Q_n(f)\right|^2 = \left(\frac{2n+1}{2}\right)\frac{J_{n+\frac{1}{2}}^2(\pi T f)}{T f}.
\end{equation}
After fitting and subtracting N modes the speckle intensity PSD will be high-pass filtered with squared transfer function
\begin{equation}
\left| H_\mathrm{HPF}(f;N) \right| ^2 = 1 - \frac{1}{2}\sum_{n=0}^{n=N-1} (2n+1) \frac{J_{n+\frac{1}{2}}^2(\pi T f)}{T f}.
\end{equation}
We show examples of the transfer functions for this filter in Figure \ref{fig:leghpf}.

\begin{figure}
\centering
\includegraphics[width=4in]{hpf.pdf}
\caption{The Legendre high pass filter rejection functions for 1, 10, and 100 modes. Fitting and subtracting legendre polynomials reduces low-frequency power, which lowers speckle lifetimes.\label{fig:leghpf}}
\end{figure}

% Now the measurement of the $a_n$ will have uncertainty due to measurement noise in the WFS.   In time $T$ there will be $Tf_s$ WFS measurements.  It then follows from Equations (\ref{eqn:expansecoeff}) and (\ref{eqn:photon_noise}) that the variance in the measurements of $a_n$ will be
% \begin{equation}
% \sigma_{a_n}^2 = \frac{4}{Tf_s} (\sigma_{ph}^2)^2.
% \end{equation}
% The reconstruction process may be noisy in excess of pure photon noise, say due to a variable calibration.  We model this by multiplicative factor $\gamma_{r} \ge 1$.  The PSD due to this measurement noise will be
% \begin{equation}
% \mathcal{T}_{a_n}(f;N,\gamma_r) = \gamma_r^2 \frac{2}{Tf_s} (\sigma_{ph}^2)^2 \sum_{n=0}^{n=N-1}\left(2n+1\right)\frac{J_{n+\frac{1}{2}}^2(\pi T f)}{T f}.
% \end{equation}

Now the fit to the data will have a fractional uncertainty $\gamma$, so the resulting speckle intensity PSD will be
\begin{equation}
\mathcal{T}_{I_{mn},pp}(f;N,\gamma) = \left( 1 -  \frac{(1-\gamma)^2}{2}\sum_{n=0}^{n=N-1}  (2n+1) \frac{J_{n+\frac{1}{2}}^2(\pi T f)}{T f}\right) \mathcal{T}_{I_{mn}}(f).
\end{equation}

\begin{figure}
\centering
\includegraphics[width=4in]{relsens.pdf}
\caption{The relative sensitivity, in terms of exposure time, after subtracting the first N Legendre polynomials with uncertainty in the fit $\gamma$.  The horizontal line at 1 corresponds to the photon-noise limited exposure time.  Residual-atmospheric speckles increase the exposure time by a factor 55 compared to the photon-noise limit.   
WFS telemetry-based PSF subtraction of just the first $\sim$5 low-order Legendre polynomials from the speckle time-series reduces the speckle lifetime, improving sensitivity dramatically. If better than $\sim$$1\%$ accuracy in reconstruction of these modes can be achieved, the speckle noise can be reduced to below the photon noise limit. \label{fig:legrelsens}}
\end{figure}

In Figure \ref{fig:legrelsens} we show the improvement in exposure time as defined by Equation (\ref{eqn:relexptime}), after fitting and removing the first $N$ Legendre polynomials, with an error of $\gamma$ on the fit.  In the figure, the horizontal line corresponds to the photon-noise limited exposure time on an 8th magnitude star.  If the the first $\sim$5 low-order Legendre polynomials can be determined from, say, WFS-based reconstruction, the reduction in speckle lifetime from removing the low-order temporal modes dramatically improves sensitivity by orders of magnitude.  If the modes are reconstructed to better than $\sim$1\% accuracy, the photon-noise limit can be reached.

\clearpage
\section{Discussion}
For some time, long-lived quasi-static speckles have been recognized as the limiting noise source in contemporary high contrast imaging observations \citep[e.g.][]{2007ApJ...654..633H}.  Given their origin within the instrument, and comparatively long lifetimes \citep{2012A&A...541A.136M}, such speckles should be supressable using wavefront sensing and control strategies.  Recent progress in applying ``dark hole'' contrast optimization with focal plane WFS \citep{2020A&A...638A.117P} and various low-order coronagraphic WFS\&C strategites \citep[e.g.][]{2017PASP..129i5002S} show great promise in significantly reducing the impact of internal speckles.  Assuming active quasi-static speckle control becomes routine, the fundamental limit to sensitivity in ground-based high-contrast imaging is set by the Earth's atmosphere.  This limit is set by two components: uncorrelated photon noise and the spatially and temporally correlated speckle noise.  The temporal correlations of atmospheric speckles are, compared to quasi-static speckles, relatively short. This precludes conventional post-processing, but the lifetimes are long enough to significantly increase exposure times compared to the photon-noise.

We have shown that the speckle lifetime in well-corrected post-coronagraph focal planes is an order of magnitude shorter than claimed in previous studies.  This result was obtained by developing a semi-analytic modeling framework, which allows us to calculate the intensity PSD at any point in the focal plane.  This powerful technique allows us to analyze the impact of control law choice, as well as to compare the influence of various telescope and atmosphere parameters on the speckle lifetime.  Of note, we find that the simple $D/\mathcal{V}$ ``crossing-time'' concept holds only for integrator control on contemporary $D\lesssim10$ m telescopes where $D/\mathcal{V} \lesssim 1$ second.  With predictive controllers and on the larger ELT-class apertures under development the crossing-time is not useful for predicting speckle lifetime.  We  note that the parameter study in Section (\ref{sec:trends}) would be prohibitively expensive in terms of time and computing resources if conducted using end-to-end simulations.  Such studies are enabled by the semi-analytic framework we have developed, beginning in \citep{2018JATIS...4a9001M} and extended to intensity dynamics in this work.

The temporal decorrelation of speckle noise has long been recognized as one of the main aspects of PSF-subtraction post-processing \citep{2006ApJ...641..556M}, even though algorithms have typically been focused on the spatial domain.  Recently, various studies have begun to consider the time-domain in post-processing explicitly \citep{2021A&A...646A..24S}.  Knowledge of the intensity PSD allows for such algorithms to be analyzed eficiently.  We demonstrate this by considering the prospect of telemetry-based reconstruction of the speckle time-series.  Nominally such an algorithm would make use of the main WFS images, but other information sources such as accelerometers, low-order WFS telemetry, and environmental sensors could be used.  While we did not present a specific algorithm, we used the temporal intensity PSDs to predict the sensitivity gain possible.  If the first few low-order temporal modes, here described in terms of the Legendre polynomials, can be determined and subtracted from the intensity in the focal plane, orders of magnitude improvements in exposure time are possible.  The fact that only a few low-order temporal modes need to be reconstructed has the important consequence that any such post-processing algorithm does not need to accurately determine the intensity at every time interval, and does not need to be perfect.  Accuracy in the mean, slope, and other low-order terms of a few percent is all that is needed to make these gains.

\section{Conclusion}
The residual atmosphere after AO correction sets the fundamental limit for ground-based high-contrast imaging.  They key parameter for understanding this limit is the statistical speckle lifetime.  We have developed a semi-analytic method for calculating the speckle lifetime, and verified the model outputs using end-to-end simulations.  We have shown that the speckle lifetime is an order of magnitude shorter than previous studies have found, and analyzed the behavior of speckle lifetime with various telescope parameters.  The key development of the post-coronagraph intensity PSD allowed us to analyze post-processing as a time-domain filter and show that it is possible to achieve photon-noise limited sensitivity given a suitable reconstruction algorithm. In future work we will use this framework to analyze more realistic turbulence models, and extend these techniques to instrumental quasi-static aberrations.

\textbf{Acknowledgements:}

\clearpage
\appendix
\section{The Correlation Length of a 1-D PSD}
\label{app:tau_psd}

In one of the seminal analyses in the development of AO, \citet{1976JOSA...66..207N} showed how to calculate the covariance of the Zernike polynomials in Kolmogorov turbulence, using only the spatial power spectral density of the optical phase. Here we adapt that analysis to one-dimensional PSDs of arbitrary form.

As above, we have a one-dimensional, one-sided ($f \ge 0$),  PSD $\mathcal{T}(f)$, where $f$ is normalized on $0 \leq f \leq f_s = 1/\Delta t$ according to Equation (\ref{eqn:process_var}).  We obtain a time-series $I(t)$ from the process goverened by $\mathcal{T}(f)$, which we expand in terms of normalized Legendre polynomials defined in Equation (\ref{eqn:basis}).  The orthogonality condition is given in Equation (\ref{eqn:orthonormal}).

\subsection{Fourier Transform}

Note the following transform pair\footnote{\url{http://dlmf.nist.gov/10.47}}
\begin{eqnarray}
\int_{-1}^{1} \mathscr{P}_n(x) e^{i2\pi k x} dx & = & 2 i^n j_n(2\pi k)\\
2 i^n \int_{-\infty}^{\infty} j_n(2\pi k) e^{-i2\pi x k} dk &=& \mathscr{P}_n(x)
\end{eqnarray}
where the spherical Bessel functions of the first kind are given by
\begin{equation}
j_n(k) = \sqrt{\frac{\pi}{2k}}J_{n+\frac{1}{2}}(k)
\end{equation}
with $J_n$ denoting the Bessel functions of the first kind.  So we have the Fourier transforms of the $P_n(x)$ 
\begin{equation}
Q_n(k) = i^n \sqrt{\frac{2n+1}{2}}\, \frac{J_{n+\frac{1}{2}}(2\pi k)}{\sqrt{k}}
\end{equation}
Since we have normalized the Legendre polynomials, we have the useful result
\begin{equation}
\int_0^\infty \left| Q_n(k) \right|^2 dk = 1 \mbox{ for all }n.
\end{equation}

\subsection{Noll Analysis}

The expansion of $I(t)$ in terms of the $P_n(\xi)$ is given by Equations (\ref{eqn:basisexpansion}) and (\ref{eqn:expansecoeff}).  This series expansion in terms of normalized Legendre polynomials is the 1-D analog of the expansion in Zernike polynomials used by \citet{1976JOSA...66..207N}.  Following that analysis we write the covariance between any two $P_n(\xi)$ as
\begin{equation}
<a_n^* a_{n'}> = \int \int P_n(\xi) C\left(\frac{T}{2}\xi, \frac{T}{2}\xi' \right) P_{n'}(\xi) d\xi d\xi' 
\end{equation}
where
\begin{equation}
C\left(\frac{T}{2}\xi, \frac{T}{2}\xi' \right) = \left<I\left(\frac{T}{2}\xi\right)I\left(\frac{T}{2}\xi'\right) \right>.
\end{equation}
Equivalently, in the Fourier domain we have
\begin{equation}
<a_n^* a_{n'}> = \int \int Q_n^*(\kappa) \Phi \left(\frac{2}{T}\kappa, \frac{2}{T}\kappa' \right) Q_{n'}(\kappa) d\kappa d\kappa' 
\end{equation}
where
\begin{equation}
\Phi \left(\frac{2}{T}\kappa, \frac{2}{T}\kappa' \right) =  \mathcal{T}\left(\frac{2}{T}\kappa \right) \mathcal{T}\left(\frac{2}{T} \kappa' \right) \delta(\kappa-\kappa')
\end{equation}
after the change of variables $f \rightarrow (2/T)\kappa$ in the PSD $\mathcal{T}(f)$.

We now have
\begin{equation}
<a_n^* a_{n'}> = -i^n i^{n'} \frac{\sqrt{2n+1}\sqrt{2n'+1}}{T}   \int_{0}^{\infty} \frac{J_{n+\frac{1}{2}}(2\pi \kappa) J_{n'+\frac{1}{2}}(2\pi \kappa)}{\kappa} \mathcal{T}\left( \frac{2}{T} \kappa \right) d\kappa
\end{equation}
and for $n = n'$, the variance in a single Legendre mode coefficient over the sample length $T$ is
\begin{equation}
\left\langle \left|a_n\right|^2 \right\rangle = \frac{2n+1}{T}  \int_0^{\infty} \frac{ J_{n+\frac{1}{2}}^2(2\pi \kappa)}{\kappa} \mathcal{T}\left( \frac{2}{T} \kappa \right) d\kappa
\label{eqn:varcoeff}
\end{equation}

\bibliographystyle{apj}
\bibliography{specklelives}
\end{document}

### 2021-03-13 specklelives.tex
update

@@
 Recent progress in applying ``dark hole'' contrast optimization with focal plane WFS \citep{2020A&A...638A.117P} and various low-order coronagraphic WFS\&C strategites \citep[e.g.][]{2017PASP..129i5002S} show great promise in significantly reducing the impact of internal speckles.
-Assuming active quasi-static speckle control becomes routine, the fundamental limit to sensitivity in ground-based high-contrast imaging is set by the Earth's atmosphere.
+Assuming active quasi-static speckle control becomes routine, the fundamental sensitivity limit in ground-based high-contrast imaging is set by the Earth's atmosphere.
 This limit is set by two components: uncorrelated photon noise and the spatially and temporally correlated speckle noise.

### 2021-03-17 specklelives.tex
updated

@@
-\documentclass[11pt,preprint]{aastex}
+\documentclass[10pt,preprint]{aastex631}
 \usepackage{amsmath,amsfonts,amssymb}
@@
 
-\usepackage{indentfirst}
-\usepackage{lscape}
+%\usepackage{indentfirst}
+%\usepackage{lscape}
 \usepackage{afterpage}
-\usepackage{rotating}
+%\usepackage{rotating}
 \let\captionbox\undefined
@@
 %\captionsetup{width=6.3in,font=footnotesize}
-\usepackage{wrapfig}
-\usepackage{multicol}
-\usepackage{textcomp}
+%\usepackage{wrapfig}
+%\usepackage{multicol}
+%\usepackage{textcomp}
 \newcommand{\textapprox}{\raisebox{0.5ex}{\texttildelow}}
 
-\usepackage{multirow}
+%\usepackage{multirow}
 
-\usepackage{xcolor}
-\usepackage{ulem}
+%\usepackage{xcolor}
+%\usepackage{ulem}
 
-\usepackage{soul}
+%\usepackage{soul}
 
-\setlength{\textwidth}{6.5in}
-\setlength{\hoffset}{0in}
-\setlength{\oddsidemargin}{0in}
-\setlength{\evensidemargin}{0in}
+%\setlength{\textwidth}{6.5in}
+%\setlength{\hoffset}{0in}
+%\setlength{\oddsidemargin}{0in}
+%\setlength{\evensidemargin}{0in}
 
-\setlength{\textheight}{9.in}
-\setlength{\voffset}{0in}
-\setlength{\topmargin}{0in}
-\setlength{\headsep}{0.25in}
-\setlength{\headheight}{0in}
+%\setlength{\textheight}{9.in}
+%\setlength{\voffset}{0in}
+%\setlength{\topmargin}{0in}
+%\setlength{\headsep}{0.25in}
+%\setlength{\headheight}{0in}
 
@@
 \title{The Mysterious Lives Of Speckles. I. Residual atmospheric speckle lifetimes behind AO-fed coronagraphs}
-\author{Jared Males}
+
+\author{Jared R. Males}
+\affiliation{Steward Observatory, University of Arizona, Tucson, 933 N Cherry Ave, Tucson, AZ 85721, USA}
+
+\author{Michael P. Fitzgerald}
+\affiliation{Department of Physics \& Astronomy, University of California, Los Angeles, CA 90095, USA}
+
+\author{Ruslan Belikov}
+\affiliation{NASA-Ames Research Center, Moffett Blvd., Moffett Field, CA, USA}
+
+\author{Olivier Guyon}
+\affiliation{Astrobiology Center, National Institutes of Natural Sciences, 2-21-1 Osawa, Mitaka, Tokyo, JAPAN}
+\affiliation{Steward Observatory, University of Arizona, Tucson, 933 N Cherry Ave, Tucson, AZ 85721, USA}
+\affiliation{College of Optical Science, University of Arizona, 1630 E University Blvd, Tucson, AZ 85719, USA}
+\affiliation{National Astronomical Observatory of Japan, Subaru Telescope, National Institutes of Natural Sciences, Hilo, HI 96720, USA}
 
@@
 Characterization of extrasolar planets with resolved imaging is limited by the halo of starlight scattered onto the image plane by optical aberrations.
-In addition to the photon noise caused by this halo, the scattered light is coherent and produces ``speckles'', which can be thought of as copies of the instrument point spread function (PSF).
-This adds structured noise, with both spatial and temporal correlations, which is almost always the limiting noise source in a coronagraphic observation.
+In addition to the photon noise caused by this halo, the scattered light is coherent and produces ``speckles'', which can be thought of as copies of the instrument point spread function (PSF) \citep{1995PASP..107..386M}.
+This adds structured noise, with both spatial and temporal correlations, which is the limiting noise source in a coronagraphic observation.
 Both in space and on the ground, imperfections within the telescope and instrument optics cause speckles.
-On the Earth's surface, the aberrations include those imposed by turbulence in the Earth's atmosphere.
+For telescopes on the Earth's surface, the aberrations include those imposed by turbulence in the Earth's atmosphere.
 
@@
 The technique of adaptive optics (AO) is used to correct the resulting time-variable optical aberrations in real time.
-Even with so-called ``extreme'' AO systems (ExAO, Guyon), this correction is not perfect.
+Even with so-called ``extreme'' AO systems \citep[ExAO, ][]{2018ARAA..56..315G} this correction is not perfect.
 The AO-corrected wavefront, though significantly improved, still causes a halo of speckles significantly brighter than the instrumental speckles in a modern high performance coronagraph system using wavefront control.
@@
 In order to analyze the capabilities of ground-based high-contrast imaging, we must understand the spatial and temporal evolution of the residual atmospheric speckles behind an ExAO coronagraph.
-Speckle formation in the spatial-domain, as structures within the coronagraphic PSF, is well understood (xxxxx).
+Speckle formation in the spatial-domain, as structures within the coronagraphic PSF, is well understood \citep{2001ApJ...558L..71B,2003ApJ...596..702P,2007ApJ...669..642S}.
 In the time-domain, we can imagine two ideal regimes for these structures.
@@
 
-Here we are concerned with the actual temporal behavior of atmospheric speckles, which vary on short time-scales making standard PSF-subtraction essentially impossible, but exhibit non-negligible temporal correlations which cause them to average much more slowly than photon noise.
+In reality, the temporal behavior of atmospheric speckles falls between these regimes.
+These speckles vary on short time-scales making standard PSF-subtraction essentially impossible, but exhibit non-negligible temporal correlations which cause them to average much more slowly than photon noise.
 The statistical speckle lifetime, that is the time-scale which controls the averaging of speckle noise, is the quantity that sets the fundamental sensitivity limit of ground-based high-contrast imaging.
@@
 Using focal plane data from SPHERE, \citet{2016SPIE.9909E..4ZM} found evidence for speckle decorrelation on several time-scales, but were unable to identify an atmospheric component given the various instrument-induced time-scales and the relatively low cadence used.
-In a study using SCExAO at much higher frame-rates, \citet{2018PASP..130j4502G} analyzed the short time-scale evolution of the intensity using the variance of difference images (essentially a form of autocorrelation).
+In a study using SCExAO at much higher frame-rates, \citet{2018PASP..130j4502G} analyzed the evolution of the intensity using the variance of difference images (essentially a form of autocorrelation).
 Here short time-scale (a few ms) speckle evolution was attributed to the atmosphere, and the authors argued that the AO system modulated the speckles on longer time-scales.
@@
 For some time, long-lived quasi-static speckles have been recognized as the limiting noise source in contemporary high contrast imaging observations \citep[e.g.][]{2007ApJ...654..633H}.
-Given their origin within the instrument, and comparatively long lifetimes \citep{2012A&A...541A.136M}, such speckles should be supressable using wavefront sensing and control strategies.
-Recent progress in applying ``dark hole'' contrast optimization with focal plane WFS \citep{2020A&A...638A.117P} and various low-order coronagraphic WFS\&C strategites \citep[e.g.][]{2017PASP..129i5002S} show great promise in significantly reducing the impact of internal speckles.
+Given their origin within the instrument, and comparatively long lifetimes \citep{2012A&A...541A.136M}, such speckles should be supressable using wavefront sensing and control (WFS\&C) strategies.
+Recent progress in applying ``dark hole'' contrast optimization with focal plane WFS \citep{2020A&A...638A.117P} and various focal-plane \citep{2010A&A...509A..31G} and low-order coronagraphic WFS\&C strategites \citep[e.g.][]{2017PASP..129i5002S} show great promise in significantly reducing the impact of internal speckles.
 Assuming active quasi-static speckle control becomes routine, the fundamental sensitivity limit in ground-based high-contrast imaging is set by the Earth's atmosphere.

### 2021-03-20 specklelives.tex
updated
Updates from Overleaf

@@
 \newcommand{\textapprox}{\raisebox{0.5ex}{\texttildelow}}
+
+% custom commands for Olivier's comments
+\usepackage{ulem}
+\newcommand{\ogadd}[1]{\textcolor{orange}{#1}}
+\newcommand{\ogrmv}[1]{\textcolor{orange}{\sout{#1}}}
 
@@
 These results motivate continued effort in the area of wavefront sensor and telemetry based reconstruction and speckle sensing for PSF subtraction, which promise to eliminate the long temporal correlations which currently limit high contrast imaging.
-Last updated \today.
 \end{abstract}
@@
 \section{Introduction}
-
 Characterization of extrasolar planets with resolved imaging is limited by the halo of starlight scattered onto the image plane by optical aberrations.
 In addition to the photon noise caused by this halo, the scattered light is coherent and produces ``speckles'', which can be thought of as copies of the instrument point spread function (PSF) \citep{1995PASP..107..386M}.
-This adds structured noise, with both spatial and temporal correlations, which is the limiting noise source in a coronagraphic observation.
+This adds structured noise, with both spatial and temporal correlations, which is the limiting noise source in \ogrmv{a}\ogadd{long exposure} coronagraphic observation\ogadd{s}.
 Both in space and on the ground, imperfections within the telescope and instrument optics cause speckles.
-For telescopes on the Earth's surface, the aberrations include those imposed by turbulence in the Earth's atmosphere.
+For \ogadd{ground-based} telescopes \ogrm{on the Earth's surface}, the aberrations \ogadd{also} include those imposed by \ogadd{atmospheric} turbulence \ogrmv{in the Earth's atmosphere}.
 
@@
 
-In order to analyze the capabilities of ground-based high-contrast imaging, we must understand the spatial and temporal evolution of the residual atmospheric speckles behind an ExAO coronagraph.
+In order to \ogrmv{analyze}\ogadd{estimate} the capabilities of ground-based high-contrast imaging, we must understand the spatial and temporal evolution of the residual atmospheric speckles behind an ExAO coronagraph.
 Speckle formation in the spatial-domain, as structures within the coronagraphic PSF, is well understood \citep{2001ApJ...558L..71B,2003ApJ...596..702P,2007ApJ...669..642S}.
@@
 In reality, the temporal behavior of atmospheric speckles falls between these regimes.
-These speckles vary on short time-scales making standard PSF-subtraction essentially impossible, but exhibit non-negligible temporal correlations which cause them to average much more slowly than photon noise.
+These speckles vary on short time-scales \ogrmv{making standard PSF-subtraction essentially impossible}\ogadd{so they do are not removed by standard PSF subtraction}, but exhibit non-negligible temporal correlations which cause them to average much more slowly than photon noise.
 The statistical speckle lifetime, that is the time-scale which controls the averaging of speckle noise, is the quantity that sets the fundamental sensitivity limit of ground-based high-contrast imaging.
@@
 called there the ``speckle boiling time'', where $r_0$ is Fried's parameter and $\sigma^2_\mathcal{V}$ is the variance with respect to height of the layer wind velocities $\mathcal{V}$  above the telescope.
-At good astronomical sites such as MKO and LCO, values of $\tau_{boil}$ should range from $\sim$5 to $\sim$20 ms.
+At good astronomical sites \ogrmv{such as MKO and LCO}, values of $\tau_{boil}$ should range from $\sim$5 to $\sim$20 ms.
 \citet{1999PASP..111..587R} employed Equation (\ref{eqn:e-fold}) for the speckle lifetime and showed that speckles dominate photon noise by factors of 100 or more in high contrast imaging.
@@
 \end{equation}
-For argument's sake, assume a star having $I_* = 10^8$ photons/sec/$(\lambda/D)^2$ in the peak (roughly an 8th mag star on a 10 m telescope in the near-IR), and $I_{as} = 1\times10^{-4}$ (see \citet{2018JATIS...4a9001M}).
+For argument's sake, assume a star having $I_* = 10^8$ photon\ogrmv{s}/sec/$(\lambda/D)^2$ in the peak (roughly an 8th mag star on a 10 m telescope in the near-IR), and $I_{as} = 1\times10^{-4}$ (see \citet{2018JATIS...4a9001M}).
 Next we (for now) assume the \citet{2005SPIE.5903..170M} crossing-time speckle lifetime for a 10 m telescope with 10 m/s wind, giving $\tau_{as}$ = 0.3 s.
@@
 \end{equation}
-where $p=\pm 1$, $\vec{q}$ is the position vector in the pupil plance and the vector spatial-frequency is 
+where $p=\pm 1$, $\vec{q}$ is the position vector in the pupil \ogrmv{plance}\ogadd{plane} and the vector spatial-frequency is 
 \begin{equation}
@@
 
-\clearpage
 \section{Discussion}
@@
 \end{equation}
-where
+where for a wide-sense stationary process,
 \begin{equation}
-\Phi \left(\frac{2}{T}\kappa, \frac{2}{T}\kappa' \right) =  \mathcal{T}\left(\frac{2}{T}\kappa \right) \mathcal{T}\left(\frac{2}{T} \kappa' \right) \delta(\kappa-\kappa')
+\Phi \left(\frac{2}{T}\kappa, \frac{2}{T}\kappa' \right) =  \mathcal{T}\left(\frac{2}{T}\kappa \right) \delta(\kappa-\kappa')
 \end{equation}

### 2021-03-21 specklelives.tex
partial update with some comments from Rus

@@
 \newcommand{\textapprox}{\raisebox{0.5ex}{\texttildelow}}
+
+% custom commands for Jared's changes
+\usepackage{ulem}
+\newcommand{\jrmadd}[1]{\textcolor{blue}{#1}}
+\newcommand{\jrmrmv}[1]{\textcolor{blue}{\sout{#1}}}
 
@@
 Both in space and on the ground, imperfections within the telescope and instrument optics cause speckles.
-For \ogadd{ground-based} telescopes \ogrm{on the Earth's surface}, the aberrations \ogadd{also} include those imposed by \ogadd{atmospheric} turbulence \ogrmv{in the Earth's atmosphere}.
+For \ogadd{ground-based} telescopes \ogrmv{on the Earth's surface}, the aberrations \ogadd{also} include those imposed by \ogadd{atmospheric} turbulence \ogrmv{in the Earth's atmosphere}.
 
@@
 
-The total variance, considering both speckles and and photon noise, is \citep{2007ApJ...669..642S} 
+The total variance, considering both speckles and and photon noise, \jrmadd{at a point in the focal plane (fp)} is \citep{2007ApJ...669..642S} 
 \begin{equation}
@@
 
-Through the Wiener-Khinchin theorem, the PSD is equivalent to the autocorrelation (they are a Fourier transform pair).
+According to the Wiener-Khinchin theorem, the PSD \jrmadd{is the Fourier transform of the autocorrelation} \jrmrmv{(they are a Fourier transform pair)}.
 This means that the PSD describes how the measurements derived from the process are correlated.
+\jrmadd{\citet{1976JOSA...66..207N} exploited this concept to derive the coefficient covariances in the Zernike polynomial expansion of Kolmogorov phase screens.
+Here we adapt that analysis to arbitrary 1-dimensional PSDs using the Legendre polynomials.}
 
-\begin{figure}[h]
-\centering
-\includegraphics[width=3.25in]{onLegPol.pdf}
-\caption{The orthonormal Legendre polynomials, which are used as a basis for analyzing the intensity time-series.  We show the first six. \label{fig:onlegpol}}
-\end{figure}
-
-Assume that we obtain a sample $I(t)$ of length $T$ consisting of measurements uniformly spaced by $\Delta t$.
-For notational simplicity we define the time coordinate such that $-T/2 \le t \le T/2$, with no loss of generality.
-Next, we express this time-series as a linear expansion using normalized Legendre polynomials as a signal processing basis:
+\jrmadd{In principle, any basis could be used.
+At present our main goal is analyze the statistics of the mean, which corresponds to the first Legendre polynomial.
+Similar to the Zernike polynomials, this basis has an intuitive progression through slope, quadratic, cubic, etc, modes.
+We later utilize these 1-deminsional modes to analyze post-processing as a time-domain filter.
+For that analysis it will be convenient to use a normalized form of Legendre polynomials defined as:}
 \begin{equation}
@@
 
-Now changing variables with
+\begin{figure}[h]
+\centering
+\includegraphics[width=3.25in]{onLegPol.pdf}
+\caption{The orthonormal Legendre polynomials, which are used as a basis for analyzing the intensity time-series.  We show the first six. \label{fig:onlegpol}}
+\end{figure}
+
+\jrmadd{Now we} assume that we obtain a sample $I(t)$ of length $T$ consisting of measurements uniformly spaced by $\Delta t$.
+For notational simplicity we define the time coordinate such that $-T/2 \le t \le T/2$, with no loss of generality.
+Next, we express this time-series as a linear expansion using the normalized Legendre polynomials.
+Changing variables with
 \begin{equation}
@@
 where $\langle \cdot \rangle$ denotes the ensemble average.
-\citet{1976JOSA...66..207N} derived the coefficient covariances in the Zernike polynomial expansion of Kolmogorov phase screens.
-We can apply the same procedure to our Legendre expansion of an arbitrary one dimensional PSD.
-We give the step-by-step derivation in Appendix \ref{app:tau_psd}, and here give the result that:
+
+Next we use the techniques developed in \citet{1976JOSA...66..207N} to calculate $\left\langle \left|a_0\right|^2 \right\rangle$.
+The step-by-step derivation is given in Appendix \ref{app:tau_psd}, and here we give the result that:
 \begin{equation}
@@
 \end{equation}
-Evaluating Equation (\ref{eqn:varmean}), the variance of the mean after a time $T$ is 
+Evaluating Equation (\ref{eqn:varmean}), the variance of the mean after a time $t$ is 
 \begin{equation}
-\sigma_\mathrm{mean}^2 = \frac{\beta}{2T}.
+\sigma_\mathrm{mean}^2 = \frac{\beta}{2t}.
 \label{eqn:vm_wn}
 \end{equation}
-Now from Equation (\ref{eqn:process_var}) the total variance, which gives the variance of a single measurement, is
+Now from Equation (\ref{eqn:process_var}) the total variance\jrmrmv{, which gives the variance of a single measurement,} is
 \begin{equation}
@@
 \begin{equation}
-\sigma_\mathrm{mean}^2 =  \frac{\sigma_o^2 \Delta t}{T} = \frac{\sigma_o^2}{N}
+\sigma_\mathrm{mean}^2 =  \frac{\sigma_o^2 \Delta t}{t} = \frac{\sigma_o^2}{N}
 \end{equation}
-where $N = T/\Delta t$ is the number of measurements.  This is exactly the result we expect for the variance of the mean in a white noise process with variance $\sigma^2_o$.
+where $N = t/\Delta t$ is the number of measurements.  This is exactly the result we expect for the variance of the mean in a white noise process with variance $\sigma^2_o$.
 
-To generalize this result we define $\tau$ as the correlation length of the process over the period $T$.  For an arbitrary PSD
+To generalize this result we define $\tau$ as the correlation length of the process over the period $T$.  Then for an arbitrary PSD
 \begin{equation}
@@
 \end{equation}
-As long as $t >> \tau$, the variance of the mean decreases with time according to
+\jrmrmv{As long as $t >> \tau$, }The variance of the mean decreases with time according to
 \begin{equation}
@@
 \end{equation}
+as long as $t >> \tau$.
 
@@
 \begin{equation}
-\mathcal{T}_{I_{mn}}(f;\vec{r}) = \left\langle \left| \mathcal{F} \left\{ I_\Phi(\vec{r},t) \right\} \right|^2 \right\rangle.
+\mathcal{T}_{I_{mn}}(f) = \left\langle \left| \mathcal{F}_t \left\{ I_\Phi(\vec{r}_{mn},t) \right\} \right|^2 \right\rangle.
 \end{equation}
-where $\mathcal{F}\{\cdot\}$ denotes the Fourier transform.
+where $\mathcal{F}_t\{\cdot\}$ denotes the Fourier transform \jrmadd{in 1 dimension (the time-domain)}.
 
-The average Fourier transform of a time-series formed from the products of uncorrelated 0-mean processes will be 0, so we can employ the same logic as above and write
+The average Fourier transform of a time-series formed from the products of uncorrelated zero-mean processes will be 0, so we can employ the same logic as above and write
 \begin{equation}
-\mathcal{T}_{I_{mn}}(f;\vec{r}_{mn}) \approx \left(\frac{2\pi}{\lambda}\right)^2  \left\langle \mathcal{F}\left\{  (h_{mn}^{+})^2 + (h_{mn}^{-})^2  \right\} \right\rangle  .
+\mathcal{T}_{I_{mn}}(f) \approx \left(\frac{2\pi}{\lambda}\right)^2  \left\langle \left| \mathcal{F}_t\left\{   (h_{mn}^{+})^2 + (h_{mn}^{-})^2  \right\} \right|^2 \right\rangle  .
 \label{eqn:psd_h2}
 \end{equation}
-We thus arrive at the need to determine the PSDs of $(h_{mn}^{+})^2$ and $(h_{mn}^{-})^2$.  Next we review how the PSDs of $h_{mn}^{+}$ and $h_{mn}^{-}$ can be calculated.
+We thus arrive at the need to determine the PSD of $(h_{mn}^{+})^2 + (h_{mn}^{-})^2$.  Next we review how the PSDs of $h_{mn}^{+}$ and $h_{mn}^{-}$ can be calculated.
 
@@
 \begin{equation}
-\mathcal{F}\left\{h_{mn}^-\right\} = \mathcal{F}\left\{h_{mn}^+\right\} e^{i\frac{\pi}{2}}
+\mathcal{F}_t\left\{h_{mn}^-\right\} = \mathcal{F}_t\left\{h_{mn}^+\right\} e^{i\frac{\pi}{2}}.
 \end{equation}
-which we employ next when calculating the PSDs of the intensity time-series.
+\jrmadd{This temporal-phase relationhip between the two modes at each spatial frequency must be included in the PSD calculation in order to obtain accurate speckle lifetimes}.\jrmrmv{which we employ next when calculating the PSDs of the intensity time-series.}
 
 \subsection{A Recipe for Intensity PSDs}
+\label{sect:recipe}
 We are now ready to calculate $\mathcal{T}_{I_{mn}}(f)$.
-Unfortunately, we have found no purely analytic way to calculate $\left| \mathcal{F}\left\{ h^2 \right\} \right|^2$ given only $\left| \mathcal{F}\left\{ h \right\} \right|^2$.
-Instead, we employ a straightforward Monte Carlo procedure:
+Unfortunately, we have found no purely analytic way to calculate $\langle \left| \mathcal{F}_t\left\{ h^2 \right\} \right|^2 \rangle$ given only $\langle \left| \mathcal{F}_t\left\{ h \right\} \right|^2 \rangle$.
+\jrmadd{ This is further complicated by the spatial correlations and temporal relationship between modes.}  We therefore employ a straightforward Monte Carlo procedure:
 \begin{enumerate}
-\item Generate an open-loop time-series for each $h_{mn}^p$ \label{step:olh}:
+\item Generate an open-loop time-series for each $h_{mn}^p$ for both parities.
+\jrmadd{Note that $h_{mn}^-$ is needed for the spatial correlation step, even though it is discarded eventually.} \label{step:olh}
 \begin{enumerate}
@@
 \item Fourier transform the white-noise time-series
-\item Multiply the Fourier transform by the square-root of the open-loop PSD.  We now have $\mathcal{F}\left\{h_{mn}^p\right\}$ \citep[cf.][]{kasdin_falpha}.
+\item Multiply the Fourier transform by the square-root of the open-loop PSD.  We now have $\mathcal{F}_t\left\{h_{mn}^p\right\}$ \citep[cf.][]{kasdin_falpha}.
 \item Apply the inverse Fourier transform to produce $h_{mn}^p(t)$, and then normalize to unit variance.
@@
 \begin{eqnarray}
-\mathcal{F}\{h_{mn,cl}^+\} = \mathcal{F}\{h_{mn}^+\} \mathrm{ETF}(f) + \mathcal{F}\{\eta_{mn}^+\} \mathrm{NTF}(f) \\
-\mathcal{F}\{h_{mn,cl}^-\} = \mathcal{F}\{h_{mn}^-\} \mathrm{ETF}(f) + \mathcal{F}\{\eta_{mn}^-\} \mathrm{NTF}(f) \nonumber
+\mathcal{F}_t\{h_{mn,cl}^+\} = \mathcal{F}_t\{h_{mn}^+\} \mathrm{ETF}(f) + \mathcal{F}_t\{\eta_{mn}^+\} \mathrm{NTF}(f) \\
+\mathcal{F}_t\{h_{mn,cl}^-\} = \mathcal{F}_t\{h_{mn}^-\} \mathrm{ETF}(f) + \mathcal{F}_t\{\eta_{mn}^-\} \mathrm{NTF}(f) \nonumber
 \end{eqnarray}
@@
 \end{equation}
-\item Finally, we calculate $|\mathcal{F}\{I_{mn}(t)\}|^2$.  For best results this is done with a window function applied.
+\item Finally, we calculate $|\mathcal{F}_t\{I_{mn}(t)\}|^2$.  For best results this is done with a window function applied.
 \item Now the entire procedure is repeated N times, averaging the result to form the estimate of $\mathcal{T}_{I_{mn}}(f)$.
@@
 \subsection{Simulations}
-
+\label{sect:simulations}
 We conducted a series of end-to-end numerical simulations of an idealized MagAO-X-like instrument to validate our calculations.
@@
 An idealized deformable mirror was used, which exactly produced commanded shapes as projection of modes.
-The DM had 2400 degrees of freedom, corresponding to a 48x48 actuator illuminated pupil.
+The DM \jrmadd{was capable of making 2400 Fourier modes}, corresponding \jrmadd{to the Nyquist limit of a} 48x48 actuator illuminated pupil.
 The wavefront sensor (WFS) worked by simply projecting modes onto the phase screen.
@@
 Turbulence parameters corresponded to the median LCO model used in \citet{2018JATIS...4a9001M}, where $r_0 = 0.16$ m, $L_0 = 25$ m, and layer averaged windspeed was 18.7 m/s.
+The control laws, both SI and LP with 21 coefficients, were optimized on the open-loop PSDs measured by running the simulation with the control loop off.
 
-The control laws, both SI and LP with 21 coefficients, were optimized on the open-loop PSDs measured by running the simulation with the control loop off.
+\jrmadd{Simulations were conducted first for an infinitely bright star, that is without photon noise, to analyze the impact of only the input disturbance dynamics.
+Next, an 8th magnitude star was simulated to show the effects of photon noise in the WFS.
+For both cases, simulations with both the SI and LP controllers were conducted.
+Individual simulation runs covered an elapsed time of 65 seconds, and the final 60 seconds were analyzed to be sure that the control system had stabilized.
+Each of these four cases were run 10 times with a new turbulence phase screen, and the results averaged.}
 
@@
 We show results for an $\infty$ mag guidestar to show the effects of the input turbulence alone, and for an $8$th mag guidestar.
+Figure \ref{fig:contrast_radprof} shows the median radial profile for each of the maps in Figure \ref{fig:contrast_2Dcomp}.
 The agreement is very good, though some small differences are visible in the LP (bottom row).
-Figure \ref{fig:contrast_radprof} shows the median radial profile for each of the maps in Figure \ref{fig:contrast_2Dcomp}.
+\jrmadd{These differences, apparent in both the maps and the profile, are likely due to the LP conroller optimization leading to slightly different performance due to the different modal basis actually under control in the simulations.
+Another possible source of these (slight) differences is spatial-temporal correlation not fully accounted for in the semi-analytic calculations.}
 
@@
 \includegraphics[width=3.54in]{contrast8mag.pdf}
-\caption{Comparison of long-exposure raw contrast from control theory predictions and end-to-end simulation for a 6.5 m telescope.
+\caption{Comparison of long-exposure raw contrast from semi-analytic calculations (Section \ref{sect:recipe}) and end-to-end simulation (Section \ref{sect:simulations}) for a 6.5 m telescope.
 Left: without photon noise, considering the input turbulence only.
@@
 Note that this required re-fitting the Fourier coefficient after the simulations completed, since the control basis was different.
+\jrmadd{For consistency and brevity we use only this mode for illustration of single-mode dynamics.
+It was arbitrarily chosen, and is representative of typical results.}
 
@@
 This better low-frequency performance is true on the 8th mag star for both SI and LP.
+\jrmadd{The difference in integrated power at low frequencies is small and does not appreciably impact final contrast.}
 
@@
 Additionally, the use of an orthogonalized basis in the simulation can be expected to produce different noise rejection characteristics on the 8th mag star.
-Overall the comparison of outputs presented here give us confidence that the semi-analytic model is valid, and can be used to analyze the closed-loop contrast performance and wavefront dynamics in a coronagraphic AO system.
+
+Overall the comparison of outputs presented here give us confidence that the semi-analytic model gives \jrmadd{reasonable} results \jrmadd{[JRM NOTE: softened from valid]}, and can be used to analyze the closed-loop contrast performance and wavefront dynamics in a coronagraphic AO system.
 
@@
 There is a departure at high frequencies, where the calculations predict stronger peaks than occur in simulations.
-The agreement at low-frequencies, which are most important for determining speckle lifetimes, is very good.
+The agreements in overal power and the fraction of power conatined in low-frequencies, which are most important for determining speckle lifetimes, is very good.
 
@@
 In Figure \ref{fig:binvarcomp} we show the evolution of the variance of the mean with time.
-The lines show the result of applying Equation (\ref{eqn:varmean}).
-The points show the direct calculation of the variance in the simulations, as a function of increasing averaging length.
+The lines show the result of applying Equation (\ref{eqn:varmean})  \jrmadd{to PSDs of the intensity in the simulations}.
+\jrmadd{To test the validity of the derivation of speckle lifetime,} the points show the direct calculation of the variance in the simulations, as a function of increasing averaging length.
 Importantly, the solid lines are in nearly perfect agreement with the points, showing the validity of Equation (\ref{eqn:varmean}).
-The dot-dashed lines are slightly offset for the LP, which is due to the difference in simulated vs.
+The dot-dashed lines, \jrmadd{which are the output of the semi-analytic calculations}, are slightly offset for the LP, which is due to the difference in simulated vs.
 predicted contrast (equivalent to the overall variance level), but the shape of the predicted evolution with time for both SI and LP controllers is in good agreement with the simulations.
@@
 \includegraphics[width=6.5in]{binVarComp_lp_10_14.pdf}
-\caption{Comparison of predicted and simulated variance evolution for speckle intensity at position (m=10,n=14). \label{fig:binvarcomp}}
+\caption{Comparison of predicted and simulated variance evolution for speckle intensity at position (m=10,n=14).
+\jrmadd{Here ``calculated'' is the output of the method in Section \ref{sect:recipe}.
+The ``sim.
+measured'' points (filled circles) show the direct measurements of variance as a function of time in the simulations.
+The ``sim.
+PSD based'' lines show the results of using Equantion (\ref{eqn:psd_lifetime}) to predict the variance as a function of time using the simulated intensity PSDs.
+The comparison between the two simulation-based results shows that the analytic expression for speckle lifetime we derived is valid.
+These are compared to the calculated line to show the result of the semi-analytic method for calculating the intensity PSD without an end-to-end simulation.}  \label{fig:binvarcomp}}
 \end{figure}
 
-Finnally in Figure \ref{fig:lifetime_2Dcomp} we present maps of speckle lifetime, and in Figure \ref{fig:lifetime_radprof} the median radial profile of speckle lifetime.
+Finally in Figure \ref{fig:lifetime_2Dcomp} we present maps of speckle lifetime, and in Figure \ref{fig:lifetime_radprof} the median radial profile of speckle lifetime.
 Overall these results show that Equation (\ref{eqn:psd_lifetime}) correctly describes the statistical speckle lifetime for a given intensity PSD, which in turn can be predicted by our semi-analytic model of an ExAO coronagraph.
@@
 For the integrator agreement is very good.
-For the predictor on the infinitely bright star, the agreement is good overall but the calculations over-predict speckle lifetimes long a diagonal assocated with the wind vectors.
+For the predictor on the infinitely bright star, the agreement is good overall but the calculations \jrmadd{modestly} over-predict speckle lifetimes \jrmadd{in a wedge assocated with (but not parallel to)} the wind vectors.
 \label{fig:lifetime_2Dcomp}}
@@
 
-Note the following transform pair\footnote{\url{http://dlmf.nist.gov/10.47}}
-\begin{eqnarray}
-\int_{-1}^{1} \mathscr{P}_n(x) e^{i2\pi k x} dx & = & 2 i^n j_n(2\pi k)\\
-2 i^n \int_{-\infty}^{\infty} j_n(2\pi k) e^{-i2\pi x k} dk &=& \mathscr{P}_n(x)
-\end{eqnarray}
-where the spherical Bessel functions of the first kind are given by
+The Fourier transform of the Legendre polynomials is\footnote{see \url{https://dlmf.nist.gov/18.17\#v} 18.17.19} \jrmadd{[simplified:]}
 \begin{equation}
-j_n(k) = \sqrt{\frac{\pi}{2k}}J_{n+\frac{1}{2}}(k)
+\int_{-1}^{1} \mathscr{P}_n(x) e^{i2\pi k x} dx  =  i^n \frac{J_{n+\frac{1}{2}}(2\pi k)}{\sqrt{k}}
 \end{equation}
-with $J_n$ denoting the Bessel functions of the first kind.  So we have the Fourier transforms of the $P_n(x)$ 
+with $J_n$ denoting the Bessel functions of the first kind.  We can then write the Fourier transforms of the orthonormal $P_n(x)$ as 
 \begin{equation}
@@
 \end{equation}
-Since we have normalized the Legendre polynomials, we have the useful result
+which have the property that \jrmadd{[changed limits]}
 \begin{equation}
-\int_0^\infty \left| Q_n(k) \right|^2 dk = 1 \mbox{ for all }n.
+\int_{-\infty}^\infty \left| Q_n(k) \right|^2 dk = 1 \mbox{ for all }n.
 \end{equation}
@@
 \end{equation}
-where
+where \jrmadd{the autocorrelation is given by}
 \begin{equation}
@@
 \end{equation}
-where for a wide-sense stationary process,
+where for a wide-sense stationary process
 \begin{equation}
@@
 
-We now have
+We now have \jrmadd{the covariances of the Legendre coefficients}
 \begin{equation}
@@
 \begin{equation}
-\left\langle \left|a_n\right|^2 \right\rangle = \frac{2n+1}{T}  \int_0^{\infty} \frac{ J_{n+\frac{1}{2}}^2(2\pi \kappa)}{\kappa} \mathcal{T}\left( \frac{2}{T} \kappa \right) d\kappa
+\left\langle \left|a_n\right|^2 \right\rangle = \frac{2n+1}{T}  \int_0^{\infty} \frac{ J_{n+\frac{1}{2}}^2(2\pi \kappa)}{\kappa} \mathcal{T}\left( \frac{2}{T} \kappa \right) d\kappa.
 \label{eqn:varcoeff}

### 2021-03-21 specklelives.tex
partial update with some comments from Rus

@@
 \end{equation}
-where $\Ji$ is the Jinc function
+at location $\vec{r} = r_u \hat{\mathpzc{u}} + r_v\hat{\mathpzc{v}}$. $\Ji$ is the Jinc function
 \begin{equation}
@@
 \end{equation}
-The cross-terms are a series of terms including products of $h_{mn}^+h_{mn}^-$ and $h_{mn}^ph_{m'n'}^{p'}$ ($mn \neq m'n'$) with $\Ji(x)\Ji(x')$.
-As long as the aperture truncation is accounted for when the long-exposure statistics are calculated,  these small cross-terms vanish and the long exposure intensity is
+\jrmadd{and the modified spatial frequencies are given by} 
+\begin{eqnarray}
+k_{mn}^+ &=& \sqrt{\left(\frac{r_u}{\lambda} + \frac{m}{d}\right) + \left(\frac{r_v}{\lambda} + \frac{n}{d}\right)} \\ 
+k_{mn}^- &=& \sqrt{\left(\frac{r_u}{\lambda} - \frac{m}{d}\right) + \left(\frac{r_v}{\lambda} - \frac{n}{d}\right)}. \nonumber
+\end{eqnarray}
+\jrmadd{Equation \ref{eqn:cross_terms} is equivalent to a convolution with the PSF, adding contributions from nearby speckles.}  The cross-terms are a series of terms including products of $h_{mn}^+h_{mn}^-$ and $h_{mn}^ph_{m'n'}^{p'}$ ($mn \neq m'n'$) with $\Ji(x)\Ji(x')$.
+As long as the aperture truncation is accounted for when the long-exposure statistics of $h$ are calculated, the sum can be dropped and these small cross-terms vanish.
+The long exposure intensity is
 \begin{equation}
@@
 On the 6.5 m aperture using the SI controller, windspeed has a strong effect, but on the larger ELT scale apertures wind has no significant effect.
-We note, that given the range of plausible mean windspeeds, the ELTs are in a different regime of $\mathcal{V} < D$.
-That is the crossing time $D/\mathcal{V}$ > 1 s.
-Were a mean windspeed significantly higher than 30 m/s to occur, the speckle lifetimes on the ELT will be shorter.
+Given the range of plausible mean windspeeds, \jrmadd{the ELTs appear to be in a different regime}.
+Were a mean windspeed significantly higher than 30 m/s to occur, the speckle lifetimes on the ELT would be shorter.
 
@@
 Diameter has a weak effect, with lifetimes dropping as $D$ increases.
-When combined as ``crossing time'' $D/V$ there is some evidence for the $\tau \propto D/V$ relation predicted by \citet{2005SPIE.5903..170M} on smaller apertures, but this does not hold for larger telescopes and does not apply to predictive control.
+When combined as ``crossing time'' $D/V$ there is some evidence for the $\tau \propto D/\mathcal{V}$ relation predicted by \citet{2005SPIE.5903..170M} on smaller apertures, but this does not hold for larger telescopes and does not apply to predictive control.
 
@@
 \caption{Changes in speckle lifetime with turbulence parameters.
-Here absolute spatial frequency is held constant.
-The $(m,n)=(10,14)$ mode on the 6.5 m telescope is compared to the $(m,n)=(60,84)$ mode on the 39 m aperture.
+Here absolute spatial frequency is held constant \jrmadd{for all aperture sizes}.
+\jrmadd{For instance}, the $(m,n)=(10,14)$ mode on the 6.5 m telescope is compared to the $(m,n)=(60,84)$ mode on the 39 m aperture.
 This means the input temporal PSD is identical in shape, though the integrated variance changes $\propto D^{-2}$.
@@
 Under SI control, speckle lifetime increases with $D$, but shows little change under LP control.
-As $D/V$, we see that there is a scaling up to $D/\mathcal{V} \sim 2$ for the SI, but this breaks down at longer crossing times (slower winds).
+There is a scaling up to $D/\mathcal{V} \sim 2$ for the SI, but this breaks down at longer crossing times (slower winds).
 For the LP controller speckle lifetime is essentially flat vs.

### 2021-03-21 specklelives.tex
rewrites

@@
 High contrast imaging observations are fundamentally limited by the spatially and temporally correlated noise source called speckles.
-Suppression and removal of speckles are the key goals of wavefront control and adaptive optics (AO), coronagraphy, and a host of post-processing techniques.
-The formation of speckles as spatial components of the instantaneous point-spread function (PSF) is well understood.
-However, while many empirical and simulation studies have been carried out, there has been relatively little effort focused on a theoretical understanding of the time-domain behavior of the intensity behind a coronagraphic wavefront control system.
+Suppression of speckles is the key goal of wavefront control and adaptive optics (AO), coronagraphy, and a host of post-processing techniques.
 Here we present a method for calculating the temporal power spectral density (PSD) of the intensity, and then show how the PSD can be used to calculate the statistical speckle lifetime.
-Considering a frozen-flow turbulence model, we use these techniques to analyze the speckle lifetimes behind a MagAO-X-like AO system and extremely large telescope (ELT) scale systems.
+Speckle-limited integration time is directly proportional to this lifetime.
+Considering a frozen-flow turbulence model, we analyze the speckle lifetimes behind a MagAO-X-like AO system and extremely large telescope (ELT) scale systems.
 We find that standard AO control shortens atmospheric speckle lifetime from $\sim$130 ms to $\sim$50 ms, and predictive control will further shorten the lifetime to $\sim$20 ms on MagAO-X.
-Speckle-limited integration time is directly proportional to these lifetimes.
-While the lifetime varies with telescope diameter, windspeed, and turbulence strength, there are no simple scaling laws among these quantities.
-Finally, we describe how the intensity PSD can be used to explore the fundamental limits of post-processing in high contrast imaging.
-These results motivate continued effort in the area of wavefront sensor and telemetry based reconstruction and speckle sensing for PSF subtraction, which promise to eliminate the long temporal correlations which currently limit high contrast imaging.
+Finally, we use these results to describe how ground-based telescopes will achieve the photon-noise limit in high contrast imaging.
 \end{abstract}
+
+%\jrmrmv{The formation of speckles as spatial components of the instantaneous point-spread function (PSF) is well understood.
+However, while many empirical and simulation studies have been carried out, there has been relatively little effort focused on a theoretical understanding of the time-domain behavior of the intensity behind a coronagraphic wavefront control system.}
 
@@
 Both in space and on the ground, imperfections within the telescope and instrument optics cause speckles.
-For \ogadd{ground-based} telescopes \ogrmv{on the Earth's surface}, the aberrations \ogadd{also} include those imposed by \ogadd{atmospheric} turbulence \ogrmv{in the Earth's atmosphere}.
+For \ogadd{ground-based} telescopes \ogrmv{on the Earth's surface}, the aberrations \ogadd{also} include those imposed by \ogadd{atmospheric} turbulence.
+\ogrmv{in the Earth's atmosphere}
 
@@
 In reality, the temporal behavior of atmospheric speckles falls between these regimes.
-These speckles vary on short time-scales \ogrmv{making standard PSF-subtraction essentially impossible}\ogadd{so they do are not removed by standard PSF subtraction}, but exhibit non-negligible temporal correlations which cause them to average much more slowly than photon noise.
+These speckles vary on short time-scales \ogrmv{making standard PSF-subtraction essentially impossible}\ogadd{so they are not removed by standard PSF subtraction}, but exhibit non-negligible temporal correlations which cause them to average much more slowly than photon noise.
 The statistical speckle lifetime, that is the time-scale which controls the averaging of speckle noise, is the quantity that sets the fundamental sensitivity limit of ground-based high-contrast imaging.
 
+\subsection{Atmospheric Coherence}
 There have been several definitions of speckle lifetime presented in the literature.
@@
 A common definition for this coherence time was based on a $1/e$ reduction in the autocorrelation of the intensity, which was experimentally found to be on the order of a few ms \citep{1978ApOpt..17.3779S,1990JMOp...37.1247D}.
-\citet{1982JOpt...13..263R} derived the following expression for the autocorrelation e-folding time in multi-layer Kolmogorov atmosphere:
+\citet{1982JOpt...13..263R} derived the following expression for the autocorrelation e-folding time in a multi-layer Kolmogorov atmosphere:
 \begin{equation}
@@
 
+\subsection{Variance of the Mean}
 The autocorrelation e-folding time is a somewhat arbitrary choice.
@@
 \end{equation}
-Here $\tau$ is given by the integral of the autocorrelation \citep{2006ApJ...637..541F}, and in this expression determines the number of uncorrelated realizations of the speckle pattern included in the mean.
+\jrmadd{This simply states that the error in a measurement of the mean improves with the number of uncorrelated trials $t/\tau$.
+Defined this way,} $\tau$ is given by the integral of the autocorrelation \citep{2006ApJ...637..541F}, and in this expression determines the number of uncorrelated realizations of the speckle pattern included in the mean.
 Note that this assumes $t >> \tau$.
@@
 \end{equation}
-for use in Equation (\ref{eqn:varmean_def_intro}).
+for use in Equation (\ref{eqn:varmean_def_intro})\footnote{\citet{2005SPIE.5903..170M} reported a speckle decorrelation time which would be a factor of two longer than the statistical lifetime}.
 This result is quite different from the previous predictions, but coarsely in agreement with the experiments by \citet{2006ApJ...637..541F} (it is challenging to make more definitive experimental comparisons due to the uncertainty in atmospheric parameters).
@@
 
-In the present study, we develop a semi-analytic method for predicting the residual speckle lifetime according to Equation (\ref{eqn:varmean_def_intro}).
+\subsection{Goals of This Study}
+In this study, we seek to describe the time-domain behavior of speckle noise in AO-fed coronagraphic images and its impact on sensitivity from as close to first principles as possible.
+This requires a traversal of intensity statistics, signal processing, wavefront propagation, and control theory, and so is somewhat involved.
+To accomplish our goals we will work through the following tasks, organized by section:
+\begin{itemize}
+\item Section \ref{sec:noise}: Review speckle statistics to understand how atmospheric speckles dramatically reduce sensitivity. \\
+\item Section \ref{sec:psd_lifetime}: Develop an analytic framework for analyzing power spectral densities (PSDs), which allows us to derive an expression for the statistical (mean-averaging) lifetime from a PSD.
+\item Section \ref{sec:intensity}: Show how to calculate the intensity PSD in the coronagraphic focal plane behind a closed-loop AO system with a semi-analytic model.
+\item Section \ref{sec:results}: Employ an end-to-end AO system simulation to (a) validate the statistical lifetime calculations and (b) assess the performance of the semi-analytic model.
+\item Section \ref{sec:trends}: Use the analytic framework to analyze speckle lifetime over a range of telescope diameters and atmospheric parameters.
+\item Section \ref{sec:reconstruction}: We propose a method to analyze post-processing in the time-domain and discuss its implications for the limits of ground-based high-contrast imaging sensitivity.
+\item Section \ref{sec:discussion}: We then discuss various implications of these results.
+\end{itemize}
+For the impatient reader, Sections \ref{sec:psd_lifetime} through \ref{sec:results} and Section \ref{sec:reconstruction} contain the technical details.
+The key points can be found in Sections \ref{sec:noise}, \ref{sec:trends}, and \ref{sec:discussion}.
+
+%In the present study, we develop a semi-analytic method for predicting the residual speckle lifetime according to Equation (\ref{eqn:varmean_def_intro}).
 This model explicitly accounts for closed-loop AO control, including advanced predictive control laws.
@@
 \section{Noise Dynamics in Coronagraphic Focal Planes}
-
-The quantity of interest in this analysis is the intensity in the focal plane as a function of time, $I(\vec{r}, t)$, where $\vec{r}$ is the position.
-Useful units for this analysis are photons/sec/$(\lambda/D)^2$.
-Following \citet{2007ApJ...669..642S} we consider three contributions to the intensity, such that
-\begin{equation}
-I(\vec{r},t) = I_{*}(t)\left[I_c(\vec{r},t) + I_{as}(\vec{r},t) + I_{qs}(\vec{r},t)\right]
-\end{equation}
-where $I_{*}$ is the peak intensity of the star's PSF without a coronagraph, itself a function of time due to Strehl ratio variation.
+\label{sec:noise}
+\jrmadd{The key metric for sensitivity in a high-contrast imaging observation is the signal-to-noise ratio, S/N.  The signal we collect from a planet, which has planet to star contrast of $C_p$, is }
+\begin{equation}
+\mbox{S}_p = I_{*}(t) C_p \Delta t
+\end{equation}
+\jrmadd{where $I_{*}$ is the peak intensity of the star's PSF without a coronagraph, which is a function of time due to Strehl ratio variation, and $\Delta t$ is the exposure time.
+Here we will assume units of photons/$(\lambda/D)^2$/second for the intensity.}
+
+\jrmadd{The dominant noise source in such an observation is the halo of scattered starlight underlying the planet signal.
+At the planet position $\vec{r}_p$ this has intensity $I_h(\vec{r}_p,t)$.
+Following \citet{2007ApJ...669..642S} we consider three contributions to the halo intensity:}
+\begin{equation}
+I_h(\vec{r}_p,t) = I_{*}(t)\left[I_c(\vec{r},t) + I_{as}(\vec{r},t) + I_{qs}(\vec{r},t)\right]
+\end{equation}
 $I_c$ is the fractional contribution from residual diffraction from the coronagraph, as well as purely static aberrations.
 $I_{as}$ is the fractional contribution from residual atmospheric speckles, and $I_{qs}$ is the fractional contribution from quasi-static (QS) speckles.
-The instantaneous raw contrast is
-\begin{equation}
-C(\vec{r},t) = \frac{I(\vec{r},t)}{I_*(t)} = I_c(\vec{r},t) + I_{as}(\vec{r},t) + I_{qs}(\vec{r},t).
-\end{equation}
-
-The total variance, considering both speckles and and photon noise, \jrmadd{at a point in the focal plane (fp)} is \citep{2007ApJ...669..642S} 
-\begin{equation}
-\sigma^2_{fp}  = I_* \Delta t [\underbrace{I_c + I_{as} + I_{qs}}_{\mbox{photon noise}} + I_* [ \underbrace{\tau_{as}\left(I_{as}^2 + 2[I_cI_{as} + I_{as}I_{qs}]  \right)}_{\mbox{atm.
+The halo contributes noise in two distinct ways: photon noise and correlated speckle noise.
+%From this we can state the instantaneous raw halo contrast
+%\begin{equation}
+%C_h(\vec{r},t) = \frac{I(\vec{r},t)}{I_*(t)} = I_c(\vec{r},t) + I_{as}(\vec{r},t) + I_{qs}(\vec{r},t).
+%\end{equation}
+The total variance, considering both speckles and and photon noise, \jrmadd{under the planet signal} is \citep{2007ApJ...669..642S} 
+\begin{equation}
+\sigma^2_{p}  = I_* \Delta t [\underbrace{I_c + I_{as} + I_{qs}}_{\mbox{photon noise}} + I_* [ \underbrace{\tau_{as}\left(I_{as}^2 + 2[I_cI_{as} + I_{as}I_{qs}]  \right)}_{\mbox{atm.
 speckles}} +  \underbrace{\tau_{qs}\left( I_{qs}^2 + 2 I_c I_{qs}\right)}_{\mbox{QS speckles}}]]
@@
 \end{equation}
-where $\tau_{as}$ is the lifetime of the residual atmospheric speckles, and $\tau_{qs}$ is the lifetime of the QS speckles.
-$\Delta t$ is the exposure time.
-For simplicity we have suppressed the $\vec{r}$ dependence.
+where $\tau_{as}$ is the lifetime of the residual atmospheric speckles, and $\tau_{qs}$ is the lifetime of the QS speckles.  For simplicity we have suppressed the $\vec{r}$ dependence.
 
@@
 \begin{equation}
-\sigma^2_{fp}  = I_* \Delta t \left(I_{as} + I_* \tau_{as} I_{as}^2 \right)
+\sigma^2_{p}  = I_* \Delta t \left(I_{as} + I_* \tau_{as} I_{as}^2 \right)
 \label{eqn:variance_as}
 \end{equation}
-If we attempt to observe a planet, with a contrast with respect to the star of $C_p$, then the peak signal-to-noise ratio of the planet detection will be
+
+Now the peak signal-to-noise ratio of the planet detection will be
 \begin{equation}
@@
 
-\subsection{Speckle Lifetime From A PSD}
+\section{Speckle Lifetime From A PSD}
 \label{sec:psd_lifetime}
@@
 This means that the PSD describes how the measurements derived from the process are correlated.
-\jrmadd{\citet{1976JOSA...66..207N} exploited this concept to derive the coefficient covariances in the Zernike polynomial expansion of Kolmogorov phase screens.
-Here we adapt that analysis to arbitrary 1-dimensional PSDs using the Legendre polynomials.}
+\jrmadd{\citet{1976JOSA...66..207N} exploited this concept in two dimensions to derive the coefficient covariances in the Zernike polynomial expansion of Kolmogorov phase screens.
+Here we adapt that analysis to arbitrary one dimensional PSDs using the Legendre polynomials.}
 
@@
 \section{Speckle Intensity PSDs}
-
+\label{sec:intensity}
 \subsection{Intensity as a Function of Time}
@@
 \subsection{A Recipe for Intensity PSDs}
-\label{sect:recipe}
+\label{sec:recipe}
 We are now ready to calculate $\mathcal{T}_{I_{mn}}(f)$.
@@
 \section{Results}
+\label{sec:results}
 
@@
 \subsection{Simulations}
-\label{sect:simulations}
+\label{sec:simulations}
 We conducted a series of end-to-end numerical simulations of an idealized MagAO-X-like instrument to validate our calculations.
@@
 \includegraphics[width=3.54in]{contrast8mag.pdf}
-\caption{Comparison of long-exposure raw contrast from semi-analytic calculations (Section \ref{sect:recipe}) and end-to-end simulation (Section \ref{sect:simulations}) for a 6.5 m telescope.
+\caption{Comparison of long-exposure raw contrast from semi-analytic calculations (Section \ref{sec:recipe}) and end-to-end simulation (Section \ref{sec:simulations}) for a 6.5 m telescope.
 Left: without photon noise, considering the input turbulence only.
@@
 \caption{Comparison of predicted and simulated variance evolution for speckle intensity at position (m=10,n=14).
-\jrmadd{Here ``calculated'' is the output of the method in Section \ref{sect:recipe}.
+\jrmadd{Here ``calculated'' is the output of the method in Section \ref{sec:recipe}.
 The ``sim.
@@
 
-\subsection{Trends in Speckle Lifetime}
+\section{Trends in Speckle Lifetime}
 \label{sec:trends}
 
-We now investigate the behavior of speckle lifetime over range of parameters.
+We now investigate the behavior of speckle lifetime over a range of parameters.
 The first comparison we make is between control laws.
 Here we simply calculate a histogram of the lifetimes over the 48x48 $\lambda/D$ dark hole for the SI and LP controller.
-We also apply the Monte Carlo procedure to the OL case.
+We also apply the Monte Carlo procedure to the OL case, \jrmadd{by treating the input PSDs as unmodified by control laws.
+We note that this is not strictly valid as the input phase is then outside the approximation regime made in the derivation of Equation (\ref{eqn:cross_terms}).
+However it provides a useful comparison of the change in the distribution of power in the PSD caused by closed-loop control action.
+}
+
 The three resulting histograms are shown in Figure \ref{fig:slHist}.
@@
 This is because increasing $r_0$ lowers the overall power at each spatial frequency, which in turn causes the control system to perform better at lower temporal frequencies.
-This better low-frequency rejection lowers the speckle lifetime.
+This better low-frequency rejection therefore lowers the speckle lifetime.
 
@@
 \section{The Potential of Reconstruction Post-Processing}
+\label{sec:reconstruction}
 In Section \ref{sec:psd_lifetime} we developed the expansion of the intensity time-series in terms of orthonormal Legendre polynomials $P_n$.
@@
 \section{Discussion}
+\label{sec:discussion}
 For some time, long-lived quasi-static speckles have been recognized as the limiting noise source in contemporary high contrast imaging observations \citep[e.g.][]{2007ApJ...654..633H}.
@@
 This precludes conventional post-processing, but the lifetimes are long enough to significantly increase exposure times compared to the photon-noise.
+
+\subsection{Caveats}
+The results shown here are subject to several caveats.
+Our analysis is monochromatic.
+While we assumed that science and sensing wavelength were the same, neither component of an AO system can be truly monochromatic.
+Though speckle lifetime form Atmsopheric chromaticity will cause differential errors across the bandpass of the WFS
 

### 2021-03-21 specklelives.tex
various updates

@@
 Suppression of speckles is the key goal of wavefront control and adaptive optics (AO), coronagraphy, and a host of post-processing techniques.
-Here we present a method for calculating the temporal power spectral density (PSD) of the intensity, and then show how the PSD can be used to calculate the statistical speckle lifetime.
-Speckle-limited integration time is directly proportional to this lifetime.
-Considering a frozen-flow turbulence model, we analyze the speckle lifetimes behind a MagAO-X-like AO system and extremely large telescope (ELT) scale systems.
-We find that standard AO control shortens atmospheric speckle lifetime from $\sim$130 ms to $\sim$50 ms, and predictive control will further shorten the lifetime to $\sim$20 ms on MagAO-X.
-Finally, we use these results to describe how ground-based telescopes will achieve the photon-noise limit in high contrast imaging.
+The temporal correlation of speckles reduces sensitivity as they average much more slowly than photon noise.
+Speckles average at a rate set by the statistical speckle lifetime, and speckle-limited integration time in long exposures is directly proportional to this lifetime.
+Here we present a novel power spectral density (PSD) method for calculating the lifetime, and develop a semi-analytic method for predicting intensity PSDs behind a coronagraph.
+Considering a frozen-flow turbulence model, we analyze the residual atmosphere speckle lifetimes in a MagAO-X-like AO system as well as 25-39 m giant segment mirror telescope (GSMT) scale systems.
+We find that standard AO control shortens atmospheric speckle lifetime from $\sim$130 ms to $\sim$50 ms, and predictive control will further shorten the lifetime to $\sim$20 ms on 6.5 m MagAO-X.
+We find that speckle lifetimes vary somewhat with diameter, windspeed, and seeing, but stay within at range of 20 ms 20 100 ms.
+Due to control system dynamics there are no simple scaling laws which apply across a wide range of system characteristics.
+Finally, we use these results to describe how post-processing will enable ground-based telescopes to achieve the photon-noise limit in high contrast imaging.
 \end{abstract}
-
-%\jrmrmv{The formation of speckles as spatial components of the instantaneous point-spread function (PSF) is well understood.
-However, while many empirical and simulation studies have been carried out, there has been relatively little effort focused on a theoretical understanding of the time-domain behavior of the intensity behind a coronagraphic wavefront control system.}
 
@@
 
+\jrmadd{[MOVED FROM DISCUSSION] For some time, long-lived ``quasi-static'' speckles have been recognized as the limiting noise source in contemporary high contrast imaging observations \citep[e.g.][]{2007ApJ...654..633H}.
+Given their origin within the instrument, and comparatively long lifetimes \citep{2012A&A...541A.136M}, such speckles should be supressable using wavefront sensing and control (WFS\&C) strategies.
+Recent progress in applying ``dark hole'' contrast optimization with focal plane WFS \citep{2020A&A...638A.117P} and various focal-plane \citep[e.g.][]{2010A&A...509A..31G} and low-order coronagraphic WFS\&C strategies \citep[e.g.][]{2017PASP..129i5002S} show great promise in significantly reducing the impact of internal speckles.
+Once active quasi-static speckle control becomes routine, the fundamental sensitivity limit in ground-based high-contrast imaging will be set by the Earth's atmosphere.
+}
+
+%This limit is set by two components: uncorrelated photon noise and the spatially and temporally correlated speckle noise.
+The temporal correlations of atmospheric speckles are, compared to quasi-static speckles, relatively short.
+This precludes conventional post-processing, but the lifetimes are long enough to significantly increase exposure times compared to the photon-noise.
+
 The index of refraction of air depends on temperature, and turbulence causes the air temperature to stochastically vary across the telescope aperture.
@@
 Even with so-called ``extreme'' AO systems \citep[ExAO, ][]{2018ARAA..56..315G} this correction is not perfect.
-The AO-corrected wavefront, though significantly improved, still causes a halo of speckles significantly brighter than the instrumental speckles in a modern high performance coronagraph system using wavefront control.
+The AO-corrected wavefront, though significantly improved, still causes a halo of speckles significantly brighter than the instrumental speckles in a modern high performance coronagraph system using coronagraphic wavefront control.
 
@@
 In the first regime, the speckles are perfectly static.
-Were this case, it would be a simple matter of measuring the PSF structure and then subtracting it from all images.
+Were this the case, it would be a simple matter of measuring the PSF structure and then subtracting it from all images.
 While this would still leave the photon noise from the scattered light, we could perfectly remove the spatially correlated noise.
@@
 These speckles vary on short time-scales \ogrmv{making standard PSF-subtraction essentially impossible}\ogadd{so they are not removed by standard PSF subtraction}, but exhibit non-negligible temporal correlations which cause them to average much more slowly than photon noise.
-The statistical speckle lifetime, that is the time-scale which controls the averaging of speckle noise, is the quantity that sets the fundamental sensitivity limit of ground-based high-contrast imaging.
+The statistical lifetime of atmospheric speckles, that is the time-scale which controls the averaging of speckle noise, is the quantity that sets the fundamental sensitivity limit of ground-based high-contrast imaging.
 
 \subsection{Atmospheric Coherence}
-There have been several definitions of speckle lifetime presented in the literature.
+There have been several definitions of the atmospheric speckle lifetime presented in the literature.
 Early treatments of speckle dynamics were motivated by speckle interferometry, which relies on exposures short enough to ``freeze the atmosphere''.
@@
 At good astronomical sites \ogrmv{such as MKO and LCO}, values of $\tau_{boil}$ should range from $\sim$5 to $\sim$20 ms.
-\citet{1999PASP..111..587R} employed Equation (\ref{eqn:e-fold}) for the speckle lifetime and showed that speckles dominate photon noise by factors of 100 or more in high contrast imaging.
+\citet{1999PASP..111..587R} employed Equation (\ref{eqn:e-fold}) for the speckle lifetime and argued that atmospheric speckles dominate photon noise by factors of 100 or more in high contrast imaging.
 
@@
 The autocorrelation e-folding time is a somewhat arbitrary choice.
-If we instead consider the evolution of the variance in the estimate of the mean given uncorrelated trials, we can define the speckle lifetime according to
+If we instead consider the evolution of the variance in the estimate of the mean given uncorrelated trials, we can define the speckle lifetime \jrmadd{according to the familiar ``root-N averaging'' reduction in noise:}
 \begin{equation}
@@
 \end{equation}
-\jrmadd{This simply states that the error in a measurement of the mean improves with the number of uncorrelated trials $t/\tau$.
+\jrmadd{This simply states that the variance in a measurement of the mean improves with the number of uncorrelated trials $t/\tau$.
 Defined this way,} $\tau$ is given by the integral of the autocorrelation \citep{2006ApJ...637..541F}, and in this expression determines the number of uncorrelated realizations of the speckle pattern included in the mean.
@@
 Using AO corrected images, \citet{2006ApJ...637..541F} empirically found significantly longer values of $\tau$, giving an overall median of 175 ms at Lick observatory.
-Of note, the results of \citet{2006ApJ...637..541F} hinted that inside the region of the image plane controlled by the AO system, the speckle lifetime was significantly shorter.
-
-\citet{2005SPIE.5903..170M} considered the time averaging of residual atmospheric speckles behind a coronagraph, directly analyzing the reduction in variance as a function of exposure time.
+Of note, the results of \citet{2006ApJ...637..541F} hinted that inside the region of the image plane controlled by the AO system, the speckle lifetime was significantly shorter than outside.
+
+\citet{2005SPIE.5903..170M} considered this time averaging of residual atmospheric speckles behind a coronagraph, directly analyzing the reduction in variance as a function of exposure time.
 They argued that the speckle lifetime should not be affected by the properties of the AO system nor by the seeing parameter $r_0$, and instead should be determined by the wind-crossing time of a telescope with diameter $D$.
@@
 
+\jrmadd{NOTE: need to add discussion of Poyneer and Mac 2006}
 %Note for comparison that the Greenwood time-constant is
@@
 \subsection{Goals of This Study}
-In this study, we seek to describe the time-domain behavior of speckle noise in AO-fed coronagraphic images and its impact on sensitivity from as close to first principles as possible.
-This requires a traversal of intensity statistics, signal processing, wavefront propagation, and control theory, and so is somewhat involved.
+\jrmadd{Our goal in this study is to develop a self-consistent description of the time-domain behavior of speckle noise in AO-fed coronagraphic images from as close to first principles as possible.
+Our motivation is to understand the true limits of ground-based high contrast imaging and so we focus on the statistical lifetime of atmospheric speckles in coronagraphic focal planes.
+This requires a traversal of speckle statistics, signal processing, wavefront propagation, and control theory, and so is somewhat involved.
 To accomplish our goals we will work through the following tasks, organized by section:
 \begin{itemize}
-\item Section \ref{sec:noise}: Review speckle statistics to understand how atmospheric speckles dramatically reduce sensitivity. \\
+\item Section \ref{sec:noise}: Review speckle statistics to understand how atmospheric speckles dramatically reduce sensitivity.
 \item Section \ref{sec:psd_lifetime}: Develop an analytic framework for analyzing power spectral densities (PSDs), which allows us to derive an expression for the statistical (mean-averaging) lifetime from a PSD.
@@
 \end{itemize}
-For the impatient reader, Sections \ref{sec:psd_lifetime} through \ref{sec:results} and Section \ref{sec:reconstruction} contain the technical details.
-The key points can be found in Sections \ref{sec:noise}, \ref{sec:trends}, and \ref{sec:discussion}.
+The key conclusions can be gleaned from Sections \ref{sec:noise}, \ref{sec:trends}, and \ref{sec:discussion}.
+Sections \ref{sec:psd_lifetime} through \ref{sec:results} and Section \ref{sec:reconstruction} contain the technical details.
+}
 
@@
 \begin{equation}
-\mbox{S}_p = I_{*}(t) C_p \Delta t
+\mbox{Planet Signal} = I_{*}(t) C_p \Delta t
+\label{eqn:signal}
 \end{equation}
 \jrmadd{where $I_{*}$ is the peak intensity of the star's PSF without a coronagraph, which is a function of time due to Strehl ratio variation, and $\Delta t$ is the exposure time.
-Here we will assume units of photons/$(\lambda/D)^2$/second for the intensity.}
+Here we will assume units of photons/sec/$(\lambda/D)^2$ for the intensity, where $\lambda$ is the observation wavelength and $D$ is the telescope diameter.
+$\lambda/D$ is the fundamental spatial scale of diffraction limited imaging and coronagraphy, and we assume that photometry is conducted in pixels of size 1 $\lambda/D$.}
 
@@
 At the planet position $\vec{r}_p$ this has intensity $I_h(\vec{r}_p,t)$.
-Following \citet{2007ApJ...669..642S} we consider three contributions to the halo intensity:}
-\begin{equation}
-I_h(\vec{r}_p,t) = I_{*}(t)\left[I_c(\vec{r},t) + I_{as}(\vec{r},t) + I_{qs}(\vec{r},t)\right]
+Following \citet{2007ApJ...669..642S} we initially consider three contributions to the halo intensity:}
+\begin{equation}
+I_h(\vec{r}_p,t) = I_{*}(t)\left[I_c(\vec{r}_p,t) + I_{as}(\vec{r}_p,t) + I_{qs}(\vec{r}_p,t)\right]
 \end{equation}
 $I_c$ is the fractional contribution from residual diffraction from the coronagraph, as well as purely static aberrations.
-$I_{as}$ is the fractional contribution from residual atmospheric speckles, and $I_{qs}$ is the fractional contribution from quasi-static (QS) speckles.
+$I_{as}$ is the fractional contribution from residual atmospheric speckles, and $I_{qs}$ is the fractional contribution from instrumental speckles, which are usually called quasi-static speckles.
+From this we can state the instantaneous raw halo contrast
+\begin{equation}
+C_h(\vec{r}_p,t) = \frac{I_h(\vec{r}_p,t)}{I_*(t)} = I_c(\vec{r}_p,t) + I_{as}(\vec{r}_p,t) + I_{qs}(\vec{r}_p,t).
+\end{equation}
+
 The halo contributes noise in two distinct ways: photon noise and correlated speckle noise.
-%From this we can state the instantaneous raw halo contrast
-%\begin{equation}
-%C_h(\vec{r},t) = \frac{I(\vec{r},t)}{I_*(t)} = I_c(\vec{r},t) + I_{as}(\vec{r},t) + I_{qs}(\vec{r},t).
-%\end{equation}
-The total variance, considering both speckles and and photon noise, \jrmadd{under the planet signal} is \citep{2007ApJ...669..642S} 
-\begin{equation}
-\sigma^2_{p}  = I_* \Delta t [\underbrace{I_c + I_{as} + I_{qs}}_{\mbox{photon noise}} + I_* [ \underbrace{\tau_{as}\left(I_{as}^2 + 2[I_cI_{as} + I_{as}I_{qs}]  \right)}_{\mbox{atm.
+The total halo variance, considering both speckles and and photon noise, \jrmadd{under the planet signal} is \citep{2007ApJ...669..642S}
+\begin{equation}
+\sigma^2_{h}  = I_* \Delta t [\underbrace{I_c + I_{as} + I_{qs}}_{\mbox{photon noise}} + I_* [ \underbrace{\tau_{as}\left(I_{as}^2 + 2[I_cI_{as} + I_{as}I_{qs}]  \right)}_{\mbox{atm.
 speckles}} +  \underbrace{\tau_{qs}\left( I_{qs}^2 + 2 I_c I_{qs}\right)}_{\mbox{QS speckles}}]]
@@
 \end{equation}
-where $\tau_{as}$ is the lifetime of the residual atmospheric speckles, and $\tau_{qs}$ is the lifetime of the QS speckles.  For simplicity we have suppressed the $\vec{r}$ dependence.
+where $\tau_{as}$ is the lifetime of the residual atmospheric speckles, and $\tau_{qs}$ is the lifetime of the instrumental quasi-static speckles.
+For simplicity we have suppressed the $\vec{r}_p$ dependence.
 
@@
 We could assume that $\Delta t$ is sufficiently short that the $I$s can be treated as constant, and then state the long exposure variance as a sum of the variance in these short exposures.
-Alternatively, we can treat the $I$s in Equation (\ref{eqn:variance}) as mean values, and then $\sigma^2_{fp}$ is the mean of the variance.
+Alternatively, we can treat the $I$s in Equation (\ref{eqn:variance}) as mean values, and then $\sigma^2_{h}$ is the mean of the variance.
 For large total exposure time these will converge.
@@
 These (optimistic) conditions eliminate the speckle pinning \citep{2001ApJ...558L..71B} terms in Equation (\ref{eqn:variance}).
-The variance is then
-\begin{equation}
-\sigma^2_{p}  = I_* \Delta t \left(I_{as} + I_* \tau_{as} I_{as}^2 \right)
+The halo variance is then due solely to atmospheric speckles:
+\begin{equation}
+\sigma^2_{h}  = I_* \Delta t \left(I_{as} + I_* \tau_{as} I_{as}^2 \right)
 \label{eqn:variance_as}
@@
 
-Now the peak signal-to-noise ratio of the planet detection will be
+\jrmadd{The noise is the square-root of the variance, so using Equations (\ref{eqn:signal}) and  (\ref{eqn:variance_as})} the peak signal-to-noise ratio of the planet detection will be
 \begin{equation}
@@
 \end{equation}
-Such an observation is therefore speckle noise limited, and the S/N improves with square-root of the number of realizations of the speckle pattern as $\sqrt{\Delta t/ \tau_{as}}$.
+Such an observation is therefore \emph{speckle noise limited}, and the S/N improves with square-root of the number of realizations of the speckle pattern as $\sqrt{\Delta t/ \tau_{as}}$.
 The ratio of the time to reach a given S/N when speckle limited, $\Delta t_{sl}$, to the time needed when photon noise limited, $\Delta t_{pn}$ is
@@
 
-According to the Wiener-Khinchin theorem, the PSD \jrmadd{is the Fourier transform of the autocorrelation} \jrmrmv{(they are a Fourier transform pair)}.
+According to the Wiener-Khinchin theorem, the PSD \jrmadd{is the Fourier transform of the autocorrelation}.
 This means that the PSD describes how the measurements derived from the process are correlated.
-\jrmadd{\citet{1976JOSA...66..207N} exploited this concept in two dimensions to derive the coefficient covariances in the Zernike polynomial expansion of Kolmogorov phase screens.
+\jrmadd{\citet{1976JOSA...66..207N} exploited this concept in two dimensions to derive the coefficient covariances in the Zernike polynomial expansion of Kolmogorov turbulence phase screens.
 Here we adapt that analysis to arbitrary one dimensional PSDs using the Legendre polynomials.}
-
-\jrmadd{In principle, any basis could be used.
-At present our main goal is analyze the statistics of the mean, which corresponds to the first Legendre polynomial.
-Similar to the Zernike polynomials, this basis has an intuitive progression through slope, quadratic, cubic, etc, modes.
-We later utilize these 1-deminsional modes to analyze post-processing as a time-domain filter.
-For that analysis it will be convenient to use a normalized form of Legendre polynomials defined as:}
-\begin{equation}
-P(\xi) = \sqrt{\frac{2n+1}{2}}\, \mathscr{P}_n(\xi)
-\label{eqn:basis}
-\end{equation}
-where $\mathscr{P}_n(\xi)$ are the Legendre polynomials defined in the usual manner \citep[cf.][]{Olver:2010:NHMF} on the domain $-1 \le \xi \le 1$.
-The first six of these functions are shown in Figure \ref{fig:onlegpol}.
-With our normalization we have the orthogonality condition
-\begin{equation}
-\int_{-1}^{1}P_n(\xi)P_{n'}(\xi) d\xi = \delta_{nn'}
-\label{eqn:orthonormal}
-\end{equation}
-where $\delta_{nn'}$ is the Kronecker delta.
 
@@
 
+\jrmadd{In principle, any basis could be used.
+Our main goal is analyze the statistics of the mean, which corresponds to the first Legendre polynomial.
+Similar to the Zernike polynomials, this basis has an intuitive progression through slope, quadratic, cubic, etc, modes (see Figure \ref{fig:onlegpol}).
+We later utilize these 1-deminsional modes to analyze post-processing as a time-domain filter.
+For that analysis it will be convenient to use a normalized form of Legendre polynomials defined as:}
+\begin{equation}
+P(\xi) = \sqrt{\frac{2n+1}{2}}\, \mathscr{P}_n(\xi)
+\label{eqn:basis}
+\end{equation}
+where $\mathscr{P}_n(\xi)$ are the Legendre polynomials defined in the usual manner \citep[cf.][]{Olver:2010:NHMF} on the domain $-1 \le \xi \le 1$.
+The first six of these functions are shown in Figure \ref{fig:onlegpol}.
+With our normalization we have the orthogonality condition
+\begin{equation}
+\int_{-1}^{1}P_n(\xi)P_{n'}(\xi) d\xi = \delta_{nn'}
+\label{eqn:orthonormal}
+\end{equation}
+where $\delta_{nn'}$ is the Kronecker delta.
+
 \jrmadd{Now we} assume that we obtain a sample $I(t)$ of length $T$ consisting of measurements uniformly spaced by $\Delta t$.
@@
 \end{equation}
+\jrmadd{The coefficients are variable quantities.
+Depending on the time $T$ under consideration, the best fit of the polynomials will change as the speckle intensity $I(t)$ varies.
+By analyzing the temporal statistics of the $a_n$, we learn about the power in specific Legendre ``modes'' just as we do with the analgous Zernike modes in a wavefront.}
+
 The first orthonormal Legendre polynomial is $P_0(\xi) = 1/\sqrt{2}$, which means that $a_0/\sqrt{2}$ is the mean value of $I(t)$.  Thus the variance of the mean is given by
@@
 \end{equation}
-Now from Equation (\ref{eqn:process_var}) the total variance\jrmrmv{, which gives the variance of a single measurement,} is
+Now from Equation (\ref{eqn:process_var}) \jrmadd{the variance of a single measurement is}
 \begin{equation}
@@
 \end{equation}
-where $N = t/\Delta t$ is the number of measurements.  This is exactly the result we expect for the variance of the mean in a white noise process with variance $\sigma^2_o$.
+where $N = t/\Delta t$ is the number of measurements.
+This is exactly the result we expect for the variance of the mean in a white noise process with variance $\sigma^2_o$, \jrmadd{giving the familiar ``root-N'' reduction in noise.}
 
@@
 \end{equation}
-\jrmrmv{As long as $t >> \tau$, }The variance of the mean decreases with time according to
-\begin{equation}
-\sigma_\mathrm{mean}^2 =  \frac{\sigma_o^2}{t/\tau}
+The variance of the mean decreases with time according to
+\begin{equation}
+\sigma_\mathrm{mean}^2 =  \frac{\sigma_o^2}{t/\tau} \tag{\ref{eqn:varmean_def_intro} restated}
 \end{equation}
@@
 
-The lifetime is mainly set by the lowest frequency components of the PSD.
-The limit in the numerator of Equation (\ref{eqn:psd_lifetime}) means that $f = \frac{2}{T}\kappa \rightarrow 0$, so we can approximate the lifetime as $\tau \approx \mathcal{T}(0)/(2\sigma_o^2)$.
+Equation (\ref{eqn:psd_lifetime}) shows that the statistical lifetime is mainly set by the lowest frequency components of the PSD.
+The limit in the numerator means that $f = \frac{2}{T}\kappa \rightarrow 0$, so we can approximate the lifetime as $\tau \approx \mathcal{T}(0)/(2\sigma_o^2)$.
 That is, the lifetime is set by the proportion of power in the 0-frequency component of the PSD.
@@
 \end{equation}
-at location $\vec{r} = r_u \hat{\mathpzc{u}} + r_v\hat{\mathpzc{v}}$. $\Ji$ is the Jinc function
+at location $\vec{r} = r_x \hat{x} + r_y\hat{y}$ where  $\hat{x}$ and  $\hat{y}$ are the focal plane coordinate unit vectors. $\Ji$ is the Jinc function
 \begin{equation}
@@
 \begin{eqnarray}
-k_{mn}^+ &=& \sqrt{\left(\frac{r_u}{\lambda} + \frac{m}{d}\right) + \left(\frac{r_v}{\lambda} + \frac{n}{d}\right)} \\ 
-k_{mn}^- &=& \sqrt{\left(\frac{r_u}{\lambda} - \frac{m}{d}\right) + \left(\frac{r_v}{\lambda} - \frac{n}{d}\right)}. \nonumber
+k_{mn}^+ &=& \sqrt{\left(\frac{r_x}{\lambda} + \frac{m}{d}\right) + \left(\frac{r_y}{\lambda} + \frac{n}{d}\right)} \\ 
+k_{mn}^- &=& \sqrt{\left(\frac{r_x}{\lambda} - \frac{m}{d}\right) + \left(\frac{r_y}{\lambda} - \frac{n}{d}\right)}. \nonumber
 \end{eqnarray}
@@
 As long as the aperture truncation is accounted for when the long-exposure statistics of $h$ are calculated, the sum can be dropped and these small cross-terms vanish.
-The long exposure intensity is
+With this caveat, the long exposure intensity is
 \begin{equation}
@@
 \citet{2018JATIS...4a9001M} showed how adding more coefficients, using the well-known Linear Prediction (LP) technique, results in significant performance improvements in terms of post-coronagraph contrast.
-Most, if not all, controllers can be cast in the form of Equation (\ref{eqn:linfilt}) \citep{2007JOSAA..24.2645P}.
+Most, if not all, controllers can be cast in the form of Equation (\ref{eqn:linfilt}) \citep{2007JOSAA..24.2645P, 2021arXiv210307566H}.
 
@@
 \end{eqnarray}
-Under the Taylor frozen flow hypothesis the amplitude $h_{mn}^\dagger$ of each spatial frequency is constant, which means that $h_{mn}^-(\phi_{mn}(t))$ = $h_{mn}^+(\phi_{mn}(t) + \frac{\pi}{2})$.
+Under the Taylor frozen flow hypothesis the amplitude $h_{mn}^\dagger$ of each spatial frequency is approximately constant over short periods, which means that $h_{mn}^-(\phi_{mn}(t))$ = $h_{mn}^+(\phi_{mn}(t) + \frac{\pi}{2})$.
 This phase shift yields
@@
 \end{equation}
-\jrmadd{This temporal-phase relationhip between the two modes at each spatial frequency must be included in the PSD calculation in order to obtain accurate speckle lifetimes}.\jrmrmv{which we employ next when calculating the PSDs of the intensity time-series.}
+\jrmadd{This temporal-phase relationship between the two modes at each spatial frequency must be included in the PSD calculation in order to obtain accurate speckle lifetimes}.
 
@@
 Figure \ref{fig:contrast_2Dcomp} shows post-coronagraph contrast maps, comparing the results of calculations using \citet{2018JATIS...4a9001M} to the simulation outputs.
-We show results for an $\infty$ mag guidestar to show the effects of the input turbulence alone, and for an $8$th mag guidestar.
+We show results for an $-\infty$ mag guidestar to show the effects of the input turbulence alone, and for an $8$th mag guidestar.
 Figure \ref{fig:contrast_radprof} shows the median radial profile for each of the maps in Figure \ref{fig:contrast_2Dcomp}.
 The agreement is very good, though some small differences are visible in the LP (bottom row).
-\jrmadd{These differences, apparent in both the maps and the profile, are likely due to the LP conroller optimization leading to slightly different performance due to the different modal basis actually under control in the simulations.
-Another possible source of these (slight) differences is spatial-temporal correlation not fully accounted for in the semi-analytic calculations.}
+\jrmadd{These differences, apparent in both the maps and the profile, are due to the LP conroller optimization leading to slightly different controller characteristics.
+This is due to the difference in the modal basis actually under control in the simulations compared to the model.
+Another possible source of slight differences is spatial-temporal correlation not fully accounted for in the semi-analytic calculations, which may become apparent at the correction levels under predictive control.}
 
@@
 The bottom row shows the ETF for each controller.
-On the $\infty$ mag guidestar, the OL PSD shows excellent agreement at low frequencies, with a slight divergence in slope beginning at a $\sim$100 Hz.
+On the $-\infty$ mag guidestar, the OL PSD shows excellent agreement at low frequencies, with a slight divergence in slope beginning at a $\sim$100 Hz.
 The SI likewise shows nearly perfect agreement, with the exception of a lower overshoot peak in the simulation.
@@
 
-The disagreements evident in the figure are likely due to the differences between the idealized continuous-time semi-analytic calculations and the discretized simulations.
-Additionally, the use of an orthogonalized basis in the simulation can be expected to produce different noise rejection characteristics on the 8th mag star.
+The disagreements evident in the figure are due to the differences between the idealized continuous-time semi-analytic calculations and the discretized simulations.
+\jrmadd{In particular the empirical optimization of the control laws produces different dynamical behavior in the simulations compared to the model predictions} Additionally, the use of an orthogonalized basis in the simulation produces different noise rejection characteristics on the 8th mag star.
 
@@
 Note that the colorbar is adjusted in each panel to best show detail.
-For the integrator agreement is very good.
+For the integrator the agreement is very good.
 For the predictor on the infinitely bright star, the agreement is good overall but the calculations \jrmadd{modestly} over-predict speckle lifetimes \jrmadd{in a wedge assocated with (but not parallel to)} the wind vectors.
@@
 \label{sec:discussion}
-For some time, long-lived quasi-static speckles have been recognized as the limiting noise source in contemporary high contrast imaging observations \citep[e.g.][]{2007ApJ...654..633H}.
-Given their origin within the instrument, and comparatively long lifetimes \citep{2012A&A...541A.136M}, such speckles should be supressable using wavefront sensing and control (WFS\&C) strategies.
-Recent progress in applying ``dark hole'' contrast optimization with focal plane WFS \citep{2020A&A...638A.117P} and various focal-plane \citep{2010A&A...509A..31G} and low-order coronagraphic WFS\&C strategites \citep[e.g.][]{2017PASP..129i5002S} show great promise in significantly reducing the impact of internal speckles.
-Assuming active quasi-static speckle control becomes routine, the fundamental sensitivity limit in ground-based high-contrast imaging is set by the Earth's atmosphere.
-This limit is set by two components: uncorrelated photon noise and the spatially and temporally correlated speckle noise.
-The temporal correlations of atmospheric speckles are, compared to quasi-static speckles, relatively short.
-This precludes conventional post-processing, but the lifetimes are long enough to significantly increase exposure times compared to the photon-noise.
 
@@
 Our analysis is monochromatic.
-While we assumed that science and sensing wavelength were the same, neither component of an AO system can be truly monochromatic.
-Though speckle lifetime form Atmsopheric chromaticity will cause differential errors across the bandpass of the WFS
+While we assumed that the science and sensing wavelength were the same, neither component of an AO system can be truly monochromatic.
+Though speckle lifetime formally does not depend on wavelength, which only scales the PSD without changing its shape, AO performance will depend on how light is distributed in the instrument.
+Furthermore, atmospheric chromaticity will cause differential errors across the bandpass of the WFS, and between the science and WFS bandpasses.
+These errors will contribute speckles in addition to those explicitly treated here, which will have different dynamics as they are not subject to the same control filters.
+These speckles can, however, be analyzed with the same tools we developed here.
+We save this for future work.
+
+A key assumption underlying the mathematics of the PSD analysis is that the speckle intensity is wide-sense stationary (WSS), which means that the mean is constant and that the autocorrelation does not depend on time.
+In this context, WSS essentially implies that the atmosphere parameters, namely $r_0$ and the layer $\vec{\mathcal{V}}$, are constant.
+This is of course not strictly true, but can be expected to be approximately true over short periods of time -- at least long enough to estimate a PSD.
+We also expect AO control to dampen changes, keeping the output speckle formation process within a more or less well defined range.
+What this ultimately means is that we must consider the speckle lifetime as itself a time-variable parameter, but there will be a representative value over an observation that characterizes long exposure sensitivity.
+
+\subsection{The Semi-analytic Model}
 
@@
 This powerful technique allows us to analyze the impact of control law choice, as well as to compare the influence of various telescope and atmosphere parameters on the speckle lifetime.
+
+\subsection{Trends in Speckle Lifetime}
+
 Of note, we find that the simple $D/\mathcal{V}$ ``crossing-time'' concept holds only for integrator control on contemporary $D\lesssim10$ m telescopes where $D/\mathcal{V} \lesssim 1$ second.
@@
 
-The temporal decorrelation of speckle noise has long been recognized as one of the main aspects of PSF-subtraction post-processing \citep{2006ApJ...641..556M}, even though algorithms have typically been focused on the spatial domain.
-Recently, various studies have begun to consider the time-domain in post-processing explicitly \citep{2021A&A...646A..24S}.
+\subsection{Post-Processing}
+The temporal decorrelation of speckle noise is one of the main benefits of PSF-subtraction post-processing \citep{2006ApJ...641..556M}, even though algorithms have typically been focused on the spatial domain.
+Recently, studies have begun to consider the time-domain in post-processing explicitly \citep{2021A&A...646A..24S}.
 Knowledge of the intensity PSD allows for such algorithms to be analyzed eficiently.
@@
 
+\subsection{Progress Towards Fundamental Limits}
+
 \section{Conclusion}
@@
 \appendix
-\section{The Correlation Length of a 1-D PSD}
+\section{The Noise Lifetime of a 1-D PSD}
 \label{app:tau_psd}

### 2021-03-23 specklelives.tex
big updates

@@
 High contrast imaging observations are fundamentally limited by the spatially and temporally correlated noise source called speckles.
-Suppression of speckles is the key goal of wavefront control and adaptive optics (AO), coronagraphy, and a host of post-processing techniques.
+Suppression of speckle \jrmadd{noise} is the key goal of wavefront control and adaptive optics (AO), coronagraphy, and a host of post-processing techniques.
 The temporal correlation of speckles reduces sensitivity as they average much more slowly than photon noise.
 Speckles average at a rate set by the statistical speckle lifetime, and speckle-limited integration time in long exposures is directly proportional to this lifetime.
+As progress continues in post-AO coronagraphic wavefront control, residual atmospheric speckles will become the limiting noise source in high contrast imaging, so a complete understanding of their statistical behavior is crucial to optimizing high contrast imaging instruments.
 Here we present a novel power spectral density (PSD) method for calculating the lifetime, and develop a semi-analytic method for predicting intensity PSDs behind a coronagraph.
-Considering a frozen-flow turbulence model, we analyze the residual atmosphere speckle lifetimes in a MagAO-X-like AO system as well as 25-39 m giant segment mirror telescope (GSMT) scale systems.
+Considering a frozen-flow turbulence model, we analyze the residual atmosphere speckle lifetimes in a MagAO-X-like AO system as well as 25-39 m giant segmented mirror telescope (GSMT) scale systems.
 We find that standard AO control shortens atmospheric speckle lifetime from $\sim$130 ms to $\sim$50 ms, and predictive control will further shorten the lifetime to $\sim$20 ms on 6.5 m MagAO-X.
-We find that speckle lifetimes vary somewhat with diameter, windspeed, and seeing, but stay within at range of 20 ms 20 100 ms.
+We find that speckle lifetimes vary with diameter, wind speed, seeing, and location with the AO control region (dark hole).
+No broadly applicable ``rules of thumb'' are found, but on bright stars lifetimes appear to remain within a rough range of $\sim$20 ms to $\sim$100 ms.
 Due to control system dynamics there are no simple scaling laws which apply across a wide range of system characteristics.
@@
 In addition to the photon noise caused by this halo, the scattered light is coherent and produces ``speckles'', which can be thought of as copies of the instrument point spread function (PSF) \citep{1995PASP..107..386M}.
-This adds structured noise, with both spatial and temporal correlations, which is the limiting noise source in \ogrmv{a}\ogadd{long exposure} coronagraphic observation\ogadd{s}.
+This adds structured noise, with both spatial and temporal correlations, which is the limiting noise source in a \ogadd{long exposure} coronagraphic observation\ogadd{s}.
 Both in space and on the ground, imperfections within the telescope and instrument optics cause speckles.
-For \ogadd{ground-based} telescopes \ogrmv{on the Earth's surface}, the aberrations \ogadd{also} include those imposed by \ogadd{atmospheric} turbulence.
-\ogrmv{in the Earth's atmosphere}
+For \ogadd{ground-based} telescopes, the aberrations \ogadd{also} include those imposed by \ogadd{atmospheric} turbulence.
 
 \jrmadd{[MOVED FROM DISCUSSION] For some time, long-lived ``quasi-static'' speckles have been recognized as the limiting noise source in contemporary high contrast imaging observations \citep[e.g.][]{2007ApJ...654..633H}.
-Given their origin within the instrument, and comparatively long lifetimes \citep{2012A&A...541A.136M}, such speckles should be supressable using wavefront sensing and control (WFS\&C) strategies.
+Given their origin within the instrument, and comparatively long lifetimes \citep{2012A&A...541A.136M}, such speckles should be suppressible using wavefront sensing and control (WFS\&C) strategies.
 Recent progress in applying ``dark hole'' contrast optimization with focal plane WFS \citep{2020A&A...638A.117P} and various focal-plane \citep[e.g.][]{2010A&A...509A..31G} and low-order coronagraphic WFS\&C strategies \citep[e.g.][]{2017PASP..129i5002S} show great promise in significantly reducing the impact of internal speckles.
@@
 
-In order to \ogrmv{analyze}\ogadd{estimate} the capabilities of ground-based high-contrast imaging, we must understand the spatial and temporal evolution of the residual atmospheric speckles behind an ExAO coronagraph.
+In order to \ogadd{estimate} the capabilities of ground-based high-contrast imaging, we must understand the spatial and temporal evolution of the residual atmospheric speckles behind an ExAO coronagraph.
 Speckle formation in the spatial-domain, as structures within the coronagraphic PSF, is well understood \citep{2001ApJ...558L..71B,2003ApJ...596..702P,2007ApJ...669..642S}.
@@
 In reality, the temporal behavior of atmospheric speckles falls between these regimes.
-These speckles vary on short time-scales \ogrmv{making standard PSF-subtraction essentially impossible}\ogadd{so they are not removed by standard PSF subtraction}, but exhibit non-negligible temporal correlations which cause them to average much more slowly than photon noise.
+These speckles vary on short time-scales \ogadd{so they are not removed by standard PSF subtraction}, but exhibit non-negligible temporal correlations which cause them to average much more slowly than photon noise.
 The statistical lifetime of atmospheric speckles, that is the time-scale which controls the averaging of speckle noise, is the quantity that sets the fundamental sensitivity limit of ground-based high-contrast imaging.
@@
 called there the ``speckle boiling time'', where $r_0$ is Fried's parameter and $\sigma^2_\mathcal{V}$ is the variance with respect to height of the layer wind velocities $\mathcal{V}$  above the telescope.
-At good astronomical sites \ogrmv{such as MKO and LCO}, values of $\tau_{boil}$ should range from $\sim$5 to $\sim$20 ms.
+At good astronomical sites values of $\tau_{boil}$ should range from $\sim$5 to $\sim$20 ms.
 \citet{1999PASP..111..587R} employed Equation (\ref{eqn:e-fold}) for the speckle lifetime and argued that atmospheric speckles dominate photon noise by factors of 100 or more in high contrast imaging.
@@
 
-\jrmadd{NOTE: need to add discussion of Poyneer and Mac 2006}
+\citet{2006OExpr..14.7499P} analyzed speckle noise correlations under closed-loop control, in similar fashion to what follows.
+Their study found that the wavefront sensor noise and input turbulence contributed differently to speckle lifetime.
+They also proposed that the control-loop would impact the lifetime, through the optimization of the control law gain, and that considering long-exposure sensitivity (i.e.
+through Equation \ref{eqn:varmean_def_intro}) may lead to different optimizations than by considering contrast alone.
 %Note for comparison that the Greenwood time-constant is
@@
 \jrmadd{where $I_{*}$ is the peak intensity of the star's PSF without a coronagraph, which is a function of time due to Strehl ratio variation, and $\Delta t$ is the exposure time.
-Here we will assume units of photons/sec/$(\lambda/D)^2$ for the intensity, where $\lambda$ is the observation wavelength and $D$ is the telescope diameter.
+Here we will assume units of photon/sec/$(\lambda/D)^2$ for the intensity, where $\lambda$ is the observation wavelength and $D$ is the telescope diameter.
 $\lambda/D$ is the fundamental spatial scale of diffraction limited imaging and coronagraphy, and we assume that photometry is conducted in pixels of size 1 $\lambda/D$.}
@@
 \end{equation}
-For argument's sake, assume a star having $I_* = 10^8$ photon\ogrmv{s}/sec/$(\lambda/D)^2$ in the peak (roughly an 8th mag star on a 10 m telescope in the near-IR), and $I_{as} = 1\times10^{-4}$ (see \citet{2018JATIS...4a9001M}).
+For argument's sake, assume a star having $I_* = 10^8$ photon/sec/$(\lambda/D)^2$ in the peak (roughly an 8th mag star on a 10 m telescope in the near-IR), and $I_{as} = 1\times10^{-4}$ (see \citet{2018JATIS...4a9001M}).
 Next we (for now) assume the \citet{2005SPIE.5903..170M} crossing-time speckle lifetime for a 10 m telescope with 10 m/s wind, giving $\tau_{as}$ = 0.3 s.
@@
 Similar to the Zernike polynomials, this basis has an intuitive progression through slope, quadratic, cubic, etc, modes (see Figure \ref{fig:onlegpol}).
-We later utilize these 1-deminsional modes to analyze post-processing as a time-domain filter.
+We later utilize these 1-dimensional modes to analyze post-processing as a time-domain filter.
 For that analysis it will be convenient to use a normalized form of Legendre polynomials defined as:}
@@
 \jrmadd{The coefficients are variable quantities.
-Depending on the time $T$ under consideration, the best fit of the polynomials will change as the speckle intensity $I(t)$ varies.
-By analyzing the temporal statistics of the $a_n$, we learn about the power in specific Legendre ``modes'' just as we do with the analgous Zernike modes in a wavefront.}
+The best fit of the polynomials will change with time as the speckle intensity varies.
+By analyzing the temporal statistics of the $a_n$, we learn about the power in specific Legendre ``modes'' over the time $T$ just as we do with the analogous Zernike modes in a wavefront over an aperture of diameter $D$ in \citet{1976JOSA...66..207N}.}
 
@@
 \end{equation}
-where $p=\pm 1$, $\vec{q}$ is the position vector in the pupil \ogrmv{plance}\ogadd{plane} and the vector spatial-frequency is 
+where $p=\pm 1$, $\vec{q}$ is the position vector in the pupil plane and the vector spatial-frequency is 
 \begin{equation}
@@
 
+\subsection{Spatial Correlations}
+
+In the following recipe for generating PSDs, we will produce time-series of the modal coefficients $h_{mn}^p$.
+In addition to the temporal phase relationship, there is a modest spatial correlation between the modes, even under good correction.
+This is illustrated in Figures 2 and 3 of \citet{2018JATIS...4a9001M}.
+This can be accounted for using the covariance matrix $\pmb{\Sigma}$ of the $h_{mn}^p$, which is calculated as described in that paper.
+Given an initially uncorrelated vector $\tilde{h}_{mn}^p$, and the decomposition $\pmb{\Sigma} = \pmb{A}\pmb{A}^T$, the spatially correlated vector of coefficients is found from
+\begin{equation}
+h_{mn}^p = \pmb{A} \tilde{h}_{mn}^p.
+\end{equation}
+
+We note that this adds significant computational cost to the following calculations, for relatively minor benefits.
+In practice we have found that it only modestly improves the match between these calculations and simulations at higher temporal frequencies, and has little to no noticeable impact on speckle lifetime predictions.
+We include it here for completeness, and consider it optional depending on the goals for employing the model.
+
 \subsection{A Recipe for Intensity PSDs}
@@
 \item Multiply the Fourier transform by the square-root of the open-loop PSD.  We now have $\mathcal{F}_t\left\{h_{mn}^p\right\}$ \citep[cf.][]{kasdin_falpha}.
-\item Apply the inverse Fourier transform to produce $h_{mn}^p(t)$, and then normalize to unit variance.
+\item Apply the inverse Fourier transform to produce $h_{mn}^p(t)$.  \jrmadd{If the correlation matrix will be applied in Step \ref{step:corr}, then} normalize to unit variance.
 \item Repeat for each mode.
 \end{enumerate}
-\item Correlate the modal time-series by multiplying the vector of coefficients at each time-step by the decomposition of the covariance matrix.
-See \citet{2018JATIS...4a9001M} for a treatment of the covariance matrix calculation for the Fourier basis.
-\item For each spatial frequency $mn$, replace $h_{mn}^-(t)$ with the phase shifted $h_{mn}^+(t)$ as described above.
+\item \jrmadd{[Optional]} Correlate the modal time-series by multiplying the vector of coefficients at each time-step by the decomposition of the covariance matrix. \label{step:corr}
+\item For each spatial frequency $mn$, replace $h_{mn}^-(t)$ with the phase shifted $h_{mn}^+(t)$.
 \item Generate a WFS measurement noise time-series for each mode, $\eta_{mn}^+$ and $\eta_{mn}^-$, and Fourier transform them.  These are white-noise PSDs, with no phase relationship.
@@
 The wavefront sensor (WFS) worked by simply projecting modes onto the phase screen.
-The basis set used for control was the modified Fourier basis (Equation (\ref{eqn:modified_fourier})) but orthornomalized with the stabilized Gramm-Schmidt procedure.
+The basis set used for control was the modified Fourier basis (Equation (\ref{eqn:modified_fourier})) but orthonormalized with the stabilized Gramm-Schmidt procedure.
 This was necessary for the loop to be stable under the LP controller.
@@
 To achieve the 1/2 frame, the WFS used the average turbulence between two time-steps.
-Turbulence was generated in 7 layers as fixed oversized phase screens, translated according to the layer windspeed.
-Turbulence parameters corresponded to the median LCO model used in \citet{2018JATIS...4a9001M}, where $r_0 = 0.16$ m, $L_0 = 25$ m, and layer averaged windspeed was 18.7 m/s.
+Turbulence was generated in 7 layers as fixed oversized phase screens, translated according to the layer wind speed.
+Turbulence parameters corresponded to the median LCO model used in \citet{2018JATIS...4a9001M}, where $r_0 = 0.16$ m, $L_0 = 25$ m, and layer averaged wind speed was 18.7 m/s.
 The control laws, both SI and LP with 21 coefficients, were optimized on the open-loop PSDs measured by running the simulation with the control loop off.
@@
 Figure \ref{fig:contrast_2Dcomp} shows post-coronagraph contrast maps, comparing the results of calculations using \citet{2018JATIS...4a9001M} to the simulation outputs.
-We show results for an $-\infty$ mag guidestar to show the effects of the input turbulence alone, and for an $8$th mag guidestar.
+We show results for an $-\infty$ mag guide star to show the effects of the input turbulence alone, and for an $8$th mag guide star.
 Figure \ref{fig:contrast_radprof} shows the median radial profile for each of the maps in Figure \ref{fig:contrast_2Dcomp}.
 The agreement is very good, though some small differences are visible in the LP (bottom row).
-\jrmadd{These differences, apparent in both the maps and the profile, are due to the LP conroller optimization leading to slightly different controller characteristics.
+\jrmadd{These differences, apparent in both the maps and the profile, are due to the LP controller optimization leading to slightly different controller characteristics.
 This is due to the difference in the modal basis actually under control in the simulations compared to the model.
@@
 Right: with photon noise on an 8th mag star assuming the sensitivity of an unmodulated pyramid WFS.
-Note that the colorbar is adjusted to best show detail.
+Note that the color bar is adjusted to best show detail.
 Agreement between the calculations and simulations is very good.\label{fig:contrast_2Dcomp}}
@@
 The bottom row shows the ETF for each controller.
-On the $-\infty$ mag guidestar, the OL PSD shows excellent agreement at low frequencies, with a slight divergence in slope beginning at a $\sim$100 Hz.
+On the $-\infty$ mag guide star, the OL PSD shows excellent agreement at low frequencies, with a slight divergence in slope beginning at a $\sim$100 Hz.
 The SI likewise shows nearly perfect agreement, with the exception of a lower overshoot peak in the simulation.
@@
 In intensity, the agreement between calculations and simulations is even better than in phase.
-There is a departure at high frequencies, where the calculations predict stronger peaks than occur in simulations.
-The agreements in overal power and the fraction of power conatined in low-frequencies, which are most important for determining speckle lifetimes, is very good.
+There is a departure at high frequencies, where the calculations predict stronger peaks than occur in the simulations.
+The agreements in overall power and the fraction of power contained in low-frequencies, which are most important for determining speckle lifetimes, are very good.
 
@@
 The agreement between the calculations and the simulations is very good at low frequencies, which determine the speckle lifetime.
-At higher frequencies the calculations predict stronger peaks than occur in the simulations  \label{fig:sppsdcomp}}
+At higher frequencies the calculations predict stronger peaks than occur in the simulations.
+\label{fig:sppsdcomp}}
 \end{figure}
@@
 The ``sim.
-PSD based'' lines show the results of using Equantion (\ref{eqn:psd_lifetime}) to predict the variance as a function of time using the simulated intensity PSDs.
+PSD based'' lines show the results of using Equation (\ref{eqn:psd_lifetime}) to predict the variance as a function of time using the simulated intensity PSDs.
 The comparison between the two simulation-based results shows that the analytic expression for speckle lifetime we derived is valid.
@@
 \includegraphics[width=3.54in]{lifetimes8mag.pdf}
-\caption{Comparison of control theory predictions for speckle lifetime to end-to-end simulation for a 6.5 m telescope.
+\caption{Comparison of model predictions for speckle lifetime to end-to-end simulation for a 6.5 m telescope.
 Left: without photon noise, considering the input turbulence only.
 Right: with photon noise on an 8th mag star assuming the sensitivity of an unmodulated pyramid WFS.
-Note that the colorbar is adjusted in each panel to best show detail.
+Note that the color bar is adjusted in each panel to best show detail.
 For the integrator the agreement is very good.
-For the predictor on the infinitely bright star, the agreement is good overall but the calculations \jrmadd{modestly} over-predict speckle lifetimes \jrmadd{in a wedge assocated with (but not parallel to)} the wind vectors.
-\label{fig:lifetime_2Dcomp}}
+For the predictor on the infinitely bright star, the agreement is good overall but the calculations \jrmadd{modestly} over-predict speckle lifetimes \jrmadd{in a wedge associated with (but not parallel to)} the wind vectors.
+\jrmadd{This results demonstrate that AO control does affect speckle lifetimes, and that the lifetime is significantly shorter under predictive control.
+The lifetime also depends on location in the control region.}  \label{fig:lifetime_2Dcomp}}
 \end{figure}
@@
 The calculations are somewhat conservative at smaller separations in the bright star case.
-\label{fig:lifetime_radprof}}
+\jrmadd{Speckle lifetime is shorter under predictive control compared to the simple integrator.} \label{fig:lifetime_radprof}}
 \end{figure}
@@
 
-\section{Trends in Speckle Lifetime}
+\section{Speckle Lifetime Comparisons}
 \label{sec:trends}
@@
 We also apply the Monte Carlo procedure to the OL case, \jrmadd{by treating the input PSDs as unmodified by control laws.
-We note that this is not strictly valid as the input phase is then outside the approximation regime made in the derivation of Equation (\ref{eqn:cross_terms}).
+We note that this is not strictly valid as the input phase is then outside the approximation regime assumed in the derivation of Equation (\ref{eqn:cross_terms}).
 However it provides a useful comparison of the change in the distribution of power in the PSD caused by closed-loop control action.
@@
 
-Next, we analyze the change in speckle lifetime with telescope diameter, windspeed, and seeing.
-Figure \ref{fig:scalings} shows the behavior of speckle lifetime at a constant absolute spatial frequency.
+%Next, we analyze the change in speckle lifetime with telescope diameter, windspeed, and seeing.
+
+%Figure \ref{fig:scalings} shows the behavior of speckle lifetime at a constant absolute spatial frequency.
 This means that for a given spatial frequency, the shape of the input PSD is the same regardless of $D$, but with an integrated variance different by a factor of $1/D^2$.
@@
 
-On the 6.5 m aperture using the SI controller, windspeed has a strong effect, but on the larger ELT scale apertures wind has no significant effect.
+%On the 6.5 m aperture using the SI controller, windspeed has a strong effect, but on the larger ELT scale apertures wind has no significant effect.
 Given the range of plausible mean windspeeds, \jrmadd{the ELTs appear to be in a different regime}.
@@
 
-For the LP controller, increasing windspeed tends to increase speckle lifetime.
+%For the LP controller, increasing windspeed tends to increase speckle lifetime.
 This is a result of increasing power concentration in the PSD peaks at $\vec{\mathcal{V}} \cdot \vec{k}$ Hz, which causes the LP controller to optimize there at the expense of worse low-frequency rejection, thus causing the lifetime to increase.
 
-Diameter has a weak effect, with lifetimes dropping as $D$ increases.
+%Diameter has a weak effect, with lifetimes dropping as $D$ increases.
 When combined as ``crossing time'' $D/V$ there is some evidence for the $\tau \propto D/\mathcal{V}$ relation predicted by \citet{2005SPIE.5903..170M} on smaller apertures, but this does not hold for larger telescopes and does not apply to predictive control.
 
-Finally, our calculations predict that improving seeing ($r_0$ growing larger) lowers speckle lifetime.
+%Finally, our calculations predict that improving seeing ($r_0$ growing larger) lowers speckle lifetime.
 This is because increasing $r_0$ lowers the overall power at each spatial frequency, which in turn causes the control system to perform better at lower temporal frequencies.
@@
 \begin{figure}
-\includegraphics[width=5.5in]{scalings.pdf}
-\caption{Changes in speckle lifetime with turbulence parameters.
+    \includegraphics[width=3.5in]{scalings_rel_5.pdf}
+    \includegraphics[width=3.5in]{scalings_rel_8.pdf}
+    \caption{\jrmadd{Comparison of speckle lifetimes over a range of parameters.
+We compare speckle lifetimes with a 5th mag guide star (left) and 8th mag guide star (right) on telescopes of diameter $D$ up to 39 m, with wind speeds $\mathcal{V}$ up to 35 m/s, and seeing $r_0$ up to 0.25 m.
+Only the speckle lifetime for the $(m,n)$ = $(10,14)$ mode is analyzed here, for a 2 kHz AO system.
+For the simple integrator (SI, blue), speckle lifetime generally decreases on fainter stars due to the larger contribution of photon noise compared to the turbulence power in the control system.
+Additionally the SI shows some dependence on the crossing time up to $D/\mathcal{V}\sim2$.
+The linear predictor (LP, red) controller is somewhat less dependent on these parameters.
+These comparisons demonstrate that AO control has a large impact on speckle lifetime, and that there are no simple thumb rules which can be used to predict speckle lifetime.
+} \label{fig:scalings}}
+\end{figure}
+
+%\begin{figure}
+%    \includegraphics[width=5.5in]{scalings.pdf}
+%\caption{Changes in speckle lifetime with turbulence parameters.
 Here absolute spatial frequency is held constant \jrmadd{for all aperture sizes}.
@@
 Improvement in $r_0$ (larger value, better seeing) shows a general reduction in speckle lifetime in all cases.
-\label{fig:scalings}
-}
-\end{figure}
-
-We next held relative spatial frequency constant with $D$, such that the resultant speckle occurs at the same relative projected separation in $\lambda/D$ units.
+%\label{fig:scalings}
+%}
+%\end{figure}
+
+\jrmadd{Next, we analyze the change in speckle lifetime with guide star brightness, telescope diameter, wind speed, and seeing.
+Note that power per Fourier mode scales as $1/D^2$.
+We held relative spatial frequency constant with $D$, such that the resultant speckle occurs at the same relative projected separation in $\lambda/D$ units.
 This means that the input temporal PSD has a different shape as well as different power.
-While more difficult to interpret in terms of atmospheric parameters, this compares the performance of coronagraphic AO systems given the improvement in spatial resolution with $D$.
-
-Here, higher windspeed generally shortens speckle lifetime.
-Under SI control, speckle lifetime increases with $D$, but shows little change under LP control.
-There is a scaling up to $D/\mathcal{V} \sim 2$ for the SI, but this breaks down at longer crossing times (slower winds).
-For the LP controller speckle lifetime is essentially flat vs.
-$D/\mathcal{V}$.
-
-As in the absolute spatial frequency case, improving seeing lowers speckle lifetime in this relative spatial frequency comparison.
-
-\begin{figure}
-\includegraphics[width=5.5in]{scalings_rel.pdf}
-\caption{Changes in speckle lifetime with turbulence parameters.
+This compares the performance of coronagraphic AO systems given the improvement in spatial resolution with $D$.}
+
+\jrmadd{Figure \ref{fig:scalings} compares speckle lifetimes for a 5th magnitude guide star to those for an 8th mag guide star.
+The lifetime is generally shorter on the 8th mag star due to the increased influence of photon noise relative to the turbulence PSD, with the overall trends being similar.
+Lifetimes are shorter under LP control compared to SI control for both guide stars for all parameters.}
+
+\jrmadd{The SI controller, in general, shows a decrease in lifetime with wind speed $\mathcal{V}$ and an increase with diameter $D$ for both guide stars.
+When analyzed as the crossing time $D/\mathcal{V}$, the SI controller scales roughly proportional to the crossing-time up to $D/\mathcal{V}$$\sim$$2$.
+For longer crossing times the trend becomes flat.
+This implies that the control system becomes dominated by photon noise rather than servo-lag and so the optimization process no longer produces changes in the output dynamics.}
+
+\jrmadd{The LP controller shows much weaker dependence on the same parameters.
+Except at the lowest wind speeds, the speckle lifetime is $\sim$$20$ ms on the 5th magnitude guide star.
+On the 8th magnitude star photon noise makes predictive control less effective, and there is a proportional scaling up to $D/\mathcal{V}$$\sim$$3$.}
+
+\jrmadd{In most cases, improving seeing (larger $r_0$), reduces speckle lifetime.
+As the turbulence power is reduced, the control system is better able to reject low temporal frequencies, resulting in lower speckle lifetimes.
+Interestingly, on the 8th mag guide star and for the largest apertures the LP controller shows the opposite trend.
+The power in each Fourier mode is proportional to $1/D^2$.
+Projected actuator spacing was held constant, so the photon noise per mode is constant in these comparisons.
+Furthermore, given the highest jet stream driven wind speeds we can expect, the GSMTs will normally operate in the regime of $D/\mathcal{V} \gtrsim 1$.
+Thus the GSMTs are operating in a different regime compared to the 6.5 m, and the LP controller produces different dynamics  when photon noise becomes important on the 8th mag star.}
+
+%\begin{figure}
+%\includegraphics[width=5.5in]{scalings_rel.pdf}
+%\caption{Changes in speckle lifetime with turbulence parameters.
 Here relative spatial frequency is held constant.
@@
 Predictive control shows little dependence on $D$ or $V$.
-\label{fig:scalings_rel}
-}
-\end{figure}
+%\label{fig:scalings_rel}
+%}
+%\end{figure}
 %\clearpage
@@
 \end{equation}
-The outcome will be fewer long-term correlations as the low order modes are sucessively subtracted, which will in turn lower the speckle lifetime.
+The outcome will be fewer long-term correlations as the low order modes are successively subtracted, which will in turn lower the speckle lifetime.
 We can describe this outcome statistically as a high-pass filter (HPF).
@@
 \caption{The Legendre high pass filter rejection functions for 1, 10, and 100 modes.
-Fitting and subtracting legendre polynomials reduces low-frequency power, which lowers speckle lifetimes.\label{fig:leghpf}}
+Fitting and subtracting Legendre polynomials reduces low-frequency power, which lowers speckle lifetimes.\label{fig:leghpf}}
 \end{figure}
@@
 \caption{The relative sensitivity, in terms of exposure time, after subtracting the first N Legendre polynomials with uncertainty in the fit $\gamma$.
-The horizontal line at 1 corresponds to the photon-noise limited exposure time.
+The horizontal line at 1 corresponds to the photon-noise limited exposure time on an 8th mag star.
 Residual-atmospheric speckles increase the exposure time by a factor 55 compared to the photon-noise limit.
@@
 \subsection{Caveats}
-The results shown here are subject to several caveats.
+\jrmadd{The results shown here are subject to several caveats.
 Our analysis is monochromatic.
@@
 These speckles can, however, be analyzed with the same tools we developed here.
-We save this for future work.
-
-A key assumption underlying the mathematics of the PSD analysis is that the speckle intensity is wide-sense stationary (WSS), which means that the mean is constant and that the autocorrelation does not depend on time.
+We save this for future work.}
+
+\jrmadd{A key assumption underlying the mathematics of the PSD analysis is that the speckle intensity is wide-sense stationary (WSS), which means that the mean is constant and that the autocorrelation does not depend on time.
 In this context, WSS essentially implies that the atmosphere parameters, namely $r_0$ and the layer $\vec{\mathcal{V}}$, are constant.
 This is of course not strictly true, but can be expected to be approximately true over short periods of time -- at least long enough to estimate a PSD.
-We also expect AO control to dampen changes, keeping the output speckle formation process within a more or less well defined range.
-What this ultimately means is that we must consider the speckle lifetime as itself a time-variable parameter, but there will be a representative value over an observation that characterizes long exposure sensitivity.
+We also expect AO control to damp changes, keeping the output speckle formation process within a more or less well defined range.
+What this ultimately means is that we must consider the speckle lifetime as itself a time-variable parameter, but expect there to be a representative value over an observation that characterizes long exposure sensitivity.}
 
@@
 
-We have shown that the speckle lifetime in well-corrected post-coronagraph focal planes is an order of magnitude shorter than claimed in previous studies.
-This result was obtained by developing a semi-analytic modeling framework, which allows us to calculate the intensity PSD at any point in the focal plane.
+\jrmadd{We have developed a semi-analytic model of a coronagraphic AO system that allows us to estimate long-exposure contrast, the intensity dynamics, and statistical lifetimes.
+Compared to an idealized end-to-end simulation, the results of the model compare very well.
 This powerful technique allows us to analyze the impact of control law choice, as well as to compare the influence of various telescope and atmosphere parameters on the speckle lifetime.
-
-\subsection{Trends in Speckle Lifetime}
-
-Of note, we find that the simple $D/\mathcal{V}$ ``crossing-time'' concept holds only for integrator control on contemporary $D\lesssim10$ m telescopes where $D/\mathcal{V} \lesssim 1$ second.
-With predictive controllers and on the larger ELT-class apertures under development the crossing-time is not useful for predicting speckle lifetime.
-We  note that the parameter study in Section (\ref{sec:trends}) would be prohibitively expensive in terms of time and computing resources if conducted using end-to-end simulations.
-Such studies are enabled by the semi-analytic framework we have developed, beginning in \citep{2018JATIS...4a9001M} and extended to intensity dynamics in this work.
+Key benefits of such modeling are that the true long-exposure performance of a system can be determined without the expense of long simulations.
+For instance, the parameter study in Section \ref{sec:trends} would have been prohibitively expensive if conducted with simulations.
+As design efforts continue for the coming GSMTs, such models will be useful for similar parameter studies to optimize high contrast imaging instruments.}
+
+\subsection{The Impact of AO}
+
+\jrmadd{We analyzed various parameters with potential impacts on speckle lifetime.
+Overall, the simplest summary of this analysis is that there is no single thumb rule type relationship that applies to all parameters.
+This modestly contradicts the conclusions of \citet{2005SPIE.5903..170M}.
+There, the AO system was modeled as a spatial filter, with the result that while the speckle amplitude was reduced, the evolution of correlations depended only on the crossing-time.
+As a result, neither the control action nor the overall power (seeing) should impact speckle lifetime.
+Only the wind would change speckle lifetime for a given diameter.}
+
+\jrmadd{Here we have added the time dimension, fully considering the action of AO control as a temporal filter.
+Furthermore, we have modeled the impact of high order predictive control.
+With this more complete view, we find that the control action does reduce speckle lifetime.
+There is an influence of the crossing-time as predicted by \citet{2005SPIE.5903..170M} to some extent, but in addition to the reduction in amplitude, the time-domain influence of the control system is to increase the rate at which one speckle turns into the next.
+That is, the control reduces the correlation length.
+This is consistent with the findings of \citet{2006OExpr..14.7499P}, which predicted that control laws could be optimized based on speckle lifetime to maximize sensitivity.}
+
+\jrmadd{Predictive control, already expected to improve contrast \citep{2007JOSAA..24.2645P,2017JOSAA..34.1877C,2018JATIS...4a9001M, 2021arXiv210307566H}, will further improve sensitivity by decreasing the speckle averaging lifetime.
+On bright stars, predictive control nearly eliminates dependence on the crossing-time as it acts to reduce servo-lag errors.}
+
+\jrmadd{As ExAO coronagraphs approach the residual atmosphere limit, it will be increasingly important to understand the impact of correlated noise from atmospheric speckles on sensitivity.
+Our results demonstrate that this is a complicated task, that must consider all relevant parameters of the system under study.}
 
 \subsection{Post-Processing}
-The temporal decorrelation of speckle noise is one of the main benefits of PSF-subtraction post-processing \citep{2006ApJ...641..556M}, even though algorithms have typically been focused on the spatial domain.
+The temporal decorrelation of speckle noise is one of the main benefits of PSF-subtraction post-processing \citep{2006ApJ...641..556M,2008ApJ...673..647M}, even if algorithms have typically been focused on the spatial domain.
 Recently, studies have begun to consider the time-domain in post-processing explicitly \citep{2021A&A...646A..24S}.
-Knowledge of the intensity PSD allows for such algorithms to be analyzed eficiently.
+Knowledge of the intensity PSD allows for such algorithms to be analyzed efficiently.
 We demonstrate this by considering the prospect of telemetry-based reconstruction of the speckle time-series.
@@
 While we did not present a specific algorithm, we used the temporal intensity PSDs to predict the sensitivity gain possible.
-If the first few low-order temporal modes, here described in terms of the Legendre polynomials, can be determined and subtracted from the intensity in the focal plane, orders of magnitude improvements in exposure time are possible.
+If the first few low-order temporal modes, here described in terms of the Legendre polynomials, can be determined and subtracted from the intensity in the focal plane, \jrmadd{large} improvements in exposure time are possible.
 The fact that only a few low-order temporal modes need to be reconstructed has the important consequence that any such post-processing algorithm does not need to accurately determine the intensity at every time interval, and does not need to be perfect.
 Accuracy in the mean, slope, and other low-order terms of a few percent is all that is needed to make these gains.
-
-\subsection{Progress Towards Fundamental Limits}
 
@@
 As above, we have a one-dimensional, one-sided ($f \ge 0$),  PSD $\mathcal{T}(f)$, where $f$ is normalized on $0 \leq f \leq f_s = 1/\Delta t$ according to Equation (\ref{eqn:process_var}).
-We obtain a time-series $I(t)$ from the process goverened by $\mathcal{T}(f)$, which we expand in terms of normalized Legendre polynomials defined in Equation (\ref{eqn:basis}).
+We obtain a time-series $I(t)$ from the process governed by $\mathcal{T}(f)$, which we expand in terms of normalized Legendre polynomials defined in Equation (\ref{eqn:basis}).
 The orthogonality condition is given in Equation (\ref{eqn:orthonormal}).

### 2021-03-28 specklelives.tex
Updates from Overleaf

@@
 \newcommand{\ogrmv}[1]{\textcolor{orange}{\sout{#1}}}
+
+% custom commends for Mike's comments
+\definecolor{avocado}{rgb}{0.34,0.51,0.01}
+\newcommand{\mpfadd}[1]{\textcolor{avocado}{#1}}
+\newcommand{\mpfrmv}[1]{\textcolor{avocado}{\sout{#1}}}
 
@@
 
-\author{Michael P. Fitzgerald}
+\author[0000-0002-0176-8973]{Michael P. Fitzgerald}
 \affiliation{Department of Physics \& Astronomy, University of California, Los Angeles, CA 90095, USA}
@@
 \begin{abstract}
-High contrast imaging observations are fundamentally limited by the spatially and temporally correlated noise source called speckles.
+High-contrast imaging observations are fundamentally limited by the spatially and temporally correlated noise source called speckles.
 Suppression of speckle \jrmadd{noise} is the key goal of wavefront control and adaptive optics (AO), coronagraphy, and a host of post-processing techniques.
 The temporal correlation of speckles reduces sensitivity as they average much more slowly than photon noise.
-Speckles average at a rate set by the statistical speckle lifetime, and speckle-limited integration time in long exposures is directly proportional to this lifetime.
-As progress continues in post-AO coronagraphic wavefront control, residual atmospheric speckles will become the limiting noise source in high contrast imaging, so a complete understanding of their statistical behavior is crucial to optimizing high contrast imaging instruments.
+\mpfadd{In speckle-limited integrations, the rate at which speckle intensity fluctuations approach the mean intensity is a function of the second-order temporal statistics.
+Using these statistics, we can define a speckle lifetime which describes how quickly speckles average in long exposures.} \mpfrmv{Speckles average at a rate set by the statistical speckle lifetime, and speckle-limited integration time in long exposures is directly proportional to this lifetime.}  As progress continues in post-AO coronagraphic wavefront control, residual atmospheric speckles will become the limiting noise source in high-contrast imaging, so a complete understanding of their statistical behavior is crucial to optimizing high-contrast imaging instruments.
 Here we present a novel power spectral density (PSD) method for calculating the lifetime, and develop a semi-analytic method for predicting intensity PSDs behind a coronagraph.
-Considering a frozen-flow turbulence model, we analyze the residual atmosphere speckle lifetimes in a MagAO-X-like AO system as well as 25-39 m giant segmented mirror telescope (GSMT) scale systems.
+Considering a frozen-flow turbulence model, we analyze the residual atmosphere speckle lifetimes in a MagAO-X-like AO system as well as 25--39 m giant segmented mirror telescope (GSMT) scale systems.
 We find that standard AO control shortens atmospheric speckle lifetime from $\sim$130 ms to $\sim$50 ms, and predictive control will further shorten the lifetime to $\sim$20 ms on 6.5 m MagAO-X.
@@
 Due to control system dynamics there are no simple scaling laws which apply across a wide range of system characteristics.
-Finally, we use these results to describe how post-processing will enable ground-based telescopes to achieve the photon-noise limit in high contrast imaging.
+Finally, we use these results to describe how post-processing will enable ground-based telescopes to achieve the photon-noise limit in high-contrast imaging.
 \end{abstract}
@@
 Characterization of extrasolar planets with resolved imaging is limited by the halo of starlight scattered onto the image plane by optical aberrations.
-In addition to the photon noise caused by this halo, the scattered light is coherent and produces ``speckles'', which can be thought of as copies of the instrument point spread function (PSF) \citep{1995PASP..107..386M}.
+In addition to the photon noise caused by this halo, the scattered light is coherent and produces ``speckles,'' which can be thought of as copies of the instrument point spread function (PSF) \citep{1995PASP..107..386M}.
 This adds structured noise, with both spatial and temporal correlations, which is the limiting noise source in a \ogadd{long exposure} coronagraphic observation\ogadd{s}.
@@
 
-The index of refraction of air depends on temperature, and turbulence causes the air temperature to stochastically vary across the telescope aperture.
+\mpfadd{Speckles also arise from atmospheric turbulence.}  The index of refraction of air depends on temperature, and turbulence causes the air temperature to stochastically vary across the telescope aperture.
 As the stellar wavefront traverses these regions of different index of refraction, parts of the wavefront travel different optical path lengths.
-Due mainly to wind, the atmospheric temperature structure is constantly changing.
+Due mainly to wind\mpfadd{-driven bulk flow across the telescope aperture}, the atmospheric temperature structure is constantly changing.
 The technique of adaptive optics (AO) is used to correct the resulting time-variable optical aberrations in real time.
-Even with so-called ``extreme'' AO systems \citep[ExAO, ][]{2018ARAA..56..315G} this correction is not perfect.
+Even with so-called ``extreme'' AO systems \citep[ExAO, ][]{2018ARAA..56..315G} this correction is \mpfrmv{not} \mpfadd{im}perfect.
 The AO-corrected wavefront, though significantly improved, still causes a halo of speckles significantly brighter than the instrumental speckles in a modern high performance coronagraph system using coronagraphic wavefront control.
@@
 In order to \ogadd{estimate} the capabilities of ground-based high-contrast imaging, we must understand the spatial and temporal evolution of the residual atmospheric speckles behind an ExAO coronagraph.
-Speckle formation in the spatial-domain, as structures within the coronagraphic PSF, is well understood \citep{2001ApJ...558L..71B,2003ApJ...596..702P,2007ApJ...669..642S}.
-In the time-domain, we can imagine two ideal regimes for these structures.
+Speckle formation in the spatial domain, as structures within the coronagraphic PSF, is well understood \citep{2001ApJ...558L..71B,2003ApJ...596..702P,2007ApJ...669..642S}.
+In the time domain, we can imagine two ideal regimes for these structures.
 In the first regime, the speckles are perfectly static.
@@
 While this would still leave the photon noise from the scattered light, we could perfectly remove the spatially correlated noise.
-In the second imaginary regime, the speckles are a perfect temporally white noise source: every image taken has a different speckle structure.
-Similar to photon noise, it would now be a simple matter of taking enough images to average the speckles to a smooth halo.
+In the second \mpfrmv{imaginary}\mpfadd{idealized} regime, the speckles are a perfect temporally white noise source: every image taken has a different speckle structure.
+Similar to photon noise, \mpfrmv{it}\mpfadd{high-contrast imaging} would now be a simple matter of taking enough images to average the speckles to a smooth halo.
 
@@
 There have been several definitions of the atmospheric speckle lifetime presented in the literature.
-Early treatments of speckle dynamics were motivated by speckle interferometry, which relies on exposures short enough to ``freeze the atmosphere''.
-The optimum exposure time is therefore dependent on the coherence time of the intensity in the focal plane.
-A common definition for this coherence time was based on a $1/e$ reduction in the autocorrelation of the intensity, which was experimentally found to be on the order of a few ms \citep{1978ApOpt..17.3779S,1990JMOp...37.1247D}.
+Early treatments of speckle dynamics were motivated by speckle interferometry, which relies on exposures short enough to ``freeze the atmosphere.''  The optimum exposure time is therefore dependent on the \mpfrmv{coherence}\mpfadd{correlation} time of the intensity in the focal plane.
+A common definition for this \mpfrmv{coherence}\mpfadd{correlation} time was based on a $1/e$ reduction in the autocorrelation of the intensity, which was experimentally found to be on the order of a few ms \citep{1978ApOpt..17.3779S,1990JMOp...37.1247D}.
 \citet{1982JOpt...13..263R} derived the following expression for the autocorrelation e-folding time in a multi-layer Kolmogorov atmosphere:
@@
 \end{equation}
-called there the ``speckle boiling time'', where $r_0$ is Fried's parameter and $\sigma^2_\mathcal{V}$ is the variance with respect to height of the layer wind velocities $\mathcal{V}$  above the telescope.
-At good astronomical sites values of $\tau_{boil}$ should range from $\sim$5 to $\sim$20 ms.
-\citet{1999PASP..111..587R} employed Equation (\ref{eqn:e-fold}) for the speckle lifetime and argued that atmospheric speckles dominate photon noise by factors of 100 or more in high contrast imaging.
+called there the ``speckle boiling time,'' where $r_0$ is Fried's parameter and $\sigma^2_\mathcal{V}$ is the variance with respect to height of the layer wind velocities $\mathcal{V}$  above the telescope.
+At good astronomical sites, values of $\tau_\text{boil}$ should range from $\sim$5 to $\sim$20 ms.
+\citet{1999PASP..111..587R} employed Equation (\ref{eqn:e-fold}) for the speckle lifetime and argued that atmospheric speckles dominate photon noise by factors of 100 or more in high-contrast imaging.
 
@@
 Defined this way,} $\tau$ is given by the integral of the autocorrelation \citep{2006ApJ...637..541F}, and in this expression determines the number of uncorrelated realizations of the speckle pattern included in the mean.
-Note that this assumes $t >> \tau$.
+Note that this assumes $t \gg \tau$.
 \citet{1986JOSAA...3.1001A} employed this definition, and argued for longer speckle lifetimes, replacing $0.36$ by $~$$1.14$ in Equation (\ref{eqn:e-fold}) based on arguments similar to \citet{1982JOpt...13..263R}.
-Using AO corrected images, \citet{2006ApJ...637..541F} empirically found significantly longer values of $\tau$, giving an overall median of 175 ms at Lick observatory.
+Using AO-corrected images, \citet{2006ApJ...637..541F} empirically found significantly longer values of $\tau$, giving an overall median of 175 ms at Lick observatory.
 Of note, the results of \citet{2006ApJ...637..541F} hinted that inside the region of the image plane controlled by the AO system, the speckle lifetime was significantly shorter than outside.
@@
 They also proposed that the control-loop would impact the lifetime, through the optimization of the control law gain, and that considering long-exposure sensitivity (i.e.
-through Equation \ref{eqn:varmean_def_intro}) may lead to different optimizations than by considering contrast alone.
+through Equation \ref{eqn:varmean_def_intro}) may lead to different optimizations than by considering raw contrast alone.
 %Note for comparison that the Greenwood time-constant is
@@
 More recent experimental studies have used on-sky data with modern ExAO systems.
-Using focal plane data from SPHERE, \citet{2016SPIE.9909E..4ZM} found evidence for speckle decorrelation on several time-scales, but were unable to identify an atmospheric component given the various instrument-induced time-scales and the relatively low cadence used.
-In a study using SCExAO at much higher frame-rates, \citet{2018PASP..130j4502G} analyzed the evolution of the intensity using the variance of difference images (essentially a form of autocorrelation).
-Here short time-scale (a few ms) speckle evolution was attributed to the atmosphere, and the authors argued that the AO system modulated the speckles on longer time-scales.
+Using focal plane data from SPHERE, \citet{2016SPIE.9909E..4ZM} found evidence for speckle decorrelation on several timescales, but were unable to identify an atmospheric component given the various instrument-induced timescales and the relatively low cadence used.
+In a study using SCExAO at much higher frame rates, \citet{2018PASP..130j4502G} analyzed the evolution of the intensity using the variance of difference images (\mpfrmv{essentially }a form of autocorrelation).
+Here short timescale (a few ms) speckle evolution was attributed to the atmosphere, and the authors argued that the AO system modulated the speckles on longer timescales.
 \citet{2017JATIS...3b5001S} present a study of speckle lifetimes at the LBT, where they found short-lived speckles attributed to the atmosphere and show some evidence for a difference in lifetime inside and outside the AO control radius.
@@
 \jrmadd{Our goal in this study is to develop a self-consistent description of the time-domain behavior of speckle noise in AO-fed coronagraphic images from as close to first principles as possible.
-Our motivation is to understand the true limits of ground-based high contrast imaging and so we focus on the statistical lifetime of atmospheric speckles in coronagraphic focal planes.
-This requires a traversal of speckle statistics, signal processing, wavefront propagation, and control theory, and so is somewhat involved.
+Our motivation is to understand the true limits of ground-based high-contrast imaging and so we focus on the statistical lifetimes of atmospheric speckles in coronagraphic focal planes.
+This requires a traversal of speckle statistics, signal processing, wavefront propagation, and control theory\mpfrmv{, and so is somewhat involved}.
 To accomplish our goals we will work through the following tasks, organized by section:
@@
 \item Section \ref{sec:trends}: Use the analytic framework to analyze speckle lifetime over a range of telescope diameters and atmospheric parameters.
-\item Section \ref{sec:reconstruction}: We propose a method to analyze post-processing in the time-domain and discuss its implications for the limits of ground-based high-contrast imaging sensitivity.
-\item Section \ref{sec:discussion}: We then discuss various implications of these results.
+\item Section \ref{sec:reconstruction}: \jrmrmv{We propose} \jrmadd{Develop} a method to analyze post-processing in the time-domain and discuss its implications for the limits of ground-based high-contrast imaging sensitivity.
+\item Section \ref{sec:discussion}: \jrmrmv{We then d} \jrmadd{D}iscuss various implications of these results.
 \end{itemize}
@@
 \label{sec:noise}
-\jrmadd{The key metric for sensitivity in a high-contrast imaging observation is the signal-to-noise ratio, S/N.  The signal we collect from a planet, which has planet to star contrast of $C_p$, is }
-\begin{equation}
-\mbox{Planet Signal} = I_{*}(t) C_p \Delta t
+\jrmadd{\mpfrmv{The}\mpfadd{A} key metric for sensitivity in a high-contrast imaging observation is the signal-to-noise ratio, S/N.
+The \jrmrmv{signal we collect from} intensity of a planet, which has planet to star contrast of $C_p$, is }
+\begin{equation}
+I_p(t) = I_{*}(t) C_p
 \label{eqn:signal}
 \end{equation}
-\jrmadd{where $I_{*}$ is the peak intensity of the star's PSF without a coronagraph, which is a function of time due to Strehl ratio variation, and $\Delta t$ is the exposure time.
-Here we will assume units of photon/sec/$(\lambda/D)^2$ for the intensity, where $\lambda$ is the observation wavelength and $D$ is the telescope diameter.
+\jrmadd{where $I_{*}$ is the peak intensity of the star's PSF without a coronagraph, which is a function of time due to Strehl ratio variation.\jrmrmv{, and $\Delta t$ is the exposure time.}  Here we will assume units of photon/sec/$(\lambda/D)^2$ for the intensity, where $\lambda$ is the observation wavelength and $D$ is the telescope diameter.
 $\lambda/D$ is the fundamental spatial scale of diffraction limited imaging and coronagraphy, and we assume that photometry is conducted in pixels of size 1 $\lambda/D$.}
@@
 \begin{equation}
-I_h(\vec{r}_p,t) = I_{*}(t)\left[I_c(\vec{r}_p,t) + I_{as}(\vec{r}_p,t) + I_{qs}(\vec{r}_p,t)\right]
-\end{equation}
-$I_c$ is the fractional contribution from residual diffraction from the coronagraph, as well as purely static aberrations.
-$I_{as}$ is the fractional contribution from residual atmospheric speckles, and $I_{qs}$ is the fractional contribution from instrumental speckles, which are usually called quasi-static speckles.
+I_h(\vec{r}_p,t) = I_{*}(t)\left[\mathcal{I}_c(\vec{r}_p,t) + \mathcal{I}_{as}(\vec{r}_p,t) + \mathcal{I}_{qs}(\vec{r}_p,t)\right]
+\end{equation}
+$\mathcal{I}_c$ is the fractional contribution from residual diffraction from the coronagraph, as well as purely static aberrations.
+$\mathcal{I}_{as}$ is the fractional contribution from residual atmospheric speckles, and $\mathcal{I}_{qs}$ is the fractional contribution from instrumental speckles, which are usually called quasi-static speckles.
 From this we can state the instantaneous raw halo contrast
 \begin{equation}
-C_h(\vec{r}_p,t) = \frac{I_h(\vec{r}_p,t)}{I_*(t)} = I_c(\vec{r}_p,t) + I_{as}(\vec{r}_p,t) + I_{qs}(\vec{r}_p,t).
+C_h(\vec{r}_p,t) = \frac{I_h(\vec{r}_p,t)}{I_*(t)} = \mathcal{I}_c(\vec{r}_p,t) + \mathcal{I}_{as}(\vec{r}_p,t) + \mathcal{I}_{qs}(\vec{r}_p,t).
 \end{equation}
@@
 \begin{equation}
-\sigma^2_{h}  = I_* \Delta t [\underbrace{I_c + I_{as} + I_{qs}}_{\mbox{photon noise}} + I_* [ \underbrace{\tau_{as}\left(I_{as}^2 + 2[I_cI_{as} + I_{as}I_{qs}]  \right)}_{\mbox{atm.
-speckles}} +  \underbrace{\tau_{qs}\left( I_{qs}^2 + 2 I_c I_{qs}\right)}_{\mbox{QS speckles}}]]
+\sigma^2_{h}  = I_* \Delta t [\underbrace{\mathcal{I}_c + \mathcal{I}_{as} + \mathcal{I}_{qs}}_{\mbox{photon noise}} + I_* [ \underbrace{\tau_{as}\left(\mathcal{I}_{as}^2 + 2[\mathcal{I}_c\mathcal{I}_{as} + \mathcal{I}_{as}\mathcal{I}_{qs}]  \right)}_{\mbox{atm.
+speckles}} +  \underbrace{\tau_{qs}\left( \mathcal{I}_{qs}^2 + 2 \mathcal{I}_c \mathcal{I}_{qs}\right)}_{\mbox{QS speckles}}]]
 \label{eqn:variance}
@@
 
-Our present focus is on the atmospheric residual $I_{as}$.
-From here on we will assume a perfect coronagraph and instrument, such that $I_c = 0$ and $I_{qs}$ = 0.
+Our present focus is on the atmospheric residual $\mathcal{I}_{as}$.
+From here on we will assume a perfect coronagraph and instrument, such that $\mathcal{I}_c = 0$ and $\mathcal{I}_{qs}$ = 0.
 These (optimistic) conditions eliminate the speckle pinning \citep{2001ApJ...558L..71B} terms in Equation (\ref{eqn:variance}).
@@
 \begin{equation}
-\sigma^2_{h}  = I_* \Delta t \left(I_{as} + I_* \tau_{as} I_{as}^2 \right)
+\sigma^2_{h}  = I_* \Delta t \left(\mathcal{I}_{as} + I_* \tau_{as} \mathcal{I}_{as}^2 \right)
 \label{eqn:variance_as}
@@
 \begin{equation}
-\mbox{S/N} = \frac{C_p \sqrt{I_* \Delta t}}{ \sqrt{I_{as} (1 + I_*I_{as} \tau_{as})}} 
-\end{equation}
-For argument's sake, assume a star having $I_* = 10^8$ photon/sec/$(\lambda/D)^2$ in the peak (roughly an 8th mag star on a 10 m telescope in the near-IR), and $I_{as} = 1\times10^{-4}$ (see \citet{2018JATIS...4a9001M}).
-Next we (for now) assume the \citet{2005SPIE.5903..170M} crossing-time speckle lifetime for a 10 m telescope with 10 m/s wind, giving $\tau_{as}$ = 0.3 s.
-We now have $I_*I_{as} \tau_{as} = 3000$.
-We see that under such conditions  $I_*I_{as} \tau_{as} >> 1$, which would hold even for significantly smaller $\tau_{as}$.
+\mbox{S/N} = \frac{C_p \sqrt{I_* \Delta t}}{ \sqrt{\mathcal{I}_{as} (1 + I_*\mathcal{I}_{as} \tau_{as})}} 
+\end{equation}
+For argument's sake, assume a star having $I_* = 10^8$ photon/sec/$(\lambda/D)^2$ in the peak (roughly an 8th mag star on a 10 m telescope in the near-IR), and $\mathcal{I}_{as} = 1\times10^{-4}$ (see \citealt{2018JATIS...4a9001M}).
+Next we (for now) assume the \citet{2005SPIE.5903..170M} crossing-time speckle lifetime for a 10-m telescope with 10-m/s wind, giving $\tau_{as}$ = 0.3 s.
+We now have $I_*\mathcal{I}_{as} \tau_{as} = 3000$.
+We see that under such conditions  $I_*\mathcal{I}_{as} \tau_{as} >> 1$, which would hold even for significantly smaller $\tau_{as}$.
 This leads us to
 \begin{equation}
-\mbox{S/N} \approx \frac{C_p \sqrt{\Delta t}}{ I_{as} \sqrt{\tau_{as}}}. 
-\end{equation}
-Such an observation is therefore \emph{speckle noise limited}, and the S/N improves with square-root of the number of realizations of the speckle pattern as $\sqrt{\Delta t/ \tau_{as}}$.
+\mbox{S/N} \approx \frac{C_p \sqrt{\Delta t}}{ \mathcal{I}_{as} \sqrt{\tau_{as}}}. 
+\end{equation}
+Such an observation is therefore \emph{speckle-noise limited}, and the S/N improves with square-root of the number of realizations of the speckle pattern as $\sqrt{\Delta t/ \tau_{as}}$.
 The ratio of the time to reach a given S/N when speckle limited, $\Delta t_{sl}$, to the time needed when photon noise limited, $\Delta t_{pn}$ is
 \begin{equation}
-\frac{\Delta t_{sl}}{\Delta t_{pn}} = I_*I_{as} \tau_{as} >> 1.
+\frac{\Delta t_{sl}}{\Delta t_{pn}} = I_*\mathcal{I}_{as} \tau_{as} \gg 1.
 \label{eqn:relexptime}
@@
 
-Understanding the speckle lifetime is therefore crucial to understanding the limits of high contrast imaging.
+Understanding the speckle lifetime is therefore crucial to understanding the limits of high-contrast imaging.
 To that end, we next show how to calculate the statistical correlation time in a stochastic process from its PSD, and then show how to apply it to post-coronagraph speckles.
@@
 \label{sec:psd_lifetime}
-We will denote the temporal PSD as $\mathcal{T}(f)$ where $f$ is frequency.
+\jrmadd{Starting with the intensity time series $I(t)$ measured at the location of a planet,} the temporal PSD $\mathcal{T}(f)$ \jrmadd{is defined as the ensemble average of the modulus-squared of its Fourier transform}:
+\begin{equation}
+\mathcal{T}(f) = \left\langle \left| \mathcal{F}_t \left\{ I(t) \right\} \right|^2 \right\rangle.
+\label{eqn:psd_def}
+\end{equation}
+\jrmadd{where $\mathcal{F}_t\{\cdot\}$ denotes the Fourier transform in one dimension (the time-domain) and $f$ is temporal frequency}.
 We will restrict ourselves to one-sided PSDs, that is $f \ge 0$.
@@
 
-According to the Wiener-Khinchin theorem, the PSD \jrmadd{is the Fourier transform of the autocorrelation}.
-This means that the PSD describes how the measurements derived from the process are correlated.
+\jrmadd{We assume that $I(t)$ is a wide-sense stationary (WSS) process.
+This means that the mean is constant and the autocorrelation does not depend on time.
+With the WSS assumption,} the Wiener-Khinchin theorem states that the PSD \jrmadd{is the Fourier transform of the autocorrelation}.
+The PSD \jrmadd{then} describes how the measurements derived from the process are correlated.
 \jrmadd{\citet{1976JOSA...66..207N} exploited this concept in two dimensions to derive the coefficient covariances in the Zernike polynomial expansion of Kolmogorov turbulence phase screens.
@@
 Similar to the Zernike polynomials, this basis has an intuitive progression through slope, quadratic, cubic, etc, modes (see Figure \ref{fig:onlegpol}).
-We later utilize these 1-dimensional modes to analyze post-processing as a time-domain filter.
-For that analysis it will be convenient to use a normalized form of Legendre polynomials defined as:}
-\begin{equation}
-P(\xi) = \sqrt{\frac{2n+1}{2}}\, \mathscr{P}_n(\xi)
+We later utilize these one dimensional modes to analyze post-processing as a time-domain filter.
+For that analysis it will be convenient to use a normalized form of Legendre polynomials, defined as}
+\begin{equation}
+P(\xi) = \sqrt{\frac{2n+1}{2}}\, \mathscr{P}_n(\xi),
 \label{eqn:basis}
@@
 
-\jrmadd{Now we} assume that we obtain a sample $I(t)$ of length $T$ consisting of measurements uniformly spaced by $\Delta t$.
-For notational simplicity we define the time coordinate such that $-T/2 \le t \le T/2$, with no loss of generality.
+\jrmadd{Now we} obtain a sample of $I(t)$ of length $T$ consisting of measurements uniformly spaced by $\Delta t$.
+For simplicity in notation we define the time coordinate such that $-T/2 \le t \le T/2$, with no loss of generality.
 Next, we express this time-series as a linear expansion using the normalized Legendre polynomials.
@@
 \end{equation}
-\jrmadd{The coefficients are variable quantities.
-The best fit of the polynomials will change with time as the speckle intensity varies.
-By analyzing the temporal statistics of the $a_n$, we learn about the power in specific Legendre ``modes'' over the time $T$ just as we do with the analogous Zernike modes in a wavefront over an aperture of diameter $D$ in \citet{1976JOSA...66..207N}.}
-
-The first orthonormal Legendre polynomial is $P_0(\xi) = 1/\sqrt{2}$, which means that $a_0/\sqrt{2}$ is the mean value of $I(t)$.  Thus the variance of the mean is given by
+\jrmadd{The coefficients $a_n$ are variable quantities, meaning that they will change with time as the speckle intensity varies.
+By analyzing the} \jrmrmv{temporal} \jrmadd{ensemble statistics of the $a_n$, we learn about the power in specific Legendre ``modes'' over the time $T$ just as we do with the analogous Zernike modes in a wavefront over an aperture of diameter $D$ \citep{1976JOSA...66..207N}.}
+
+The first orthonormal Legendre polynomial is $P_0(\xi) = 1/\sqrt{2}$, which means that $a_0/\sqrt{2}$ is the mean value of $I(t)$ \jrmadd{over a time-period of length $T$}.
+Thus the variance of the mean is given by
 \begin{equation}
@@
 \end{equation}
-where $J_n$ denotes the cylindrical Bessel functions of the first kind.  This expression gives the variance of the mean in a process governed by $\mathcal{T}(f)$.
+where $J_n$ denotes the cylindrical Bessel functions of the first kind.  This expression gives the variance of the mean in a process governed by $\mathcal{T}(f)$ sampled in chunks of length $T$.
 
@@
 \end{equation}
-Evaluating Equation (\ref{eqn:varmean}), the variance of the mean after a time $t$ is 
-\begin{equation}
-\sigma_\mathrm{mean}^2 = \frac{\beta}{2t}.
+Evaluating Equation (\ref{eqn:varmean}), the variance of the mean after a time $T$ is 
+\begin{equation}
+\sigma_\mathrm{mean}^2 = \frac{\beta}{2T}.
 \label{eqn:vm_wn}
@@
 \begin{equation}
-\sigma_\mathrm{mean}^2 =  \frac{\sigma_o^2 \Delta t}{t} = \frac{\sigma_o^2}{N}
-\end{equation}
-where $N = t/\Delta t$ is the number of measurements.
+\sigma_\mathrm{mean}^2 =  \frac{\sigma_o^2 \Delta t}{T} = \frac{\sigma_o^2}{N}
+\end{equation}
+where $N = T/\Delta t$ is the number of measurements.
 This is exactly the result we expect for the variance of the mean in a white noise process with variance $\sigma^2_o$, \jrmadd{giving the familiar ``root-N'' reduction in noise.}
@@
 \end{equation}
-The variance of the mean decreases with time according to
-\begin{equation}
-\sigma_\mathrm{mean}^2 =  \frac{\sigma_o^2}{t/\tau} \tag{\ref{eqn:varmean_def_intro} restated}
-\end{equation}
-as long as $t >> \tau$.
+The variance of the mean then decreases with time according to
+\begin{equation}
+\sigma_\mathrm{mean}^2 =  \frac{\sigma_o^2}{T/\tau} \tag{\ref{eqn:varmean_def} restated}
+\end{equation}
+as long as $T \gg \tau$.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Speckle Intensity PSDs}
+\section{Speckle Intensity Temporal PSDs}
 \label{sec:intensity}
@@
 \end{equation}
-where $m$ and $n$ are integer indices, $D$ is the aperture diameter, and $(\hat{\mathpzc{u}},\hat{\mathpzc{v}})$ are unit vectors.
+where $m$ and $n$ are integer indices\footnote{Note that this $n$ is different from that used in the Legendre indices.
+The difference will be clear from the context.}, $D$ is the aperture diameter, and $(\hat{\mathpzc{u}},\hat{\mathpzc{v}})$ are unit vectors.
 The modal basis described by Equation (\ref{eqn:modified_fourier}) is normalized, and has the advantage that on the unobscured circular aperture both $p$ modes at a given $mn$ have the same symmetric spatial and temporal PSD.
@@
 We note that this basis is the one derived for the unobscured circular aperture in a 2D Lomb-Scargle analysis \citep{2020arXiv200110200S}, and, as such, phase-shifted Fourier bases with the same properties exist for more complicated apertures.
-In this modal basis the phase in the pupil plane is
+In this modal basis, the phase in the pupil plane is
 \begin{equation}
@@
 \end{equation}
-where $h_{mn}^{+}$ and $h_{mn}^{-}$ are the coefficients of the modes in the same units as wavelength, and the phase $\Phi$ has units of radians.
+where $h_{mn}^{+}$ and $h_{mn}^{-}$ are the \mpfadd{real} coefficients of the modes in the same units as wavelength, and the phase $\Phi$ has units of radians.
 
@@
 \begin{equation}
-I_\Phi(\vec{r},t) \approx \left(\frac{2\pi}{\lambda}\right)^2 2 \sum_{mn} \left\{  \left[ (h_{mn}^{+})^2 + (h_{mn}^{-})^2 \right] \left[ \Ji^2(\pi D k_{mn}^+) +\Ji^2(\pi D k_{mn}^-)\right] + \mbox{cross-terms} \right\}
+I_\Phi(\vec{r},t) \approx \left(\frac{2\pi}{\lambda}\right)^2 2 \sum_{mn} \left\{  \left[ (h_{mn}^{+}(t))^2 + (h_{mn}^{-}(t))^2 \right] \left[ \Ji^2(\pi D k_{mn}^+) +\Ji^2(\pi D k_{mn}^-)\right] + \mbox{cross-terms} \right\}
 \label{eqn:cross_terms}
@@
 \begin{eqnarray}
-k_{mn}^+ &=& \sqrt{\left(\frac{r_x}{\lambda} + \frac{m}{d}\right) + \left(\frac{r_y}{\lambda} + \frac{n}{d}\right)} \\ 
-k_{mn}^- &=& \sqrt{\left(\frac{r_x}{\lambda} - \frac{m}{d}\right) + \left(\frac{r_y}{\lambda} - \frac{n}{d}\right)}. \nonumber
+k_{mn}^+ &=& \sqrt{\left(\frac{r_x}{\lambda} + \frac{m}{D}\right) + \left(\frac{r_y}{\lambda} + \frac{n}{D}\right)} \\ 
+k_{mn}^- &=& \sqrt{\left(\frac{r_x}{\lambda} - \frac{m}{D}\right) + \left(\frac{r_y}{\lambda} - \frac{n}{D}\right)}. \nonumber
 \end{eqnarray}
@@
 
-The quantity we seek is the temporal PSD of the intensity time-series described by Equation (\ref{eqn:cross_terms}).  This is the ensemble average of the modulus-squared of its Fourier transform:
-\begin{equation}
-\mathcal{T}_{I_{mn}}(f) = \left\langle \left| \mathcal{F}_t \left\{ I_\Phi(\vec{r}_{mn},t) \right\} \right|^2 \right\rangle.
-\end{equation}
-where $\mathcal{F}_t\{\cdot\}$ denotes the Fourier transform \jrmadd{in 1 dimension (the time-domain)}.
+The quantity we seek is the temporal PSD of the intensity time-series described by Equation (\ref{eqn:cross_terms}).
+\jrmadd{The PSD $\mathcal{T}_{I_{mn}}(f)$ is defined as in Equation (\ref{eqn:psd_def})}.
+%is the ensemble average of the modulus-squared of its %Fourier transform:
+%\begin{equation}
+%\mathcal{T}_{I_{mn}}(f) = \left\langle \left| %\mathcal{F}_t \left\{ I_\Phi(\vec{r}_{mn},t) \right\} %\right|^2 \right\rangle.
+%\end{equation}
+%where $\mathcal{F}_t\{\cdot\}$ denotes the Fourier %transform \jrmadd{in 1 dimension (the time-domain)}.
 
@@
 \end{equation}
-We thus arrive at the need to determine the PSD of $(h_{mn}^{+})^2 + (h_{mn}^{-})^2$.  Next we review how the PSDs of $h_{mn}^{+}$ and $h_{mn}^{-}$ can be calculated.
+\mpfrmv{We thus arrive at the need to determine the PSD of}\mpfadd{We therefore directly relate the temporal PSD of $I_\phi$ to the temporal PSD of} $(h_{mn}^{+})^2 + (h_{mn}^{-})^2$.
+\mpfrmv{Next we review how the PSDs of $h_{mn}^{+}$ and $h_{mn}^{-}$ can be calculated.}  \mpfadd{This can be expressed in terms of the PSDs for $h_{mn}^{+}$ and $h_{mn}^{-}$; we review how these can be calculated next.}
 
@@
 
-We express the PSDs of the $h_{mn}^p$ in closed loop ($\mathcal{T}_{\mathrm{cl},mn}(f;g)$) as functions of the open-loop PSD ($\mathcal{T}_{\mathrm{ol},mn}(f)$), the error transfer function (ETF), the measurement noise PSD ($\mathcal{T}_{\mathrm{ph},mn}(f)$), and the noise transfer function (NTF):
+We express the PSDs of the $h_{mn}^p$ in closed loop ($\mathcal{T}_{\mathrm{cl},mn}(f;g)$) as functions of the open-loop (OL) PSD ($\mathcal{T}_{\mathrm{ol},mn}(f)$), the error transfer function (ETF), the measurement noise PSD ($\mathcal{T}_{\mathrm{ph},mn}(f)$), and the noise transfer function (NTF):
 \begin{equation}
@@
 where the control loop gain $g$ is optimized to minimize the residual variance in the closed-loop PSD \citep{1999aoa..book.....R_ch6,2016ApOpt..55..323P}.
-\citet{2018JATIS...4a9001M} derived open-loop PSDs of Fourier modes in frozen-flow turbulence.
+\citet{2018JATIS...4a9001M} derived OL PSDs of Fourier modes in frozen-flow turbulence.
 In simplified form, the wavefront sensor (WFS) noise is given by
@@
 
-We consider ETFs and NTFs for control laws based on linear filters of the form
+We consider ETFs and NTFs for control laws based on \jrmadd{general} linear filters of the form
 \begin{equation}
@@
 
-With $J=1$, $a_1 = 1$, $L=0$, and $b_0 = 1$ Equation (\ref{eqn:linfilt}) represents the simple integrator (SI), the simplest control law.
+With $J=1$, $a_1 = 1$, $L=0$, and $b_0 = 1$, Equation (\ref{eqn:linfilt}) represents the \mpfadd{simplest control law, the} simple integrator (SI)\mpfrmv{, the simplest control law}.
 \citet{2018JATIS...4a9001M} showed how adding more coefficients, using the well-known Linear Prediction (LP) technique, results in significant performance improvements in terms of post-coronagraph contrast.
-Most, if not all, controllers can be cast in the form of Equation (\ref{eqn:linfilt}) \citep{2007JOSAA..24.2645P, 2021arXiv210307566H}.
+Most, if not all, linear controllers can be cast in the form of Equation (\ref{eqn:linfilt}) \citep{2007JOSAA..24.2645P, 2021arXiv210307566H}.
 
@@
 \begin{enumerate}
-\item Generate an open-loop time-series for each $h_{mn}^p$ for both parities.
+\item Generate an OL time-series for each $h_{mn}^p$ for both parities.
 \jrmadd{Note that $h_{mn}^-$ is needed for the spatial correlation step, even though it is discarded eventually.} \label{step:olh}
@@
 \item Fourier transform the white-noise time-series
-\item Multiply the Fourier transform by the square-root of the open-loop PSD.  We now have $\mathcal{F}_t\left\{h_{mn}^p\right\}$ \citep[cf.][]{kasdin_falpha}.
+\item Multiply the Fourier transform by the square-root of the OL PSD.  We now have $\mathcal{F}_t\left\{h_{mn}^p\right\}$ \citep[cf.][]{kasdin_falpha}.
 \item Apply the inverse Fourier transform to produce $h_{mn}^p(t)$.  \jrmadd{If the correlation matrix will be applied in Step \ref{step:corr}, then} normalize to unit variance.
@@
 \end{enumerate}
-\item \jrmadd{[Optional]} Correlate the modal time-series by multiplying the vector of coefficients at each time-step by the decomposition of the covariance matrix. \label{step:corr}
+\item \jrmadd{Optionally,} Correlate the modal time-series by multiplying the vector of coefficients at each time-step by the decomposition of the covariance matrix. \label{step:corr}
 \item For each spatial frequency $mn$, replace $h_{mn}^-(t)$ with the phase shifted $h_{mn}^+(t)$.
@@
 An idealized deformable mirror was used, which exactly produced commanded shapes as projection of modes.
-The DM \jrmadd{was capable of making 2400 Fourier modes}, corresponding \jrmadd{to the Nyquist limit of a} 48x48 actuator illuminated pupil.
+The DM \jrmadd{was capable of making 2,400 Fourier modes}, corresponding \jrmadd{to the Nyquist limit of a} 48$\times$48 actuator illuminated pupil.
 The wavefront sensor (WFS) worked by simply projecting modes onto the phase screen.
@@
 
-The simulated system operated at 2000 Hz with a 1.5 frame delay in addition to the 1 frame total sample and hold.
+The simulated system operated at 2,000 Hz with a 1.5 frame delay in addition to the 1 frame total sample and hold \jrmadd{for WFS integration}.
 To achieve the 1/2 frame, the WFS used the average turbulence between two time-steps.
@@
 Turbulence parameters corresponded to the median LCO model used in \citet{2018JATIS...4a9001M}, where $r_0 = 0.16$ m, $L_0 = 25$ m, and layer averaged wind speed was 18.7 m/s.
-The control laws, both SI and LP with 21 coefficients, were optimized on the open-loop PSDs measured by running the simulation with the control loop off.
+The control laws, both SI and LP with 21 coefficients, were optimized on the OL PSDs measured by running the simulation with the control loop off.
 
@@
 Figure \ref{fig:contrast_2Dcomp} shows post-coronagraph contrast maps, comparing the results of calculations using \citet{2018JATIS...4a9001M} to the simulation outputs.
-We show results for an $-\infty$ mag guide star to show the effects of the input turbulence alone, and for an $8$th mag guide star.
+We show results for an \jrmadd{infinitely bright} ($-\infty$ mag) guide star to show the effects of the input turbulence alone, and for an $8$th-mag guide star.
 Figure \ref{fig:contrast_radprof} shows the median radial profile for each of the maps in Figure \ref{fig:contrast_2Dcomp}.
@@
 Note that the color bar is adjusted to best show detail.
-Agreement between the calculations and simulations is very good.\label{fig:contrast_2Dcomp}}
+\jrmadd{The agreement between the calculations and simulations shows that the model is accurate}.\label{fig:contrast_2Dcomp}}
 \end{figure}
@@
 Left: without photon noise, considering the input turbulence only.
-Right: with photon noise on an 8th mag star assuming the sensitivity of an unmodulated pyramid WFS.
-Agreement between the calculations and simulations is very good.\label{fig:contrast_radprof}}
+Right: with photon noise on an 8th-mag star assuming the sensitivity of an unmodulated pyramid WFS.
+\jrmadd{The agreement between the calculations and simulations shows that the model is accurate}.
+\label{fig:contrast_radprof}}
 \end{figure}
@@
 
-In the top row, the black lines compare the calculated (dot-dashed) and simulated (solid) open-loop (OL) PSD.
+In the top row, the black lines compare the calculated (dot-dashed) and simulated (solid) OL PSD.
 The blue lines compare the output of the calculations to the simulations for the SI, and red lines show the LP.
@@
 This better low-frequency performance is true on the 8th mag star for both SI and LP.
-\jrmadd{The difference in integrated power at low frequencies is small and does not appreciably impact final contrast.}
+\jrmadd{The difference in integrated power at low frequencies is small and does not appreciably impact final contrast, and as we show in the following section leads to good results in intensity.}
 
@@
 
-Overall the comparison of outputs presented here give us confidence that the semi-analytic model gives \jrmadd{reasonable} results \jrmadd{[JRM NOTE: softened from valid]}, and can be used to analyze the closed-loop contrast performance and wavefront dynamics in a coronagraphic AO system.
+Overall the comparison of outputs presented here give us confidence that the semi-analytic model gives \jrmadd{reasonable} results, and can be used to analyze the closed-loop contrast performance and wavefront dynamics in a coronagraphic AO system.
 
@@
 Importantly, the solid lines are in nearly perfect agreement with the points, showing the validity of Equation (\ref{eqn:varmean}).
-The dot-dashed lines, \jrmadd{which are the output of the semi-analytic calculations}, are slightly offset for the LP, which is due to the difference in simulated vs.
-predicted contrast (equivalent to the overall variance level), but the shape of the predicted evolution with time for both SI and LP controllers is in good agreement with the simulations.
+The dot-dashed lines, \jrmadd{which are the output of the semi-analytic calculations}, are slightly offset for the LP, which is due to differences in simulated vs.
+predicted contrast (equivalent to the overall variance level) as found in the previous section, but the shape of the predicted evolution with time for both SI and LP controllers is in good agreement with the simulations.
 
@@
 The first comparison we make is between control laws.
-Here we simply calculate a histogram of the lifetimes over the 48x48 $\lambda/D$ dark hole for the SI and LP controller.
+Here we simply calculate a histogram of the lifetimes over the 48$\times$48 $\lambda/D$ dark hole for the SI and LP controller.
 We also apply the Monte Carlo procedure to the OL case, \jrmadd{by treating the input PSDs as unmodified by control laws.
@@
 \jrmadd{Next, we analyze the change in speckle lifetime with guide star brightness, telescope diameter, wind speed, and seeing.
-Note that power per Fourier mode scales as $1/D^2$.
+Note that power per Fourier mode scales as $D^{-2}$.
 We held relative spatial frequency constant with $D$, such that the resultant speckle occurs at the same relative projected separation in $\lambda/D$ units.
@@
 
-\jrmadd{Figure \ref{fig:scalings} compares speckle lifetimes for a 5th magnitude guide star to those for an 8th mag guide star.
-The lifetime is generally shorter on the 8th mag star due to the increased influence of photon noise relative to the turbulence PSD, with the overall trends being similar.
+\jrmadd{Figure \ref{fig:scalings} compares speckle lifetimes for a 5th-magnitude guide star to those for an 8th-mag guide star.
+The lifetime is generally shorter on the 8th-mag star due to the increased influence of photon noise relative to the turbulence PSD, with the overall trends being similar.
 Lifetimes are shorter under LP control compared to SI control for both guide stars for all parameters.}
@@
 For longer crossing times the trend becomes flat.
-This implies that the control system becomes dominated by photon noise rather than servo-lag and so the optimization process no longer produces changes in the output dynamics.}
+This implies that the control system becomes dominated by photon noise rather than servo lag and so the optimization process no longer produces changes in the output dynamics.}
 
 \jrmadd{The LP controller shows much weaker dependence on the same parameters.
-Except at the lowest wind speeds, the speckle lifetime is $\sim$$20$ ms on the 5th magnitude guide star.
-On the 8th magnitude star photon noise makes predictive control less effective, and there is a proportional scaling up to $D/\mathcal{V}$$\sim$$3$.}
+Except at the lowest wind speeds, the speckle lifetime is $\sim$$20$ ms on the 5th-magnitude guide star.
+On the 8th-magnitude star photon noise makes predictive control less effective, and there is a proportional scaling up to $D/\mathcal{V}$$\sim$$3$.}
 
@@
 As the turbulence power is reduced, the control system is better able to reject low temporal frequencies, resulting in lower speckle lifetimes.
-Interestingly, on the 8th mag guide star and for the largest apertures the LP controller shows the opposite trend.
+Interestingly, on the 8th-mag guide star and for the largest apertures the LP controller shows the opposite trend.
 The power in each Fourier mode is proportional to $1/D^2$.
@@
 Furthermore, given the highest jet stream driven wind speeds we can expect, the GSMTs will normally operate in the regime of $D/\mathcal{V} \gtrsim 1$.
-Thus the GSMTs are operating in a different regime compared to the 6.5 m, and the LP controller produces different dynamics  when photon noise becomes important on the 8th mag star.}
+Thus the GSMTs are operating in a different regime compared to the 6.5 m, and the LP controller produces different dynamics  when photon noise becomes important on the 8th-mag star.}
 
@@
 \end{equation}
-After fitting and subtracting N modes the speckle intensity PSD will be high-pass filtered with squared transfer function
+After fitting and subtracting $N$ modes, the speckle intensity PSD will be high-pass filtered with squared transfer function
 \begin{equation}
@@
 
-Now the fit to the data will have a fractional uncertainty $\gamma$, so the resulting speckle intensity PSD will be
+Now due to measurement noise the fit to the data will have a fractional uncertainty $\gamma$, so the resulting speckle intensity PSD will be
 \begin{equation}
@@
 \includegraphics[width=4in]{relsens.pdf}
-\caption{The relative sensitivity, in terms of exposure time, after subtracting the first N Legendre polynomials with uncertainty in the fit $\gamma$.
+\caption{The relative sensitivity, in terms of exposure time, after subtracting the first $N$ Legendre polynomials with uncertainty in the fit $\gamma$ for each $a_n$.
 The horizontal line at 1 corresponds to the photon-noise limited exposure time on an 8th mag star.
 Residual-atmospheric speckles increase the exposure time by a factor 55 compared to the photon-noise limit.
-WFS telemetry-based PSF subtraction of just the first $\sim$5 low-order Legendre polynomials from the speckle time-series reduces the speckle lifetime, improving sensitivity dramatically.
+WFS telemetry-based PSF subtraction of just the first $\sim$5 low-order Legendre polynomials from the speckle time series reduces the speckle lifetime, improving sensitivity dramatically.
 If better than $\sim$$1\%$ accuracy in reconstruction of these modes can be achieved, the speckle noise can be reduced to below the photon noise limit.
@@
 If the the first $\sim$5 low-order Legendre polynomials can be determined from, say, WFS-based reconstruction, the reduction in speckle lifetime from removing the low-order temporal modes dramatically improves sensitivity by orders of magnitude.
-If the modes are reconstructed to better than $\sim$1\% accuracy, the photon-noise limit can be reached.
+If the modes are reconstructed to better than $\sim$1\% accuracy in the $a_n$, the photon-noise limit \jrmadd{in the focal plane} can be reached.
 
@@
 In this context, WSS essentially implies that the atmosphere parameters, namely $r_0$ and the layer $\vec{\mathcal{V}}$, are constant.
-This is of course not strictly true, but can be expected to be approximately true over short periods of time -- at least long enough to estimate a PSD.
-We also expect AO control to damp changes, keeping the output speckle formation process within a more or less well defined range.
+This is of course not strictly true, but can be expected to be approximately true over short periods of time --- at least long enough to estimate a PSD.
+We also expect AO control to damp changes, keeping the output speckle formation process within a more or less well-defined range.
 What this ultimately means is that we must consider the speckle lifetime as itself a time-variable parameter, but expect there to be a representative value over an observation that characterizes long exposure sensitivity.}
@@
 \jrmadd{We have developed a semi-analytic model of a coronagraphic AO system that allows us to estimate long-exposure contrast, the intensity dynamics, and statistical lifetimes.
-Compared to an idealized end-to-end simulation, the results of the model compare very well.
-This powerful technique allows us to analyze the impact of control law choice, as well as to compare the influence of various telescope and atmosphere parameters on the speckle lifetime.
-Key benefits of such modeling are that the true long-exposure performance of a system can be determined without the expense of long simulations.
+Compared to an idealized end-to-end simulation, the results of the model compare favorably.
+This powerful technique allows us to analyze the impact of control-law choice, as well as to compare the influence of various telescope and atmosphere parameters on the speckle lifetime.
+Key benefits of such modeling are that the true long-exposure performance of a system can be determined without the expense of simulations.
 For instance, the parameter study in Section \ref{sec:trends} would have been prohibitively expensive if conducted with simulations.
@@
 \jrmadd{We analyzed various parameters with potential impacts on speckle lifetime.
-Overall, the simplest summary of this analysis is that there is no single thumb rule type relationship that applies to all parameters.
+Overall, the simplest summary of this analysis is that there is no single thumb-rule-type relationship that applies to all parameters.
 This modestly contradicts the conclusions of \citet{2005SPIE.5903..170M}.
@@
 As a result, neither the control action nor the overall power (seeing) should impact speckle lifetime.
-Only the wind would change speckle lifetime for a given diameter.}
+Only the wind would change speckle lifetime for a given \mpfadd{pupil} diameter.}
 
 \jrmadd{Here we have added the time dimension, fully considering the action of AO control as a temporal filter.
-Furthermore, we have modeled the impact of high order predictive control.
+Furthermore, we have modeled the impact of high-order predictive control.
 With this more complete view, we find that the control action does reduce speckle lifetime.
@@
 \subsection{Post-Processing}
-The temporal decorrelation of speckle noise is one of the main benefits of PSF-subtraction post-processing \citep{2006ApJ...641..556M,2008ApJ...673..647M}, even if algorithms have typically been focused on the spatial domain.
+The temporal decorrelation of speckle noise is one of the main benefits of PSF-subtraction post-processing \citep{2006ApJ...641..556M,2008ApJ...673..647M}, even though algorithms have typically focused on exploiting spatial correlations.
 Recently, studies have begun to consider the time-domain in post-processing explicitly \citep{2021A&A...646A..24S}.
@@
 The key development of the post-coronagraph intensity PSD allowed us to analyze post-processing as a time-domain filter and show that it is possible to achieve photon-noise limited sensitivity given a suitable reconstruction algorithm.
-In future work we will use this framework to analyze more realistic turbulence models, and extend these techniques to instrumental quasi-static aberrations.
+In future work we will use this framework to analyze more realistic turbulence models, extend these techniques to instrumental quasi-static aberrations, \jrmadd{and analyze various post-processing strategies}.
 
@@
 
-We now have \jrmadd{the covariances of the Legendre coefficients}
-\begin{equation}
-<a_n^* a_{n'}> = -i^n i^{n'} \frac{\sqrt{2n+1}\sqrt{2n'+1}}{T}   \int_{0}^{\infty} \frac{J_{n+\frac{1}{2}}(2\pi \kappa) J_{n'+\frac{1}{2}}(2\pi \kappa)}{\kappa} \mathcal{T}\left( \frac{2}{T} \kappa \right) d\kappa
+We now have \jrmadd{the covariance of the Legendre coefficients}
+\begin{equation}
+\langle a_n^* a_{n'}\rangle = -\mpfrmv{i^n i^{n'}}\mpfadd{i^{n+n'}} \frac{\sqrt{2n+1}\sqrt{2n'+1}}{T}   \int_{0}^{\infty} \frac{J_{n+\frac{1}{2}}(2\pi \kappa) J_{n'+\frac{1}{2}}(2\pi \kappa)}{\kappa} \mathcal{T}\left( \frac{2}{T} \kappa \right) d\kappa
 \end{equation}

### 2021-05-23 specklelives.tex
update recon method

@@
 
-%\usepackage{indentfirst}
-%\usepackage{lscape}
 \usepackage{afterpage}
-%\usepackage{rotating}
 \let\captionbox\undefined
-%\usepackage{caption}
-
-%\captionsetup{width=6.3in,font=footnotesize}
-%\usepackage{wrapfig}
-%\usepackage{multicol}
-%\usepackage{textcomp}
+
 \newcommand{\textapprox}{\raisebox{0.5ex}{\texttildelow}}
+\newcommand{\Ji}{\mathrm{Ji}}
 
@@
 % custom commands for Olivier's comments
-\usepackage{ulem}
 \newcommand{\ogadd}[1]{\textcolor{orange}{#1}}
@@
 
-%\usepackage{multirow}
-
-%\usepackage{xcolor}
-%\usepackage{ulem}
-
-%\usepackage{soul}
-
-%\setlength{\textwidth}{6.5in}
-%\setlength{\hoffset}{0in}
-%\setlength{\oddsidemargin}{0in}
-%\setlength{\evensidemargin}{0in}
-
-%\setlength{\textheight}{9.in}
-%\setlength{\voffset}{0in}
-%\setlength{\topmargin}{0in}
-%\setlength{\headsep}{0.25in}
-%\setlength{\headheight}{0in}
-
-%Some handy-dandy commands
-%***Distance and Length***
-\newcommand{\meters}{\mbox{m}}
-\newcommand{\cm}{\mbox{cm}}
-\newcommand{\km}{\mbox{km}}
-\newcommand{\dpc}{d_{pc}}
-\newcommand{\au}{\mbox{AU}}
-\newcommand{\microns}{\mu\mbox{m}}
-\newcommand{\rsun}{R_{\sun}}
-\newcommand{\rjup}{R_J}
-\newcommand{\rearth}{R_{\earth}}
-
-%***Times***
-\newcommand{\hours}{\mbox{ hrs}}
-\newcommand{\seconds}{\mbox{ s}}
-\newcommand{\years}{\mbox{ yrs}}
-
-%**Mass**
-\newcommand{\kg}{\mbox{kg}}
-\newcommand{\msun}{M_{\sun}}
-\newcommand{\mearth}{M_{\earth}}
-\newcommand{\mjup}{M_J}
-
-\newcommand{\lambdasci}{\lambda_\mathrm{sci}}
-\newcommand{\lambdawfs}{\lambda_\mathrm{wfs}}
-\newcommand{\tautd}{\tau_\mathrm{td}}
-\newcommand{\tauwfs}{\tau_\mathrm{wfs}}
-\newcommand{\dopt}{d_\mathrm{opt}}
-\newcommand{\mmax}{m_\mathrm{max}}
-\newcommand{\nmax}{n_\mathrm{max}}
-\newcommand{\Ji}{\mathrm{Ji}}
-\newcommand{\pone}{ {\scalebox{0.6}{+1}}}
-\newcommand{\mone}{ {\scalebox{0.6}{-1}} }
-
-\newcommand{\karman}{K\'{a}rm\'{a}n }
-
-\newcommand{\argmax}{\operatornamewithlimits{argmax}}
-\newcommand{\argmin}{\operatornamewithlimits{argmin}}
+% custom commends for Res's comments
+\newcommand{\rbadd}[1]{\textcolor{red}{#1}}
+\newcommand{\rbrmv}[1]{\textcolor{red}{\sout{#1}}}
 
@@
 
+% \usepackage{background}
+% \backgroundsetup{
+%   position=current page.north,
+%   angle=0,
+%   nodeanchor=north,
+%   vshift=10mm,
+%   hshift=50mm,
+%   opacity=1,
+%   scale=1,
+%   color=black,
+%   contents={Draft -- submitted to PASP 2021 March 29}
+% }
+
 \begin{document}
 
-\title{The Mysterious Lives Of Speckles. I. Residual atmospheric speckle lifetimes behind AO-fed coronagraphs}
+\title{The Mysterious Lives Of Speckles. I. Residual atmospheric speckle lifetimes in ground-based coronagraphs}
 
@@
 High-contrast imaging observations are fundamentally limited by the spatially and temporally correlated noise source called speckles.
-Suppression of speckle \jrmadd{noise} is the key goal of wavefront control and adaptive optics (AO), coronagraphy, and a host of post-processing techniques.
+Suppression of speckle noise is the key goal of wavefront control and adaptive optics (AO), coronagraphy, and a host of post-processing techniques.
 The temporal correlation of speckles reduces sensitivity as they average much more slowly than photon noise.
-\mpfadd{In speckle-limited integrations, the rate at which speckle intensity fluctuations approach the mean intensity is a function of the second-order temporal statistics.
-Using these statistics, we can define a speckle lifetime which describes how quickly speckles average in long exposures.} \mpfrmv{Speckles average at a rate set by the statistical speckle lifetime, and speckle-limited integration time in long exposures is directly proportional to this lifetime.}  As progress continues in post-AO coronagraphic wavefront control, residual atmospheric speckles will become the limiting noise source in high-contrast imaging, so a complete understanding of their statistical behavior is crucial to optimizing high-contrast imaging instruments.
+Speckles average at a rate set by the statistical speckle lifetime, and speckle-limited integration time in long exposures is directly proportional to this lifetime.
+As progress continues in post-AO coronagraphic wavefront control, residual atmospheric speckles will become the limiting noise source in high-contrast imaging, so a complete understanding of their statistical behavior is crucial to optimizing high-contrast imaging instruments.
 Here we present a novel power spectral density (PSD) method for calculating the lifetime, and develop a semi-analytic method for predicting intensity PSDs behind a coronagraph.
@@
 We find that standard AO control shortens atmospheric speckle lifetime from $\sim$130 ms to $\sim$50 ms, and predictive control will further shorten the lifetime to $\sim$20 ms on 6.5 m MagAO-X.
-We find that speckle lifetimes vary with diameter, wind speed, seeing, and location with the AO control region (dark hole).
-No broadly applicable ``rules of thumb'' are found, but on bright stars lifetimes appear to remain within a rough range of $\sim$20 ms to $\sim$100 ms.
+We find that speckle lifetimes vary with diameter, wind speed, seeing, and location within the AO control region.
+On bright stars lifetimes remain within a rough range of $\sim$20 ms to $\sim$100 ms.
 Due to control system dynamics there are no simple scaling laws which apply across a wide range of system characteristics.
-Finally, we use these results to describe how post-processing will enable ground-based telescopes to achieve the photon-noise limit in high-contrast imaging.
+Finally, we use these results to argue that telemetry-based post-processing should enable ground-based telescopes to achieve the photon-noise limit in high-contrast imaging.
 \end{abstract}
+
+%\mpfadd{In speckle-limited integrations, the rate at which speckle intensity fluctuations approach the mean intensity is a function of the second-order temporal statistics.
+Using these statistics, we can define a speckle lifetime which describes how quickly speckles average in long exposures.}
 
@@
 In addition to the photon noise caused by this halo, the scattered light is coherent and produces ``speckles,'' which can be thought of as copies of the instrument point spread function (PSF) \citep{1995PASP..107..386M}.
-This adds structured noise, with both spatial and temporal correlations, which is the limiting noise source in a \ogadd{long exposure} coronagraphic observation\ogadd{s}.
+This adds structured noise, with both spatial and temporal correlations, which is the limiting noise source in long exposure coronagraphic observations.
 Both in space and on the ground, imperfections within the telescope and instrument optics cause speckles.
-For \ogadd{ground-based} telescopes, the aberrations \ogadd{also} include those imposed by \ogadd{atmospheric} turbulence.
-
-\jrmadd{[MOVED FROM DISCUSSION] For some time, long-lived ``quasi-static'' speckles have been recognized as the limiting noise source in contemporary high contrast imaging observations \citep[e.g.][]{2007ApJ...654..633H}.
+For ground-based telescopes, the aberrations also include those imposed by atmospheric turbulence.
+
+For some time, long-lived ``quasi-static'' speckles have been recognized as the limiting noise source in contemporary high contrast imaging observations \citep[e.g.][]{2007ApJ...654..633H}.
 Given their origin within the instrument, and comparatively long lifetimes \citep{2012A&A...541A.136M}, such speckles should be suppressible using wavefront sensing and control (WFS\&C) strategies.
@@
 Once active quasi-static speckle control becomes routine, the fundamental sensitivity limit in ground-based high-contrast imaging will be set by the Earth's atmosphere.
-}
 
@@
 
-\mpfadd{Speckles also arise from atmospheric turbulence.}  The index of refraction of air depends on temperature, and turbulence causes the air temperature to stochastically vary across the telescope aperture.
+Speckles also arise from atmospheric turbulence.
+The index of refraction of air depends on temperature, and turbulence causes the air temperature to stochastically vary across the telescope aperture.
 As the stellar wavefront traverses these regions of different index of refraction, parts of the wavefront travel different optical path lengths.
-Due mainly to wind\mpfadd{-driven bulk flow across the telescope aperture}, the atmospheric temperature structure is constantly changing.
+Due mainly to wind-driven bulk flow across the telescope aperture, the atmospheric temperature structure is constantly changing.
 The technique of adaptive optics (AO) is used to correct the resulting time-variable optical aberrations in real time.
-Even with so-called ``extreme'' AO systems \citep[ExAO, ][]{2018ARAA..56..315G} this correction is \mpfrmv{not} \mpfadd{im}perfect.
+Even with so-called ``extreme'' AO systems \citep[ExAO, ][]{2018ARAA..56..315G} this correction is imperfect.
 The AO-corrected wavefront, though significantly improved, still causes a halo of speckles significantly brighter than the instrumental speckles in a modern high performance coronagraph system using coronagraphic wavefront control.
 
-In order to \ogadd{estimate} the capabilities of ground-based high-contrast imaging, we must understand the spatial and temporal evolution of the residual atmospheric speckles behind an ExAO coronagraph.
+In order to estimate the capabilities of ground-based high-contrast imaging, we must understand the spatial and temporal evolution of the residual atmospheric speckles behind an ExAO coronagraph.
 Speckle formation in the spatial domain, as structures within the coronagraphic PSF, is well understood \citep{2001ApJ...558L..71B,2003ApJ...596..702P,2007ApJ...669..642S}.
@@
 While this would still leave the photon noise from the scattered light, we could perfectly remove the spatially correlated noise.
-In the second \mpfrmv{imaginary}\mpfadd{idealized} regime, the speckles are a perfect temporally white noise source: every image taken has a different speckle structure.
-Similar to photon noise, \mpfrmv{it}\mpfadd{high-contrast imaging} would now be a simple matter of taking enough images to average the speckles to a smooth halo.
+In the second idealized regime, the speckles are a perfect temporally white noise source: every image taken has a different speckle structure.
+Similar to photon noise, high-contrast imaging would now be a simple matter of taking enough images to average the speckles to a smooth halo.
 
 In reality, the temporal behavior of atmospheric speckles falls between these regimes.
-These speckles vary on short time-scales \ogadd{so they are not removed by standard PSF subtraction}, but exhibit non-negligible temporal correlations which cause them to average much more slowly than photon noise.
+These speckles vary on short time-scales so they are not removed by standard PSF subtraction, but exhibit non-negligible temporal correlations which cause them to average much more slowly than photon noise.
 The statistical lifetime of atmospheric speckles, that is the time-scale which controls the averaging of speckle noise, is the quantity that sets the fundamental sensitivity limit of ground-based high-contrast imaging.
@@
 There have been several definitions of the atmospheric speckle lifetime presented in the literature.
-Early treatments of speckle dynamics were motivated by speckle interferometry, which relies on exposures short enough to ``freeze the atmosphere.''  The optimum exposure time is therefore dependent on the \mpfrmv{coherence}\mpfadd{correlation} time of the intensity in the focal plane.
-A common definition for this \mpfrmv{coherence}\mpfadd{correlation} time was based on a $1/e$ reduction in the autocorrelation of the intensity, which was experimentally found to be on the order of a few ms \citep{1978ApOpt..17.3779S,1990JMOp...37.1247D}.
+Early treatments of speckle dynamics were motivated by speckle interferometry, which relies on exposures short enough to ``freeze the atmosphere.''  The optimum exposure time is therefore dependent on the correlation time of the intensity in the focal plane.
+A common definition for this correlation time was based on a $1/e$ reduction in the autocorrelation of the intensity, which was experimentally found to be on the order of a few ms \citep{1978ApOpt..17.3779S,1990JMOp...37.1247D}.
 \citet{1982JOpt...13..263R} derived the following expression for the autocorrelation e-folding time in a multi-layer Kolmogorov atmosphere:
@@
 The autocorrelation e-folding time is a somewhat arbitrary choice.
-If we instead consider the evolution of the variance in the estimate of the mean given uncorrelated trials, we can define the speckle lifetime \jrmadd{according to the familiar ``root-N averaging'' reduction in noise:}
+If we instead consider the evolution of the variance in the estimate of the mean given uncorrelated trials, we can define the speckle lifetime according to the familiar ``root-N'' reduction in noise:
 \begin{equation}
@@
 \end{equation}
-\jrmadd{This simply states that the variance in a measurement of the mean improves with the number of uncorrelated trials $t/\tau$.
-Defined this way,} $\tau$ is given by the integral of the autocorrelation \citep{2006ApJ...637..541F}, and in this expression determines the number of uncorrelated realizations of the speckle pattern included in the mean.
+This simply states that the variance in a measurement of the mean improves with the number of uncorrelated trials $t/\tau$.
+Defined this way, $\tau$ is given by the integral of the autocorrelation \citep{2006ApJ...637..541F}, and in this expression determines the number of uncorrelated realizations of the speckle pattern included in the mean.
 Note that this assumes $t \gg \tau$.
@@
 through Equation \ref{eqn:varmean_def_intro}) may lead to different optimizations than by considering raw contrast alone.
-%Note for comparison that the Greenwood time-constant is
-%\begin{equation}
-%\tau_0 = 0.31 \frac{r_0}{\bar{v}}
-%\end{equation}
-%where $\bar{v}$ is the 5/3-moment weighted mean velocity (cf. Hardy sec 3.3.6).  $\tau_0$ usually ranges from $\sim$2 to $\sim$8 ms, and is a characteristic time-scale 
-%which sets the update rate requirement for AO systems.
 
@@
 Using focal plane data from SPHERE, \citet{2016SPIE.9909E..4ZM} found evidence for speckle decorrelation on several timescales, but were unable to identify an atmospheric component given the various instrument-induced timescales and the relatively low cadence used.
-In a study using SCExAO at much higher frame rates, \citet{2018PASP..130j4502G} analyzed the evolution of the intensity using the variance of difference images (\mpfrmv{essentially }a form of autocorrelation).
+In a study using SCExAO at much higher frame rates, \citet{2018PASP..130j4502G} analyzed the evolution of the intensity using the variance of difference images (a form of autocorrelation).
 Here short timescale (a few ms) speckle evolution was attributed to the atmosphere, and the authors argued that the AO system modulated the speckles on longer timescales.
@@
 \subsection{Goals of This Study}
-\jrmadd{Our goal in this study is to develop a self-consistent description of the time-domain behavior of speckle noise in AO-fed coronagraphic images from as close to first principles as possible.
+Our goal in this study is to develop a self-consistent description of the time-domain behavior of speckle noise in AO-fed coronagraphic images from as close to first principles as possible.
 Our motivation is to understand the true limits of ground-based high-contrast imaging and so we focus on the statistical lifetimes of atmospheric speckles in coronagraphic focal planes.
-This requires a traversal of speckle statistics, signal processing, wavefront propagation, and control theory\mpfrmv{, and so is somewhat involved}.
+This requires a traversal of speckle statistics, signal processing, wavefront propagation, and control theory.
 To accomplish our goals we will work through the following tasks, organized by section:
@@
 \item Section \ref{sec:trends}: Use the analytic framework to analyze speckle lifetime over a range of telescope diameters and atmospheric parameters.
-\item Section \ref{sec:reconstruction}: \jrmrmv{We propose} \jrmadd{Develop} a method to analyze post-processing in the time-domain and discuss its implications for the limits of ground-based high-contrast imaging sensitivity.
-\item Section \ref{sec:discussion}: \jrmrmv{We then d} \jrmadd{D}iscuss various implications of these results.
+\item Section \ref{sec:reconstruction}: Develop a method to analyze post-processing in the time-domain and discuss its implications for the limits of ground-based high-contrast imaging sensitivity.
+\item Section \ref{sec:discussion}: Discuss various implications of these results.
 \end{itemize}
@@
 Sections \ref{sec:psd_lifetime} through \ref{sec:results} and Section \ref{sec:reconstruction} contain the technical details.
-}
-
-%In the present study, we develop a semi-analytic method for predicting the residual speckle lifetime according to Equation (\ref{eqn:varmean_def_intro}).
-This model explicitly accounts for closed-loop AO control, including advanced predictive control laws.
-We verify the results of the model by comparison to end-to-end simulations of an ideal closed-loop AO system.
-We here concentrate on the speckles due to the atmosphere under the frozen flow hypothesis, and do not consider more complicated turbulence evolution (e.g.
-boiling) nor do we consider instrumental speckles.
-We begin with a review of the theory of speckle noise in coronagraph images, followed by a derivation of the statistical speckle lifetime based on the power spectral density (PSD) of the intensity.
-We then extend the methods of \citet{2018JATIS...4a9001M} to calculate the intensity PSD.
-The model predictions for contrast and speckle lifetimes are compared, and we use the model to predict the behavior of specle lifetimes under various conditions.
-Finally, we use the intensity PSDs to consider the benefits of telemetry-based post-processing.
 
@@
 \label{sec:noise}
-\jrmadd{\mpfrmv{The}\mpfadd{A} key metric for sensitivity in a high-contrast imaging observation is the signal-to-noise ratio, S/N.
-The \jrmrmv{signal we collect from} intensity of a planet, which has planet to star contrast of $C_p$, is }
+A key metric for sensitivity in a high-contrast imaging observation is the signal-to-noise ratio, S/N.  The intensity of a planet, which has planet to star contrast of $C_p$, is 
 \begin{equation}
@@
 \end{equation}
-\jrmadd{where $I_{*}$ is the peak intensity of the star's PSF without a coronagraph, which is a function of time due to Strehl ratio variation.\jrmrmv{, and $\Delta t$ is the exposure time.}  Here we will assume units of photon/sec/$(\lambda/D)^2$ for the intensity, where $\lambda$ is the observation wavelength and $D$ is the telescope diameter.
-$\lambda/D$ is the fundamental spatial scale of diffraction limited imaging and coronagraphy, and we assume that photometry is conducted in pixels of size 1 $\lambda/D$.}
-
-\jrmadd{The dominant noise source in such an observation is the halo of scattered starlight underlying the planet signal.
+where $I_{*}$ is the peak intensity of the star's PSF without a coronagraph, which is a function of time due to Strehl ratio variation.
+Here we will assume units of photon/sec/$(\lambda/D)^2$ for the intensity, where $\lambda$ is the observation wavelength and $D$ is the telescope diameter.
+$\lambda/D$ is the fundamental spatial scale of diffraction limited imaging and coronagraphy, and we assume that photometry is conducted in pixels of size 1 $\lambda/D$.
+
+The dominant noise source in such an observation is the halo of scattered starlight underlying the planet signal.
 At the planet position $\vec{r}_p$ this has intensity $I_h(\vec{r}_p,t)$.
-Following \citet{2007ApJ...669..642S} we initially consider three contributions to the halo intensity:}
+Following \citet{2007ApJ...669..642S} we initially consider three contributions to the halo intensity:
 \begin{equation}
@@
 The halo contributes noise in two distinct ways: photon noise and correlated speckle noise.
-The total halo variance, considering both speckles and and photon noise, \jrmadd{under the planet signal} is \citep{2007ApJ...669..642S}
+The total halo variance, considering both speckles and and photon noise, under the planet signal is \citep{2007ApJ...669..642S}
 \begin{equation}
@@
 
-\jrmadd{The noise is the square-root of the variance, so using Equations (\ref{eqn:signal}) and  (\ref{eqn:variance_as})} the peak signal-to-noise ratio of the planet detection will be
+The noise is the square-root of the variance, so using Equations (\ref{eqn:signal}) and  (\ref{eqn:variance_as}) the peak signal-to-noise ratio of the planet detection will be
 \begin{equation}
@@
 \label{sec:psd_lifetime}
-\jrmadd{Starting with the intensity time series $I(t)$ measured at the location of a planet,} the temporal PSD $\mathcal{T}(f)$ \jrmadd{is defined as the ensemble average of the modulus-squared of its Fourier transform}:
+Starting with the intensity time series $I(t)$ measured at the location of a planet, the temporal PSD $\mathcal{T}(f)$ is defined as the ensemble average of the modulus-squared of its Fourier transform:
 \begin{equation}
@@
 \end{equation}
-\jrmadd{where $\mathcal{F}_t\{\cdot\}$ denotes the Fourier transform in one dimension (the time-domain) and $f$ is temporal frequency}.
+where $\mathcal{F}_t\{\cdot\}$ denotes the Fourier transform in one dimension (the time-domain) and $f$ is temporal frequency.
 We will restrict ourselves to one-sided PSDs, that is $f \ge 0$.
@@
 
-\jrmadd{We assume that $I(t)$ is a wide-sense stationary (WSS) process.
+We assume that $I(t)$ is a wide-sense stationary (WSS) process.
 This means that the mean is constant and the autocorrelation does not depend on time.
-With the WSS assumption,} the Wiener-Khinchin theorem states that the PSD \jrmadd{is the Fourier transform of the autocorrelation}.
-The PSD \jrmadd{then} describes how the measurements derived from the process are correlated.
-\jrmadd{\citet{1976JOSA...66..207N} exploited this concept in two dimensions to derive the coefficient covariances in the Zernike polynomial expansion of Kolmogorov turbulence phase screens.
-Here we adapt that analysis to arbitrary one dimensional PSDs using the Legendre polynomials.}
+With the WSS assumption, the Wiener-Khinchin theorem states that the PSD is the Fourier transform of the autocorrelation.
+The PSD then describes how the measurements derived from the process are correlated.
+\citet{1976JOSA...66..207N} exploited this concept in two dimensions to derive the coefficient covariances in the Zernike polynomial expansion of Kolmogorov turbulence phase screens.
+Here we adapt that analysis to arbitrary one dimensional PSDs using the Legendre polynomials.
 
@@
 
-\jrmadd{In principle, any basis could be used.
+In principle, any basis could be used.
 Our main goal is analyze the statistics of the mean, which corresponds to the first Legendre polynomial.
@@
 We later utilize these one dimensional modes to analyze post-processing as a time-domain filter.
-For that analysis it will be convenient to use a normalized form of Legendre polynomials, defined as}
-\begin{equation}
-P(\xi) = \sqrt{\frac{2n+1}{2}}\, \mathscr{P}_n(\xi),
+For that analysis it will be convenient to use a normalized form of Legendre polynomials, defined as
+\begin{equation}
+P_{\eta}(\xi) = \sqrt{\frac{2\eta+1}{2}}\, \mathscr{P}_\eta(\xi),
 \label{eqn:basis}
 \end{equation}
-where $\mathscr{P}_n(\xi)$ are the Legendre polynomials defined in the usual manner \citep[cf.][]{Olver:2010:NHMF} on the domain $-1 \le \xi \le 1$.
+where $\mathscr{P}_\eta(\xi)$ are the Legendre polynomials defined in the usual manner \citep[cf.][]{Olver:2010:NHMF} on the domain $-1 \le \xi \le 1$.
 The first six of these functions are shown in Figure \ref{fig:onlegpol}.
@@
 \begin{equation}
-\int_{-1}^{1}P_n(\xi)P_{n'}(\xi) d\xi = \delta_{nn'}
+\int_{-1}^{1}P_\eta(\xi)P_{\eta'}(\xi) d\xi = \delta_{\eta\eta'}
 \label{eqn:orthonormal}
 \end{equation}
-where $\delta_{nn'}$ is the Kronecker delta.
-
-\jrmadd{Now we} obtain a sample of $I(t)$ of length $T$ consisting of measurements uniformly spaced by $\Delta t$.
+where $\delta_{\eta\eta'}$ is the Kronecker delta.
+
+Now we obtain a sample of $I(t)$ of length $T$ consisting of measurements uniformly spaced by $\Delta t$.
 For simplicity in notation we define the time coordinate such that $-T/2 \le t \le T/2$, with no loss of generality.
@@
 \begin{equation}
-I(t) = I\left(\frac{T}{2}\xi\right) = \sum_n a_n P_n(\xi)
+I(t) = I\left(\frac{T}{2}\xi\right) = \sum_\eta a_\eta P_\eta(\xi)
 \label{eqn:basisexpansion}
@@
 \begin{equation}
-a_n = \int_{-1}^{1} I\left(\frac{T}{2}\xi\right)P_n(\xi) d\xi.
+a_\eta = \int_{-1}^{1} I\left(\frac{T}{2}\xi\right)P_\eta(\xi) d\xi.
 \label{eqn:expansecoeff}
 \end{equation}
-\jrmadd{The coefficients $a_n$ are variable quantities, meaning that they will change with time as the speckle intensity varies.
-By analyzing the} \jrmrmv{temporal} \jrmadd{ensemble statistics of the $a_n$, we learn about the power in specific Legendre ``modes'' over the time $T$ just as we do with the analogous Zernike modes in a wavefront over an aperture of diameter $D$ \citep{1976JOSA...66..207N}.}
-
-The first orthonormal Legendre polynomial is $P_0(\xi) = 1/\sqrt{2}$, which means that $a_0/\sqrt{2}$ is the mean value of $I(t)$ \jrmadd{over a time-period of length $T$}.
+The coefficients $a_\eta$ are variable quantities, meaning that they will change with time as the speckle intensity varies.
+By analyzing the ensemble statistics of the $a_\eta$, we learn about the power in specific Legendre ``modes'' over the time $T$ just as we do with the analogous Zernike modes in a wavefront over an aperture of diameter $D$ \citep{1976JOSA...66..207N}.
+
+The first orthonormal Legendre polynomial is $P_0(\xi) = 1/\sqrt{2}$, which means that $a_0/\sqrt{2}$ is the mean value of $I(t)$ over a time-period of length $T$.
 Thus the variance of the mean is given by
@@
 \end{equation}
-Now from Equation (\ref{eqn:process_var}) \jrmadd{the variance of a single measurement is}
+Now from Equation (\ref{eqn:process_var}) the variance of a single measurement is
 \begin{equation}
@@
 where $N = T/\Delta t$ is the number of measurements.
-This is exactly the result we expect for the variance of the mean in a white noise process with variance $\sigma^2_o$, \jrmadd{giving the familiar ``root-N'' reduction in noise.}
+This is exactly the result we expect for the variance of the mean in a white noise process with variance $\sigma^2_o$, giving the familiar ``root-N'' reduction in noise.
 
@@
 
-% The position vector $\vec{q}$ in the pupil plane is defined by the unit vectors $(\hat{\mathpzc{u}},\hat{\mathpzc{v}})$, such that
-% \begin{equation}
-% \textstyle\vec{q} = u \hat{\mathpzc{u}} + v \hat{\mathpzc{v}} \\
-% \label{eqn:q_def}
-% \end{equation}
-% and the vector spatial-frequency is 
-% \begin{equation}
-% \vec{k} = k_u \hat{\mathpzc{u}} + k_v \hat{\mathpzc{v}}. 
-% \end{equation}
-% We discretize the wavefront control problem according to
-% \begin{equation}
-% \Delta k = \frac{1}{D}.
-% \end{equation}
-% where $D$ is the aperture diameter.  This yields
-% \begin{equation}
-% \textstyle\vec{k}_{mn} = \textstyle\frac{m}{D} \hat{\mathpzc{u}} + \textstyle\frac{n}{D} \hat{\mathpzc{v}}
-% \label{eqn:kq_def}
-% \end{equation}
-% where $m$ and $n$ are integer indices.
-
 We use a real-valued Fourier basis
@@
 \end{equation}
-where $h_{mn}^{+}$ and $h_{mn}^{-}$ are the \mpfadd{real} coefficients of the modes in the same units as wavelength, and the phase $\Phi$ has units of radians.
+where $h_{mn}^{+}$ and $h_{mn}^{-}$ are the real-valued coefficients of the modes in the same units as wavelength, and the phase $\Phi$ has units of radians.
 
@@
 \end{equation}
-\jrmadd{and the modified spatial frequencies are given by} 
+and the modified spatial frequencies are given by
 \begin{eqnarray}
@@
 \end{eqnarray}
-\jrmadd{Equation \ref{eqn:cross_terms} is equivalent to a convolution with the PSF, adding contributions from nearby speckles.}  The cross-terms are a series of terms including products of $h_{mn}^+h_{mn}^-$ and $h_{mn}^ph_{m'n'}^{p'}$ ($mn \neq m'n'$) with $\Ji(x)\Ji(x')$.
+Equation \ref{eqn:cross_terms} is equivalent to a convolution with the PSF, adding contributions from nearby speckles.
+The cross-terms are a series of terms including products of $h_{mn}^+h_{mn}^-$ and $h_{mn}^ph_{m'n'}^{p'}$ ($mn \neq m'n'$) with $\Ji(x)\Ji(x')$.
 As long as the aperture truncation is accounted for when the long-exposure statistics of $h$ are calculated, the sum can be dropped and these small cross-terms vanish.
@@
 
-The quantity we seek is the temporal PSD of the intensity time-series described by Equation (\ref{eqn:cross_terms}).
-\jrmadd{The PSD $\mathcal{T}_{I_{mn}}(f)$ is defined as in Equation (\ref{eqn:psd_def})}.
+The quantity we seek is the temporal PSD of the intensity time-series described by Equation (\ref{eqn:cross_terms}).  The PSD $\mathcal{T}_{I_{mn}}(f)$ is defined as in Equation (\ref{eqn:psd_def}).
 %is the ensemble average of the modulus-squared of its %Fourier transform:
@@
 \end{equation}
-\mpfrmv{We thus arrive at the need to determine the PSD of}\mpfadd{We therefore directly relate the temporal PSD of $I_\phi$ to the temporal PSD of} $(h_{mn}^{+})^2 + (h_{mn}^{-})^2$.
-\mpfrmv{Next we review how the PSDs of $h_{mn}^{+}$ and $h_{mn}^{-}$ can be calculated.}  \mpfadd{This can be expressed in terms of the PSDs for $h_{mn}^{+}$ and $h_{mn}^{-}$; we review how these can be calculated next.}
+We therefore directly relate the temporal PSD of $I_\phi$ to the temporal PSD of $(h_{mn}^{+})^2 + (h_{mn}^{-})^2$.
+This can be estimated given the PSDs for $h_{mn}^{+}$ and $h_{mn}^{-}$; we review how these can be calculated next.
 
@@
 
-We consider ETFs and NTFs for control laws based on \jrmadd{general} linear filters of the form
+We consider ETFs and NTFs for control laws based on general linear filters of the form
 \begin{equation}
@@
 
-With $J=1$, $a_1 = 1$, $L=0$, and $b_0 = 1$, Equation (\ref{eqn:linfilt}) represents the \mpfadd{simplest control law, the} simple integrator (SI)\mpfrmv{, the simplest control law}.
+With $J=1$, $a_1 = 1$, $L=0$, and $b_0 = 1$, Equation (\ref{eqn:linfilt}) represents the simple integrator (SI).
 \citet{2018JATIS...4a9001M} showed how adding more coefficients, using the well-known Linear Prediction (LP) technique, results in significant performance improvements in terms of post-coronagraph contrast.
 Most, if not all, linear controllers can be cast in the form of Equation (\ref{eqn:linfilt}) \citep{2007JOSAA..24.2645P, 2021arXiv210307566H}.
-
-% \begin{equation}
-% H_{wfs}(s) = H_{dm}(s) = \frac{1 - e^{-sT}}{sT}
-% \label{eqn:H_wfs}
-% \end{equation}
-% where $T = 1/f_s$, $f_s$ being the loop sampling frequency, and $s = i 2\pi f$
-% 
-% \begin{equation}
-% H_{\tau}(s) = e^{-s\tau}
-% \label{eqn:H_tau}
-% \end{equation}
-% 
-% \begin{equation}
-% H_{con}(z;g) = \frac{g \sum_{l=0}^L b_l z^{-l}}{1+\sum_{j=1}^J a_l z^{-j} }
-% \label{eqn:H_con}
-% \end{equation}
-% and we can map from the $z$-domain to the $s$-domain by the substitution $z \rightarrow e^{sT}$.
-% 
-% \begin{equation}
-% \mbox{ETF}_{ol}(s;g) = H_{wfs}(s) H_{con}(s;g) H_{\tau}(s) H_{dm}(s)
-% \label{eqn:etf_ol}
-% \end{equation}
-% From this it follows that the closed-loop ETF is
-% \begin{equation}
-% \mbox{ETF}_{cl}(s;g) = \left(1 + H_{wfs}(s) H_{con}(s;g) H_{\tau}(s) H_{dm}(s) \right)^{-1}
-% \end{equation}
-% The closed-loop noise transfer function (NTF) is
-% \begin{equation}
-% \mbox{NTF}_{cl}(s;g) = H_{dm}(s) H_{\tau}(s)  H_{con}(s;g)\:\mbox{ETF}_{cl}(s ; g)
-% \end{equation}
 
@@
 \end{equation}
-\jrmadd{This temporal-phase relationship between the two modes at each spatial frequency must be included in the PSD calculation in order to obtain accurate speckle lifetimes}.
+This temporal-phase relationship between the two modes at each spatial frequency must be included in the PSD calculation in order to obtain accurate speckle lifetimes.
 
@@
 Unfortunately, we have found no purely analytic way to calculate $\langle \left| \mathcal{F}_t\left\{ h^2 \right\} \right|^2 \rangle$ given only $\langle \left| \mathcal{F}_t\left\{ h \right\} \right|^2 \rangle$.
-\jrmadd{ This is further complicated by the spatial correlations and temporal relationship between modes.}  We therefore employ a straightforward Monte Carlo procedure:
+This is further complicated by the spatial correlations and temporal relationship between modes.
+We therefore employ a Monte Carlo procedure:
 \begin{enumerate}
 \item Generate an OL time-series for each $h_{mn}^p$ for both parities.
-\jrmadd{Note that $h_{mn}^-$ is needed for the spatial correlation step, even though it is discarded eventually.} \label{step:olh}
+Note that $h_{mn}^-$ is needed for the optional spatial correlation step, even though it is discarded eventually.
+\label{step:olh}
 \begin{enumerate}
@@
 \item Multiply the Fourier transform by the square-root of the OL PSD.  We now have $\mathcal{F}_t\left\{h_{mn}^p\right\}$ \citep[cf.][]{kasdin_falpha}.
-\item Apply the inverse Fourier transform to produce $h_{mn}^p(t)$.  \jrmadd{If the correlation matrix will be applied in Step \ref{step:corr}, then} normalize to unit variance.
+\item Apply the inverse Fourier transform to produce $h_{mn}^p(t)$.  If the correlation matrix will be applied in Step \ref{step:corr}, then normalize to unit variance.
 \item Repeat for each mode.
 \end{enumerate}
-\item \jrmadd{Optionally,} Correlate the modal time-series by multiplying the vector of coefficients at each time-step by the decomposition of the covariance matrix. \label{step:corr}
+\item Optionally, correlate the modal time-series by multiplying the vector of coefficients at each time-step by the decomposition of the covariance matrix. \label{step:corr}
 \item For each spatial frequency $mn$, replace $h_{mn}^-(t)$ with the phase shifted $h_{mn}^+(t)$.
@@
 \item Finally, we calculate $|\mathcal{F}_t\{I_{mn}(t)\}|^2$.  For best results this is done with a window function applied.
-\item Now the entire procedure is repeated N times, averaging the result to form the estimate of $\mathcal{T}_{I_{mn}}(f)$.
+\item Now the entire procedure is repeated many times, averaging the result to form the estimate of $\mathcal{T}_{I_{mn}}(f)$.
 \end{enumerate}
@@
 
-\afterpage{\clearpage}
+%\afterpage{\clearpage}
 
@@
 \label{sec:simulations}
-We conducted a series of end-to-end numerical simulations of an idealized MagAO-X-like instrument to validate our calculations.
+We conducted a series of end-to-end numerical simulations of an idealized instrument to validate our calculations, using MagAO-X \citep{2020SPIE11448E..4LM} as an example.
 Our goal is not to develop a realistic hardware simulation, rather to ensure that the the calculations produce reasonable results.
@@
 An idealized deformable mirror was used, which exactly produced commanded shapes as projection of modes.
-The DM \jrmadd{was capable of making 2,400 Fourier modes}, corresponding \jrmadd{to the Nyquist limit of a} 48$\times$48 actuator illuminated pupil.
+The DM was capable of making 2,400 Fourier modes, corresponding to the Nyquist limit of a 48$\times$48 actuator illuminated pupil.
 The wavefront sensor (WFS) worked by simply projecting modes onto the phase screen.
@@
 The WFS was spatially filtered in the Fourier domain to minimize aliasing.
-The perfect 1st order coronagraph was simulated.
+The perfect coronagraph was simulated \textbf{\citep{2006A&A...447..397C}}.
 The wavefront was spatially filtered before the coronagraph as well, to minimize Gibbs ringing on the dark hole edge due to the digitized pupil.
 
-The simulated system operated at 2,000 Hz with a 1.5 frame delay in addition to the 1 frame total sample and hold \jrmadd{for WFS integration}.
+The simulated system operated at 2,000 Hz with a 1.5 frame delay in addition to the 1 frame total sample and hold for WFS integration.
 To achieve the 1/2 frame, the WFS used the average turbulence between two time-steps.
@@
 
-\jrmadd{Simulations were conducted first for an infinitely bright star, that is without photon noise, to analyze the impact of only the input disturbance dynamics.
+Simulations were conducted first for an infinitely bright star, that is without photon noise, to analyze the impact of only the input disturbance dynamics.
 Next, an 8th magnitude star was simulated to show the effects of photon noise in the WFS.
@@
 Individual simulation runs covered an elapsed time of 65 seconds, and the final 60 seconds were analyzed to be sure that the control system had stabilized.
-Each of these four cases were run 10 times with a new turbulence phase screen, and the results averaged.}
+Each of these four cases were run 10 times with a new turbulence phase screen, and the results averaged.
 
@@
 Figure \ref{fig:contrast_2Dcomp} shows post-coronagraph contrast maps, comparing the results of calculations using \citet{2018JATIS...4a9001M} to the simulation outputs.
-We show results for an \jrmadd{infinitely bright} ($-\infty$ mag) guide star to show the effects of the input turbulence alone, and for an $8$th-mag guide star.
+We show results for an infinitely bright ($-\infty$ mag) guide star to show the effects of the input turbulence alone, and for an $8$th-mag guide star.
 Figure \ref{fig:contrast_radprof} shows the median radial profile for each of the maps in Figure \ref{fig:contrast_2Dcomp}.
 The agreement is very good, though some small differences are visible in the LP (bottom row).
-\jrmadd{These differences, apparent in both the maps and the profile, are due to the LP controller optimization leading to slightly different controller characteristics.
+These differences, apparent in both the maps and the profile, are due to the LP controller optimization leading to slightly different controller characteristics.
 This is due to the difference in the modal basis actually under control in the simulations compared to the model.
-Another possible source of slight differences is spatial-temporal correlation not fully accounted for in the semi-analytic calculations, which may become apparent at the correction levels under predictive control.}
+Another possible source of slight differences is spatial-temporal correlation not fully accounted for in the semi-analytic calculations, which may become apparent at the correction levels under predictive control.
 
@@
 Note that the color bar is adjusted to best show detail.
-\jrmadd{The agreement between the calculations and simulations shows that the model is accurate}.\label{fig:contrast_2Dcomp}}
+The agreement between the calculations and simulations shows that the model is accurate.\label{fig:contrast_2Dcomp}}
 \end{figure}
@@
 Right: with photon noise on an 8th-mag star assuming the sensitivity of an unmodulated pyramid WFS.
-\jrmadd{The agreement between the calculations and simulations shows that the model is accurate}.
+The agreement between the calculations and simulations shows that the model is accurate.
 \label{fig:contrast_radprof}}
@@
 Note that this required re-fitting the Fourier coefficient after the simulations completed, since the control basis was different.
-\jrmadd{For consistency and brevity we use only this mode for illustration of single-mode dynamics.
-It was arbitrarily chosen, and is representative of typical results.}
+For consistency and brevity we use only this mode for illustration of single-mode dynamics.
+It was arbitrarily chosen, and is representative of typical results.
 
@@
 This better low-frequency performance is true on the 8th mag star for both SI and LP.
-\jrmadd{The difference in integrated power at low frequencies is small and does not appreciably impact final contrast, and as we show in the following section leads to good results in intensity.}
+The difference in integrated power at low frequencies is small and does not appreciably impact final contrast, and as we show in the following section leads to good results in intensity.
 
 The disagreements evident in the figure are due to the differences between the idealized continuous-time semi-analytic calculations and the discretized simulations.
-\jrmadd{In particular the empirical optimization of the control laws produces different dynamical behavior in the simulations compared to the model predictions} Additionally, the use of an orthogonalized basis in the simulation produces different noise rejection characteristics on the 8th mag star.
-
-Overall the comparison of outputs presented here give us confidence that the semi-analytic model gives \jrmadd{reasonable} results, and can be used to analyze the closed-loop contrast performance and wavefront dynamics in a coronagraphic AO system.
+In particular the empirical optimization of the control laws produces different dynamical behavior in the simulations compared to the model predictions Additionally, the use of an orthogonalized basis in the simulation produces different noise rejection characteristics on the 8th mag star.
+
+Overall the comparison of outputs presented here give us confidence that the semi-analytic model gives reasonable results, and can be used to analyze the closed-loop contrast performance and wavefront dynamics in a coronagraphic AO system.
 
@@
 In Figure \ref{fig:binvarcomp} we show the evolution of the variance of the mean with time.
-The lines show the result of applying Equation (\ref{eqn:varmean})  \jrmadd{to PSDs of the intensity in the simulations}.
-\jrmadd{To test the validity of the derivation of speckle lifetime,} the points show the direct calculation of the variance in the simulations, as a function of increasing averaging length.
+The lines show the result of applying Equation (\ref{eqn:varmean}) to PSDs of the intensity in the simulations.
+To test the validity of the derivation of speckle lifetime, the points show the direct calculation of the variance in the simulations, as a function of increasing averaging length.
 Importantly, the solid lines are in nearly perfect agreement with the points, showing the validity of Equation (\ref{eqn:varmean}).
-The dot-dashed lines, \jrmadd{which are the output of the semi-analytic calculations}, are slightly offset for the LP, which is due to differences in simulated vs.
+The dot-dashed lines, which are the output of the semi-analytic calculations, are slightly offset for the LP, which is due to differences in simulated vs.
 predicted contrast (equivalent to the overall variance level) as found in the previous section, but the shape of the predicted evolution with time for both SI and LP controllers is in good agreement with the simulations.
@@
 \caption{Comparison of predicted and simulated variance evolution for speckle intensity at position (m=10,n=14).
-\jrmadd{Here ``calculated'' is the output of the method in Section \ref{sec:recipe}.
+Here ``calculated'' is the output of the method in Section \ref{sec:recipe}.
 The ``sim.
@@
 The comparison between the two simulation-based results shows that the analytic expression for speckle lifetime we derived is valid.
-These are compared to the calculated line to show the result of the semi-analytic method for calculating the intensity PSD without an end-to-end simulation.}  \label{fig:binvarcomp}}
+These are compared to the calculated line to show the result of the semi-analytic method for calculating the intensity PSD without an end-to-end simulation.
+\label{fig:binvarcomp}}
 \end{figure}
@@
 For the integrator the agreement is very good.
-For the predictor on the infinitely bright star, the agreement is good overall but the calculations \jrmadd{modestly} over-predict speckle lifetimes \jrmadd{in a wedge associated with (but not parallel to)} the wind vectors.
-\jrmadd{This results demonstrate that AO control does affect speckle lifetimes, and that the lifetime is significantly shorter under predictive control.
-The lifetime also depends on location in the control region.}  \label{fig:lifetime_2Dcomp}}
+For the predictor on the infinitely bright star, the agreement is good overall but the calculations modestly over-predict speckle lifetimes in a wedge associated with (but not parallel to) the wind vectors.
+This results demonstrate that AO control does affect speckle lifetimes, and that the lifetime is significantly shorter under predictive control.
+The lifetime also depends on location in the control region.
+\label{fig:lifetime_2Dcomp}}
 \end{figure}
@@
 Right: with photon noise on an 8th mag star assuming the sensitivity of an unmodulated pyramid WFS.
-The calculations are somewhat conservative at smaller separations in the bright star case.
-\jrmadd{Speckle lifetime is shorter under predictive control compared to the simple integrator.} \label{fig:lifetime_radprof}}
-\end{figure}
-
-\clearpage
+Speckle lifetime is shorter under predictive control compared to the simple integrator.
+\label{fig:lifetime_radprof}}
+\end{figure}
+
+%\clearpage
 
@@
 Here we simply calculate a histogram of the lifetimes over the 48$\times$48 $\lambda/D$ dark hole for the SI and LP controller.
-We also apply the Monte Carlo procedure to the OL case, \jrmadd{by treating the input PSDs as unmodified by control laws.
+We also apply the Monte Carlo procedure to the OL case, by treating the input PSDs as unmodified by control laws.
 We note that this is not strictly valid as the input phase is then outside the approximation regime assumed in the derivation of Equation (\ref{eqn:cross_terms}).
 However it provides a useful comparison of the change in the distribution of power in the PSD caused by closed-loop control action.
-}
 
@@
 
-%Next, we analyze the change in speckle lifetime with telescope diameter, windspeed, and seeing.
-
-%Figure \ref{fig:scalings} shows the behavior of speckle lifetime at a constant absolute spatial frequency.
-This means that for a given spatial frequency, the shape of the input PSD is the same regardless of $D$, but with an integrated variance different by a factor of $1/D^2$.
-The speckles resulting from that spatial frequency occur at different projected separations according to $\lambda/D$.
-
-%On the 6.5 m aperture using the SI controller, windspeed has a strong effect, but on the larger ELT scale apertures wind has no significant effect.
-Given the range of plausible mean windspeeds, \jrmadd{the ELTs appear to be in a different regime}.
-Were a mean windspeed significantly higher than 30 m/s to occur, the speckle lifetimes on the ELT would be shorter.
-
-%For the LP controller, increasing windspeed tends to increase speckle lifetime.
-This is a result of increasing power concentration in the PSD peaks at $\vec{\mathcal{V}} \cdot \vec{k}$ Hz, which causes the LP controller to optimize there at the expense of worse low-frequency rejection, thus causing the lifetime to increase.
-
-%Diameter has a weak effect, with lifetimes dropping as $D$ increases.
-When combined as ``crossing time'' $D/V$ there is some evidence for the $\tau \propto D/\mathcal{V}$ relation predicted by \citet{2005SPIE.5903..170M} on smaller apertures, but this does not hold for larger telescopes and does not apply to predictive control.
-
-%Finally, our calculations predict that improving seeing ($r_0$ growing larger) lowers speckle lifetime.
-This is because increasing $r_0$ lowers the overall power at each spatial frequency, which in turn causes the control system to perform better at lower temporal frequencies.
-This better low-frequency rejection therefore lowers the speckle lifetime.
-
 \begin{figure}
@@
     \includegraphics[width=3.5in]{scalings_rel_8.pdf}
-    \caption{\jrmadd{Comparison of speckle lifetimes over a range of parameters.
+    \caption{Comparison of speckle lifetimes over a range of parameters.
 We compare speckle lifetimes with a 5th mag guide star (left) and 8th mag guide star (right) on telescopes of diameter $D$ up to 39 m, with wind speeds $\mathcal{V}$ up to 35 m/s, and seeing $r_0$ up to 0.25 m.
@@
 These comparisons demonstrate that AO control has a large impact on speckle lifetime, and that there are no simple thumb rules which can be used to predict speckle lifetime.
-} \label{fig:scalings}}
-\end{figure}
-
-%\begin{figure}
-%    \includegraphics[width=5.5in]{scalings.pdf}
-%\caption{Changes in speckle lifetime with turbulence parameters.
-Here absolute spatial frequency is held constant \jrmadd{for all aperture sizes}.
-\jrmadd{For instance}, the $(m,n)=(10,14)$ mode on the 6.5 m telescope is compared to the $(m,n)=(60,84)$ mode on the 39 m aperture.
-This means the input temporal PSD is identical in shape, though the integrated variance changes $\propto D^{-2}$.
-Results shown here for turbulence only, without WFS noise.
-There are few general trends for $D$ and $V$.
-For the integrator controller there is evidence of a $D/V$ scaling as found by \citet{2005SPIE.5903..170M} up to around $D/V$$\sim$1, however this breaks down on larger apertures and higher winds, and is not evident in predictive control.
-Improvement in $r_0$ (larger value, better seeing) shows a general reduction in speckle lifetime in all cases.
-%\label{fig:scalings}
-%}
-%\end{figure}
-
-\jrmadd{Next, we analyze the change in speckle lifetime with guide star brightness, telescope diameter, wind speed, and seeing.
+\label{fig:scalings}}
+\end{figure}
+
+Next, we analyze the change in speckle lifetime with guide star brightness, telescope diameter, wind speed, and seeing.
 Note that power per Fourier mode scales as $D^{-2}$.
@@
 This means that the input temporal PSD has a different shape as well as different power.
-This compares the performance of coronagraphic AO systems given the improvement in spatial resolution with $D$.}
-
-\jrmadd{Figure \ref{fig:scalings} compares speckle lifetimes for a 5th-magnitude guide star to those for an 8th-mag guide star.
+This compares the performance of coronagraphic AO systems given the improvement in spatial resolution with $D$.
+
+Figure \ref{fig:scalings} compares speckle lifetimes for a 5th-magnitude guide star to those for an 8th-mag guide star.
 The lifetime is generally shorter on the 8th-mag star due to the increased influence of photon noise relative to the turbulence PSD, with the overall trends being similar.
-Lifetimes are shorter under LP control compared to SI control for both guide stars for all parameters.}
-
-\jrmadd{The SI controller, in general, shows a decrease in lifetime with wind speed $\mathcal{V}$ and an increase with diameter $D$ for both guide stars.
+Lifetimes are shorter under LP control compared to SI control for both guide stars for all parameters.
+
+The SI controller, in general, shows a decrease in lifetime with wind speed $\mathcal{V}$ and an increase with diameter $D$ for both guide stars.
 When analyzed as the crossing time $D/\mathcal{V}$, the SI controller scales roughly proportional to the crossing-time up to $D/\mathcal{V}$$\sim$$2$.
 For longer crossing times the trend becomes flat.
-This implies that the control system becomes dominated by photon noise rather than servo lag and so the optimization process no longer produces changes in the output dynamics.}
-
-\jrmadd{The LP controller shows much weaker dependence on the same parameters.
+This implies that the control system becomes dominated by photon noise rather than servo lag and so the optimization process no longer produces changes in the output dynamics.
+
+The LP controller shows much weaker dependence on the same parameters.
 Except at the lowest wind speeds, the speckle lifetime is $\sim$$20$ ms on the 5th-magnitude guide star.
-On the 8th-magnitude star photon noise makes predictive control less effective, and there is a proportional scaling up to $D/\mathcal{V}$$\sim$$3$.}
-
-\jrmadd{In most cases, improving seeing (larger $r_0$), reduces speckle lifetime.
+On the 8th-magnitude star photon noise makes predictive control less effective, and there is a proportional scaling up to $D/\mathcal{V}$$\sim$$3$.
+
+In most cases, improving seeing (larger $r_0$), reduces speckle lifetime.
 As the turbulence power is reduced, the control system is better able to reject low temporal frequencies, resulting in lower speckle lifetimes.
@@
 Furthermore, given the highest jet stream driven wind speeds we can expect, the GSMTs will normally operate in the regime of $D/\mathcal{V} \gtrsim 1$.
-Thus the GSMTs are operating in a different regime compared to the 6.5 m, and the LP controller produces different dynamics  when photon noise becomes important on the 8th-mag star.}
-
-%\begin{figure}
-%\includegraphics[width=5.5in]{scalings_rel.pdf}
-%\caption{Changes in speckle lifetime with turbulence parameters.
-Here relative spatial frequency is held constant.
-The $(m,n)=(10,14)$ mode on the 6.5 m telescope is compared to the $(m,n)=(10,14)$ mode on the 39 m aperture.
-This means the input temporal PSD is different in shape.
-Results shown here for turbulence only, without WFS noise.
-In this comparison there is stronger evidence for the integrator controll showing a $D/V$ scaling as found by \citet{2005SPIE.5903..170M} up to around $D/V$$\sim$2.
-Predictive control shows little dependence on $D$ or $V$.
-%\label{fig:scalings_rel}
-%}
-%\end{figure}
-%\clearpage
+Thus the GSMTs are operating in a different regime compared to the 6.5 m, and the LP controller produces different dynamics  when photon noise becomes important on the 8th-mag star.
 
@@
 \label{sec:reconstruction}
-In Section \ref{sec:psd_lifetime} we developed the expansion of the intensity time-series in terms of orthonormal Legendre polynomials $P_n$.
-Suppose we have a technique for measuring the Legendre mode coefficients $a_n$, such as from WFS telemetry reconstruction.
-We do not specify a specific algorithm, but assuming a suitable one exists we use it determine and then subtract the first $N$ modes from the intensity time-series
-\begin{equation}
-I_{mn,pp}(t) = I_{mn}(t) - \sum_{n=0}^{n=N-1} a_n P_n. 
-\end{equation}
-The outcome will be fewer long-term correlations as the low order modes are successively subtracted, which will in turn lower the speckle lifetime.
+\textbf{A closed-loop wavefront control system measures the coefficients of the controlled modes in the residual wavefront at each loop time step.
+In this analysis we have treated the Fourier basis, with modal coefficients $h_{mn}^{p}(t)$.
+The residual phase is reconstructed as in Equation \ref{eqn:phi_expansion}.
+The control system could use any modal basis, and the phase reconstructed in equivalent fashion.
+More realistic coronagraphs can be modeled, and the wavefront propagated through the system model.
+}.
+
+\textbf{Here we continue analyzing an ExAO system controlling in the Fourier basis, which will produce measured coefficients $\Delta h_{mn}^{p}(t)$ which we assume are recorded for later analysis.
+Additionally we assume, for now, that focal plane intensity is recorded with a detector reading out at the same rate as and synchronized to the WFS, such that measurements $I_{mn}(t)$ are recorded at each pixel.
+The measured focal plane intensity may include signal from a planet in addition to scattered starlight.
+Our goal is to use the measured modal coefficients to estimate the contribution from scattered starlight, so that any planet signal can be measured.}
+
+\textbf{To accomplish this goal, at some later time we use the measured residual coefficients to reconstruct an estimate of the focal plane intensity due to scattered starlight alone.
+To begin, we reconstruct the residual phase based on Equation \ref{eqn:phi_expansion} }
+\begin{equation}
+\Delta\Phi(\vec{q},t) = \frac{2\pi}{\lambda} \sum\limits_{mn} \left[ \Delta h_{mn}^{+}(t) M_{mn}^{+}(\vec{q}) + \Delta h_{mn}^{-}(t) M_{mn}^{-}(\vec{q})\right]
+\label{eqn:phi_residual}
+\end{equation}
+\textbf{We then propagate the residual phase to the focal plane}
+\begin{equation}
+\hat{I}(\vec{r},t) = G \left\{\mathcal{A}(\vec{q}) \mbox{e}^{i\Delta\Phi(\vec{q},t)} \right\}
+\end{equation}
+\textbf{where $G\{\cdot\}$ is the system propagation operator, which includes the effect of the coronagraph.
+See \citet{2017MNRAS.467L.105H} for an example of the construction of such a model.
+In the case of the perfect coronagraph used in this work, the propagation operator reduces to  Equation \ref{eqn:cross_terms}.
+The result is an estimated intensity time series at position $\vec{r}_{mn}$ }
+\begin{equation}
+\hat{I}_{mn}(t) = \hat{I}(\vec{r}_{mn},t)
+\end{equation}
+
+In Section \ref{sec:psd_lifetime} we developed the expansion of the intensity time-series in terms of orthonormal Legendre polynomials $P_\eta$.
+\textbf{For a segment of the reconstructed intensity of length T, we measure the coefficients of the Legendre polynomials}
+\begin{equation}
+\hat{a}_\eta = \int_{-1}^1 \hat{I}_{mn}(\frac{T}{2}\xi) P_\eta(\xi) d\xi. 
+\end{equation}
+\textbf{We then subtract the first $N$ Legendre modes of the reconstructed time-series from the measured focal plane intensity time-series, giving the ``post-processed'' intensity} 
+\begin{equation}
+I_{mn,pp}(t) = I_{mn}(t) - \sum_{\eta=0}^{\eta=N-1} \hat{a}_\eta P_\eta. 
+\end{equation}
+\textbf{The outcome of this post-processing} will be fewer long-term correlations as the low order modes are successively subtracted, which will in turn lower the speckle lifetime.
 We can describe this outcome statistically as a high-pass filter (HPF).
@@
 \begin{equation}
-\mathcal{T}_n(f) = \mathcal{T}(f) \left|Q_n(f)\right|^2
-\end{equation}
-where ${Q}_n$ is the Fourier transform of $P_n$, which we derive in Appendix \ref{app:tau_psd}, giving
-\begin{equation}
- \left|Q_n(f)\right|^2 = \left(\frac{2n+1}{2}\right)\frac{J_{n+\frac{1}{2}}^2(\pi T f)}{T f}.
+\mathcal{T}_\eta(f) = \mathcal{T}(f) \left|Q_\eta(f)\right|^2
+\end{equation}
+where ${Q}_\eta$ is the Fourier transform of $P_\eta$, which we derive in Appendix \ref{app:tau_psd}, giving
+\begin{equation}
+ \left|Q_\eta(f)\right|^2 = \left(\frac{2\eta+1}{2}\right)\frac{J_{\eta+\frac{1}{2}}^2(\pi T f)}{T f}.
 \end{equation}
@@
 \begin{equation}
-\left| H_\mathrm{HPF}(f;N) \right| ^2 = 1 - \frac{1}{2}\sum_{n=0}^{n=N-1} (2n+1) \frac{J_{n+\frac{1}{2}}^2(\pi T f)}{T f}.
+\left| H_\mathrm{HPF}(f;N) \right| ^2 = 1 - \frac{1}{2}\sum_{\eta=0}^{\eta=N-1} (2\eta+1) \frac{J_{\eta+\frac{1}{2}}^2(\pi T f)}{T f}.
 \end{equation}
@@
 
-% Now the measurement of the $a_n$ will have uncertainty due to measurement noise in the WFS.
-In time $T$ there will be $Tf_s$ WFS measurements.
-It then follows from Equations (\ref{eqn:expansecoeff}) and (\ref{eqn:photon_noise}) that the variance in the measurements of $a_n$ will be
-% \begin{equation}
-% \sigma_{a_n}^2 = \frac{4}{Tf_s} (\sigma_{ph}^2)^2.
-% \end{equation}
-% The reconstruction process may be noisy in excess of pure photon noise, say due to a variable calibration.
-We model this by multiplicative factor $\gamma_{r} \ge 1$.
-The PSD due to this measurement noise will be
-% \begin{equation}
-% \mathcal{T}_{a_n}(f;N,\gamma_r) = \gamma_r^2 \frac{2}{Tf_s} (\sigma_{ph}^2)^2 \sum_{n=0}^{n=N-1}\left(2n+1\right)\frac{J_{n+\frac{1}{2}}^2(\pi T f)}{T f}.
-% \end{equation}
-
 Now due to measurement noise the fit to the data will have a fractional uncertainty $\gamma$, so the resulting speckle intensity PSD will be
 \begin{equation}
-\mathcal{T}_{I_{mn},pp}(f;N,\gamma) = \left( 1 -  \frac{(1-\gamma)^2}{2}\sum_{n=0}^{n=N-1}  (2n+1) \frac{J_{n+\frac{1}{2}}^2(\pi T f)}{T f}\right) \mathcal{T}_{I_{mn}}(f).
+\mathcal{T}_{I_{mn},pp}(f;N,\gamma) = \left( 1 -  \frac{(1-\gamma)^2}{2}\sum_{\eta=0}^{\eta=N-1}  (2\eta+1) \frac{J_{\eta+\frac{1}{2}}^2(\pi T f)}{T f}\right) \mathcal{T}_{I_{mn}}(f).
 \end{equation}
@@
 \includegraphics[width=4in]{relsens.pdf}
-\caption{The relative sensitivity, in terms of exposure time, after subtracting the first $N$ Legendre polynomials with uncertainty in the fit $\gamma$ for each $a_n$.
+\caption{The relative sensitivity, in terms of exposure time, after subtracting the first $N$ Legendre polynomials with uncertainty in the fit $\gamma$ for each $a_\eta$.
 The horizontal line at 1 corresponds to the photon-noise limited exposure time on an 8th mag star.
@@
 If the the first $\sim$5 low-order Legendre polynomials can be determined from, say, WFS-based reconstruction, the reduction in speckle lifetime from removing the low-order temporal modes dramatically improves sensitivity by orders of magnitude.
-If the modes are reconstructed to better than $\sim$1\% accuracy in the $a_n$, the photon-noise limit \jrmadd{in the focal plane} can be reached.
+If the modes are reconstructed to better than $\sim$1\% accuracy in the $a_\eta$, the photon-noise limit in the focal plane can be reached.
 
@@
 \subsection{Caveats}
-\jrmadd{The results shown here are subject to several caveats.
-Our analysis is monochromatic.
+
+The results shown here are subject to several caveats.
+All of the assumptions used in the development of \citet{2018JATIS...4a9001M} apply here.
+In particular, our analysis is monochromatic.
 While we assumed that the science and sensing wavelength were the same, neither component of an AO system can be truly monochromatic.
@@
 These speckles can, however, be analyzed with the same tools we developed here.
-We save this for future work.}
-
-\jrmadd{A key assumption underlying the mathematics of the PSD analysis is that the speckle intensity is wide-sense stationary (WSS), which means that the mean is constant and that the autocorrelation does not depend on time.
+We save this for future work.
+
+A key assumption underlying the mathematics of the PSD analysis is that the speckle intensity is wide-sense stationary (WSS), which means that the mean is constant and that the autocorrelation does not depend on time.
 In this context, WSS essentially implies that the atmosphere parameters, namely $r_0$ and the layer $\vec{\mathcal{V}}$, are constant.
@@
 We also expect AO control to damp changes, keeping the output speckle formation process within a more or less well-defined range.
-What this ultimately means is that we must consider the speckle lifetime as itself a time-variable parameter, but expect there to be a representative value over an observation that characterizes long exposure sensitivity.}
+What this ultimately means is that we must consider the speckle lifetime as itself a time-variable parameter, but expect there to be a representative value over an observation that characterizes long exposure sensitivity.
 
@@
 
-\jrmadd{We have developed a semi-analytic model of a coronagraphic AO system that allows us to estimate long-exposure contrast, the intensity dynamics, and statistical lifetimes.
+We have developed a semi-analytic model of a coronagraphic AO system that allows us to estimate long-exposure contrast, the intensity dynamics, and statistical lifetimes.
 Compared to an idealized end-to-end simulation, the results of the model compare favorably.
@@
 For instance, the parameter study in Section \ref{sec:trends} would have been prohibitively expensive if conducted with simulations.
-As design efforts continue for the coming GSMTs, such models will be useful for similar parameter studies to optimize high contrast imaging instruments.}
+As design efforts continue for the coming GSMTs, such models will be useful for similar parameter studies to optimize high contrast imaging instruments.
 
@@
 
-\jrmadd{We analyzed various parameters with potential impacts on speckle lifetime.
+We analyzed various parameters with potential impacts on speckle lifetime.
 Overall, the simplest summary of this analysis is that there is no single thumb-rule-type relationship that applies to all parameters.
@@
 As a result, neither the control action nor the overall power (seeing) should impact speckle lifetime.
-Only the wind would change speckle lifetime for a given \mpfadd{pupil} diameter.}
-
-\jrmadd{Here we have added the time dimension, fully considering the action of AO control as a temporal filter.
+Only the wind would change speckle lifetime for a given pupil diameter.
+
+Here we have added the time dimension, fully considering the action of AO control as a temporal filter.
 Furthermore, we have modeled the impact of high-order predictive control.
@@
 That is, the control reduces the correlation length.
-This is consistent with the findings of \citet{2006OExpr..14.7499P}, which predicted that control laws could be optimized based on speckle lifetime to maximize sensitivity.}
-
-\jrmadd{Predictive control, already expected to improve contrast \citep{2007JOSAA..24.2645P,2017JOSAA..34.1877C,2018JATIS...4a9001M, 2021arXiv210307566H}, will further improve sensitivity by decreasing the speckle averaging lifetime.
-On bright stars, predictive control nearly eliminates dependence on the crossing-time as it acts to reduce servo-lag errors.}
-
-\jrmadd{As ExAO coronagraphs approach the residual atmosphere limit, it will be increasingly important to understand the impact of correlated noise from atmospheric speckles on sensitivity.
-Our results demonstrate that this is a complicated task, that must consider all relevant parameters of the system under study.}
+This is consistent with the findings of \citet{2006OExpr..14.7499P}, which predicted that control laws could be optimized based on speckle lifetime to maximize sensitivity.
+
+Predictive control, already expected to improve contrast \citep{2007JOSAA..24.2645P,2017JOSAA..34.1877C,2018JATIS...4a9001M, 2021arXiv210307566H}, will further improve sensitivity by decreasing the speckle averaging lifetime.
+On bright stars, predictive control nearly eliminates dependence on the crossing-time as it acts to reduce servo-lag errors.
+
+As ExAO coronagraphs approach the residual atmosphere limit, it will be increasingly important to understand the impact of correlated noise from atmospheric speckles on sensitivity.
+Our results demonstrate that this is a complicated task, that must consider all relevant parameters of the system under study.
 
@@
 The temporal decorrelation of speckle noise is one of the main benefits of PSF-subtraction post-processing \citep{2006ApJ...641..556M,2008ApJ...673..647M}, even though algorithms have typically focused on exploiting spatial correlations.
-Recently, studies have begun to consider the time-domain in post-processing explicitly \citep{2021A&A...646A..24S}.
+Recently, studies have begun to consider the time-domain in post-processing explicitly \citep{2021A&A...646A..24S}, and several algorithms show promise in exploiting short exposure focal-plane data \citep{2019PASP..131k4506W, 2021arXiv210306898S} and WFS telemetry (Rodack et al (submitted); Frazin and Rodack (submitted)) to estimate speckle intensity.
 Knowledge of the intensity PSD allows for such algorithms to be analyzed efficiently.
@@
 While we did not present a specific algorithm, we used the temporal intensity PSDs to predict the sensitivity gain possible.
-If the first few low-order temporal modes, here described in terms of the Legendre polynomials, can be determined and subtracted from the intensity in the focal plane, \jrmadd{large} improvements in exposure time are possible.
+If the first few low-order temporal modes, here described in terms of the Legendre polynomials, can be determined and subtracted from the intensity in the focal plane, large improvements in exposure time are possible.
 The fact that only a few low-order temporal modes need to be reconstructed has the important consequence that any such post-processing algorithm does not need to accurately determine the intensity at every time interval, and does not need to be perfect.
@@
 The key development of the post-coronagraph intensity PSD allowed us to analyze post-processing as a time-domain filter and show that it is possible to achieve photon-noise limited sensitivity given a suitable reconstruction algorithm.
-In future work we will use this framework to analyze more realistic turbulence models, extend these techniques to instrumental quasi-static aberrations, \jrmadd{and analyze various post-processing strategies}.
-
-\textbf{Acknowledgements:}
-
-\clearpage
+In future work we will use this framework to analyze more realistic turbulence models, extend these techniques to instrumental quasi-static aberrations, and analyze various post-processing strategies.
+\\
+
+\noindent \textbf{Acknowledgements:} JRM and OG acknowledge support from the NSF (\#1625441, MagAO-X).
+JRM, MPF, and OG acknowledge support from the NSF (\#1710356) and from the Heising Simons Foundation.
+
 \appendix
@@
 
-The Fourier transform of the Legendre polynomials is\footnote{see \url{https://dlmf.nist.gov/18.17\#v} 18.17.19} \jrmadd{[simplified:]}
-\begin{equation}
-\int_{-1}^{1} \mathscr{P}_n(x) e^{i2\pi k x} dx  =  i^n \frac{J_{n+\frac{1}{2}}(2\pi k)}{\sqrt{k}}
-\end{equation}
-with $J_n$ denoting the Bessel functions of the first kind.  We can then write the Fourier transforms of the orthonormal $P_n(x)$ as 
-\begin{equation}
-Q_n(k) = i^n \sqrt{\frac{2n+1}{2}}\, \frac{J_{n+\frac{1}{2}}(2\pi k)}{\sqrt{k}}
-\end{equation}
-which have the property that \jrmadd{[changed limits]}
-\begin{equation}
-\int_{-\infty}^\infty \left| Q_n(k) \right|^2 dk = 1 \mbox{ for all }n.
+The Fourier transform of the Legendre polynomials is\footnote{See \url{https://dlmf.nist.gov/18.17\#v} 18.17.19}
+\begin{equation}
+\int_{-1}^{1} \mathscr{P}_\eta(x) e^{i2\pi k x} dx  =  i^\eta \frac{J_{\eta+\frac{1}{2}}(2\pi k)}{\sqrt{k}}
+\end{equation}
+with $J_\eta$ denoting the Bessel functions of the first kind.  We can then write the Fourier transforms of the orthonormal $P_\eta(x)$ as 
+\begin{equation}
+Q_\eta(k) = i^\eta \sqrt{\frac{2\eta+1}{2}}\, \frac{J_{\eta+\frac{1}{2}}(2\pi k)}{\sqrt{k}}.
+\end{equation}
+For analysis of Legendre mode filtering these have the useful property that\footnote{See \citet{2007tisp.book.....G} 6.574 \#2.}
+\begin{equation}
+\int_{0}^\infty \left| Q_\eta(k) \right|^2 dk = 1 \mbox{ for all }\eta.
 \end{equation}
@@
 
-The expansion of $I(t)$ in terms of the $P_n(\xi)$ is given by Equations (\ref{eqn:basisexpansion}) and (\ref{eqn:expansecoeff}).
+The expansion of $I(t)$ in terms of the $P_\eta(\xi)$ is given by Equations (\ref{eqn:basisexpansion}) and (\ref{eqn:expansecoeff}).
 This series expansion in terms of normalized Legendre polynomials is the 1-D analog of the expansion in Zernike polynomials used by \citet{1976JOSA...66..207N}.
-Following that analysis we write the covariance between any two $P_n(\xi)$ as
-\begin{equation}
-<a_n^* a_{n'}> = \int \int P_n(\xi) C\left(\frac{T}{2}\xi, \frac{T}{2}\xi' \right) P_{n'}(\xi) d\xi d\xi' 
-\end{equation}
-where \jrmadd{the autocorrelation is given by}
+Following that analysis we write the covariance between any two $P_\eta(\xi)$ as
+\begin{equation}
+<a_\eta^* a_{\eta'}> = \int \int P_\eta(\xi) C\left(\frac{T}{2}\xi, \frac{T}{2}\xi' \right) P_{\eta'}(\xi) d\xi d\xi' 
+\end{equation}
+the autocorrelation is given by
 \begin{equation}
@@
 \begin{equation}
-<a_n^* a_{n'}> = \int \int Q_n^*(\kappa) \Phi \left(\frac{2}{T}\kappa, \frac{2}{T}\kappa' \right) Q_{n'}(\kappa) d\kappa d\kappa' 
+<a_\eta^* a_{\eta'}> = \int \int Q_\eta^*(\kappa) \Phi \left(\frac{2}{T}\kappa, \frac{2}{T}\kappa' \right) Q_{\eta'}(\kappa) d\kappa d\kappa' 
 \end{equation}
@@
 \begin{equation}
-\Phi \left(\frac{2}{T}\kappa, \frac{2}{T}\kappa' \right) =  \mathcal{T}\left(\frac{2}{T}\kappa \right) \delta(\kappa-\kappa')
+\Phi \left(\frac{2}{T}\kappa, \frac{2}{T}\kappa' \right) = \frac{2}{T} \mathcal{T}\left(\frac{2}{T}\kappa \right) \delta(\kappa-\kappa')
 \end{equation}
@@
 
-We now have \jrmadd{the covariance of the Legendre coefficients}
-\begin{equation}
-\langle a_n^* a_{n'}\rangle = -\mpfrmv{i^n i^{n'}}\mpfadd{i^{n+n'}} \frac{\sqrt{2n+1}\sqrt{2n'+1}}{T}   \int_{0}^{\infty} \frac{J_{n+\frac{1}{2}}(2\pi \kappa) J_{n'+\frac{1}{2}}(2\pi \kappa)}{\kappa} \mathcal{T}\left( \frac{2}{T} \kappa \right) d\kappa
-\end{equation}
-and for $n = n'$, the variance in a single Legendre mode coefficient over the sample length $T$ is
-\begin{equation}
-\left\langle \left|a_n\right|^2 \right\rangle = \frac{2n+1}{T}  \int_0^{\infty} \frac{ J_{n+\frac{1}{2}}^2(2\pi \kappa)}{\kappa} \mathcal{T}\left( \frac{2}{T} \kappa \right) d\kappa.
+We now have the covariance of the Legendre coefficients
+\begin{equation}
+\langle a_\eta^* a_{\eta'}\rangle = i^{\eta'-\eta} \frac{\sqrt{2\eta+1}\sqrt{2\eta'+1}}{T}   \int_{0}^{\infty} \frac{J_{\eta+\frac{1}{2}}(2\pi \kappa) J_{\eta'+\frac{1}{2}}(2\pi \kappa)}{\kappa} \mathcal{T}\left( \frac{2}{T} \kappa \right) d\kappa
+\end{equation}
+and for $\eta = \eta'$, the variance in a single Legendre mode coefficient over the sample length $T$ is
+\begin{equation}
+\left\langle \left|a_\eta\right|^2 \right\rangle = \frac{2\eta+1}{T}  \int_0^{\infty} \frac{ J_{\eta+\frac{1}{2}}^2(2\pi \kappa)}{\kappa} \mathcal{T}\left( \frac{2}{T} \kappa \right) d\kappa.
 \label{eqn:varcoeff}

### 2021-05-24 response.txt
updated

This paper presents a semi-analytic method for calculating speckle lifetimes in idealized
coronagraphic adaptive optics systems by way of calculating the intensity PSD in the
coronagraphic focal plane. The predictions of this semi-analytic method are favorably compared
with the results of an end-to-end simulation, demonstrating the validity of the less
computationally intensive semi-analytic approach. The semi-analytic approach is used to
analyze the dependence of the speckle lifetime on parameters such as the windspeed, crossing
time, and seeing for telescopes of different aperture sizes, where complex relationships are
identified (counter to previous work). This analysis of speckle lifetimes is motivated by the
argument that the statistical lifetimes of atmospheric speckles limit the ultimate sensitivity of
ground-based high contrast imaging systems.

Overall, this paper presents useful tools that will undoubtedly be used for the planning of GSMT
extreme-AO systems. However, one of the central arguments of the paper -- that reconstruction
post-processing can lead to photon-noise limited sensitivity in the focal plane -- is weakened by
a lack of discussion of how such a reconstruction could be accomplished. While it may be
beyond the scope of this paper to present such a method in detail, the paper would be
strengthened by some discussion of the prospects of making the relevant measurements.

Response: We thank the referee for the careful review and insightful comments.  We have addressed each below.  In particular we have significantly expanded the discussion of post-processing to include an outline of a workable algorithm.

----------------------------------------------
Introduction

"Coronagraphic wavefront control"
Comment: This term should be defined.

Response: We agree that this is not well defined, and have changed to "post-coronagraph".

In paragraph 2 we changed the introduction of FPWFS and LOWFS to read:
".. various post-coronagraph focal-plane (e.g. Galicher et al. 2010) and low-order WFS&C strategies (e.g. Singh et al. 2017) show great promise in significantly reducing the impact of internal speckles."

And in paragraph 3 the last sentence now reads:
" The AO-corrected wavefront, though significantly improved, still causes a halo of speckles which will be
significantly brighter than the instrumental speckles in a high performance coronagraph
system employing the WFS&C strategies discussed above."

----------------------------------------------

Equation 2
Comment: Please give explicit definitions of all terms in the equation.

Response: The terms are now defined after the Equation:

"...where $\tau$ is the noise lifetime, $\sigma_o^2$ is the variance of a single measurement, $\sigma_{mean}^2$ is the variance in the estimate of the mean after time $t$"

----------------------------------------------

"Using AO-corrected images, Fitzgerald & Graham (2006) empirically found significantly longer
values of τ, giving an overall median of 175 ms at Lick observatory. Of note, the results of
Fitzgerald & Graham (2006) hinted that inside the region of the image plane controlled by the
AO system, the speckle lifetime was significantly shorter than outside."

Comment: It would be helpful to give the value of tau both inside and outside the control region
of the AO system here.

Response: We have removed this statement as it is over-interpreting the results of the cited paper.

----------------------------------------------

Section 2
"We could assume that delta_t is sufficiently short that the Is can be treated as a constant"
Comment: Perhaps I_s is a typo here, and should refer to I_*.

Response: This was meant to be a plural s and is confusing.  We have re-worded to:
"We could assume that delta_t is sufficiently short that the each of the I terms can be treated as constant"

and changed "Is" to "I terms" in the following sentence.

----------------------------------------------

Section 3
"This means that the mean is constant and the autocorrelation does not depend on time"
Comment: Given the variation of atmospheric turbulence, it would be helpful to justify this
assumption, and more broadly to justify the wide-sense stationary process.

Response: We have moved text from the Discussion to this point, adding "In this context, WSS essentially implies that the atmosphere parameters, namely $r_0$ and the layer $\vec{\mathcal{V}}$, are constant.  This is of course not strictly true, but can be expected to be approximately true over short periods of time, by which we mean at least long enough to estimate a PSD.  WSS also implies that AO control is stable, and we expect the control system to damp changes, keeping the output speckle formation process within a more or less well-defined range."

The relevant text in the discussion section now reads: "A key assumption underlying the mathematics of the PSD analysis is that the speckle intensity is wide-sense stationary (WSS), which means that the mean is constant and that the autocorrelation does not depend on time.  As discussed above, this assumption implies that $r_0$ and the layer $\vec{\mathcal{V}}$ are constant such that the statistics of the AO control output process are WSS.  In the simulations presented these assumptions are true by construction, and so the comparison to our semi-analytic predictions is valid.  On-sky input turbulence parameters will vary, and so the process will not be WSS over long periods.  However, these variations occur over finite times.  For our results to be useful the time-scales of these variations would need to be long enough to estimate a PSD which meaningfully captures the intensity dynamics.   What this ultimately means is that we must consider the speckle lifetime as itself a time-variable parameter, but we can expect there to be a representative value over an observation that characterizes long exposure sensitivity."

----------------------------------------------

Section 5
"The perfect 1st order coronagraph was simulated."
Comment: Please provide a citation here.

Response: We ave droped "1st order" and added a citation.  Te sentence now reads:
"The perfect coronagraph was simulated (Cavarroc et al. 2006)"

----------------------------------------------
Figure 7 caption "Note that the color bar is adjusted in each panel to best show detail."
Comment: Is the colorbar different for all eight panels, or is it the same for each pair of images
(e.g. is the colorbar for the calculated and simulated simple integrator panels the same)? If all
eight panels have different colorbars, they should be included in the figure.

Response: in both Figures 2 and 7 we have changed the text to read "The same colorbar is used for both calculated and simulated within a combination of controller and guidestar, but is adjusted to best show detail for each combination"
----------------------------------------------

Figure 8
Comment: Why do the calculations under-predict the speckle lifetimes at separations greater
than 10L/D for the infinitely bright, LP case?

Response: To address this, we have added additional text at the end of Section 5.4:
"The calculated and simulated lifetimes agree throughout most of the control region.  The noteworthy exception occurs in the LP roughly along a diagonal.  At wider separations the simulations predict lifetimes $\sim 50\%$ higher in a wedged shape region.  As with the differences in the phase dynamics discussed above, we attribute this to differences between the orthogonalized modal basis used in the simulations and the Fourier modes used in the calculations, and the resultant changes in the output of the controller coefficients and gain optimization."

And to the caption of Figure 8 we added: "The large difference in median lifetime between simulation and calculation for the LP is due to the wedge shaped region where the simulation predicts higher lifetimes (see Figure \ref{fig:lifetime_2Dcomp}.  See text for discussion."

----------------------------------------------
Section 6
"The LP controller shows much weaker dependence on the same parameters"

Comment: I find it particularly striking that the speckle lifetimes with the LP controller have a
much weaker dependence on the telescope diameter than the SI controller. Further discussion
of this point would be useful.

Response: We expanded discussion of this.  The text now reads: 
"The LP controller shows much weaker dependence on the same parameters.  Except at the lowest wind speeds, the speckle lifetime is $\sim$$20$ ms on the 5th-magnitude guide star with little variation.  Here the photon noise is well below the input turbulence PSD, and the LP can effectively whiten the disturbance to the photon noise floor.  The result is that the output PSD has the same shape, giving the same speckle lifetime regardless of the parameters.  On the 8th-magnitude star photon noise makes predictive control less effective, causing some variation with the parameters. There is a proportional scaling up to $D/\mathcal{V}$$\sim$$3$."

----------------------------------------------

"Interestingly, on the 8th-mag guide star and for the largest apertures the LP controller shows
the opposite trend. The power in each Fourier mode is proportional to 1/D2. Projected actuator
spacing was held constant, so the photon noise per mode is constant in these comparisons.
Furthermore, given the highest jet stream driven wind speeds we can expect, the GSMTs will
normally operate in the regime of D/V > 1. Thus the GSMTs are operating in a different regime
compared to the 6.5 m, and the LP controller produces different dynamics when photon noise
becomes important on the 8th-mag star."

Comment: The source of this different behavior with r0 is still unclear to me after reading this
discussion. The authors have mentioned several similarities and differences between the
different aperture cases, but the takeaway point regarding the r0 trends is not entirely clear.

Response:  We have added additional explanatory text.  The paragraph introducing the parameter analysis was re-worded to say:

"Note that the input power per Fourier mode scales as $D^{-2}$, that is it will be lower given larger $D$.  This means that the input temporal PSD has a different shape due to relative spatial frequency as well as different power due both to the wind speed and diameter. The input noise PSD will also be lower for each mode, but by $D^{-1}$ since the signal to noise ratio goes as S/N $\propto \sqrt{D}$.  The balance betwen these inputs can be expected to affect the performance of the optimized controllers, resulting in variation in speckle lifetimes."

The paragraph about r_0 now says:

"{For the SI controller, improving seeing (larger $r_0$) reduces speckle lifetime.  As the turbulence power is reduced, the control system is better able to reject low temporal frequencies, resulting in lower speckle lifetimes.  As with $D$ and $\mathcal{V}$, the LP controller shows little variation in lifetimes vs. $r_0$.  Here again, the predictive controller whitens the noise to the same floor set by photon noise, which produces an output PSD of roughly the same shape resulting in similar speckle lifetimes.  The lifetime increases somewhat vs $r_0$ on the larger diameters as the turbulence power becomes relatively lower compared to photon noise, resulting in reduced effectiveness of the predictor such that more correlations remain in the output."

----------------------------------------------

Section 7
Comment: This section is based on the supposition that "we have a technique for measuring the
Legendre mode coefficients an." It is difficult to evaluate the usefulness of this discussion
without understanding the prospects for making such measurements. Perhaps the authors could
weigh in on how or whether the mode coefficients could be measured with current or future
instruments.

Response: We have added a brief introduction to AO PSF reconstruction, and provided an outline of the basic algorithm needed to perform reconstruction
for high-contrast imaging.  The new text read:

[TODO: add text once we're all happy with it]

----------------------------------------------
Section 8
"A key assumption underlying the mathematics of the PSD analysis is that the speckle intensity
is WSS"
Comment: I suggest moving this discussion to the earlier section when the WSS argument was
initially made, or summarizing it there, to help the audience understand the validity of the WSS
argument when it is first introduced.

Response: Addressed in comment above -- we added more discussion in section 3, and re-worded this paragraph to read:

"A key assumption underlying the mathematics of the PSD analysis is that the speckle intensity is wide-sense stationary (WSS), which means that the mean is constant and that the autocorrelation does not depend on time.  As discussed above, this assumption implies that $r_0$ and the layer $\vec{\mathcal{V}}$ are constant such that the statistics of the AO control output process are WSS.  In the simulations presented these assumptions are true by construction, and so the comparison to our semi-analytic predictions is valid.  On-sky input turbulence parameters will vary, and so the process will not be WSS over long periods.  However, these variations occur over finite times.  For our results to be useful the time-scales of these variations would need to be long enough to estimate a PSD which meaningfully captures the intensity dynamics.   What this ultimately means is that we must consider the speckle lifetime as itself a time-variable parameter, but we can expect there to be a representative value over an observation that characterizes long exposure sensitivity."

----------------------------------------------

Section 9
"We have shown that the speckle lifetime is an order of magnitude shorter than previous studies
have found"
Comment: Unless I'm mistaken, this point was not explicitly made in the body of the paper.

Response: This point was not made explicity.  We added the following paragraph at the end of Section 5.4:

"For comparison, the ``crossing-time'' speckle lifetime of \citet{2005SPIE.5903..170M} would be $\tau=0.3D/\mathcal{V} = 224$ ms for the turbulence averaged windspeed used in these calculations, with the caveat that the constant of propotionality is expected to vary with coronagraph. We find lifetimes much shorter than this, up to a factor of 10 shorter for the LP controller. Figures \ref{fig:lifetime_2Dcomp} and \ref{fig:lifetime_radprof} show that speckle lifetime has a strong dependence on control law, and varies with position in the image.  We investigate the $D/\mathcal{V}$ scaling next."
----------------------------------------------

### 2021-05-24 specklelives.tex
updated

@@
 Given their origin within the instrument, and comparatively long lifetimes \citep{2012A&A...541A.136M}, such speckles should be suppressible using wavefront sensing and control (WFS\&C) strategies.
-Recent progress in applying ``dark hole'' contrast optimization with focal plane WFS \citep{2020A&A...638A.117P} and various focal-plane \citep[e.g.][]{2010A&A...509A..31G} and low-order coronagraphic WFS\&C strategies \citep[e.g.][]{2017PASP..129i5002S} show great promise in significantly reducing the impact of internal speckles.
+Recent progress in applying ``dark hole'' contrast optimization with focal plane WFS \citep{2020A&A...638A.117P} and various \textbf{post-coronagraph} focal-plane \citep[e.g.][]{2010A&A...509A..31G} and low-order  WFS\&C strategies \citep[e.g.][]{2017PASP..129i5002S} show great promise in significantly reducing the impact of internal speckles.
 Once active quasi-static speckle control becomes routine, the fundamental sensitivity limit in ground-based high-contrast imaging will be set by the Earth's atmosphere.
@@
 Even with so-called ``extreme'' AO systems \citep[ExAO, ][]{2018ARAA..56..315G} this correction is imperfect.
-The AO-corrected wavefront, though significantly improved, still causes a halo of speckles significantly brighter than the instrumental speckles in a modern high performance coronagraph system using coronagraphic wavefront control.
+The AO-corrected wavefront, though significantly improved, still causes a halo of speckles \textbf{which will be significantly brighter than the instrumental speckles in a high performance coronagraph system employing the WFS\&C strategies discussed above}.
 
@@
 \end{equation}
+\textbf{where $\tau$ is the noise lifetime, $\sigma_o^2$ is the variance of a single measurement, $\sigma_{mean}^2$ is the variance in the estimate of the mean after time $t$}.
 This simply states that the variance in a measurement of the mean improves with the number of uncorrelated trials $t/\tau$.
@@
 Using AO-corrected images, \citet{2006ApJ...637..541F} empirically found significantly longer values of $\tau$, giving an overall median of 175 ms at Lick observatory.
-Of note, the results of \citet{2006ApJ...637..541F} hinted that inside the region of the image plane controlled by the AO system, the speckle lifetime was significantly shorter than outside.
+%Of note, the results of \citet{2006ApJ...637..541F} hinted that inside the region of the image plane controlled by the AO system, the speckle lifetime was significantly shorter than outside.
 
@@
 We have likewise dropped the $t$ dependence, but we note that this is somewhat subtle.
-We could assume that $\Delta t$ is sufficiently short that the $I$s can be treated as constant, and then state the long exposure variance as a sum of the variance in these short exposures.
-Alternatively, we can treat the $I$s in Equation (\ref{eqn:variance}) as mean values, and then $\sigma^2_{h}$ is the mean of the variance.
+We could assume that $\Delta t$ is sufficiently short that the \textbf{each of the $\mathcal{I}$ terms} can be treated as constant, and then state the long exposure variance as a sum of the variance in these short exposures.
+Alternatively, we can treat the $\mathcal{I}$ terms in Equation (\ref{eqn:variance}) as mean values, and then $\sigma^2_{h}$ is the mean of the variance.
 For large total exposure time these will converge.
@@
 This means that the mean is constant and the autocorrelation does not depend on time.
+\textbf{In this context, WSS essentially implies that the atmosphere parameters, namely $r_0$ and the layer $\vec{\mathcal{V}}$, are constant.
+This is of course not strictly true, but can be expected to be approximately true over short periods of time, by which we mean at least long enough to estimate a PSD.
+WSS also implies that AO control is stable, and we expect the control system to damp changes, keeping the output speckle formation process within a more or less well-defined range.}
+
 With the WSS assumption, the Wiener-Khinchin theorem states that the PSD is the Fourier transform of the autocorrelation.
@@
 Right: with photon noise on an 8th mag star assuming the sensitivity of an unmodulated pyramid WFS.
-Note that the color bar is adjusted to best show detail.
+\textbf{The same colorbar is used for both calculated and simulated within a combination of controller and guidestar, but is adjusted to best show detail for each combination}.
 The agreement between the calculations and simulations shows that the model is accurate.\label{fig:contrast_2Dcomp}}
@@
 Finally in Figure \ref{fig:lifetime_2Dcomp} we present maps of speckle lifetime, and in Figure \ref{fig:lifetime_radprof} the median radial profile of speckle lifetime.
+\textbf{The calculated and simulated lifetimes agree throughout most of the control region.
+The noteworthy exception occurs in the LP roughly along a diagonal.
+At wider separations the simulations predict lifetimes $\sim 50\%$ higher in a wedged shape region.
+As with the differences in the phase dynamics discussed above, we attribute this to differences between the orthogonalized modal basis used in the simulations and the Fourier modes used in the calcualtions, and the resultant changes in the output of the controller coefficients and gain optimization.}
+
+\textbf{For comparison, the ``crossing-time'' speckle lifetime of \citet{2005SPIE.5903..170M} would be $\tau=0.3D/\mathcal{V} = 224$ ms for the turbulence averaged windspeed used in these calculations, with the caveat that the constant of propotionality is expected to vary with coronagraph.
+We find lifetimes much shorter than this, up to a factor of 10 shorter for the LP controller.
+Figures \ref{fig:lifetime_2Dcomp} and \ref{fig:lifetime_radprof} show that speckle lifetime has a strong dependence on control law, and varies with position in the image.
+We investigate the $D/\mathcal{V}$ scaling below.}
+
 Overall these results show that Equation (\ref{eqn:psd_lifetime}) correctly describes the statistical speckle lifetime for a given intensity PSD, which in turn can be predicted by our semi-analytic model of an ExAO coronagraph.
@@
 Right: with photon noise on an 8th mag star assuming the sensitivity of an unmodulated pyramid WFS.
-Note that the color bar is adjusted in each panel to best show detail.
+\textbf{The same colorbar is used for both calculated and simulated within a combination of controller and guidestar, but is adjusted to best show detail for each combination}.
 For the integrator the agreement is very good.
 For the predictor on the infinitely bright star, the agreement is good overall but the calculations modestly over-predict speckle lifetimes in a wedge associated with (but not parallel to) the wind vectors.
-This results demonstrate that AO control does affect speckle lifetimes, and that the lifetime is significantly shorter under predictive control.
+These results demonstrate that AO control does affect speckle lifetimes, and that the lifetime is significantly shorter under predictive control.
 The lifetime also depends on location in the control region.
@@
 Speckle lifetime is shorter under predictive control compared to the simple integrator.
-\label{fig:lifetime_radprof}}
-\end{figure}
-
-%\clearpage
+\textbf{The large difference in median lifetime between simulation and calculation for the LP is due to the wedge shaped region where the simulation predicts higher lifetimes (see Figure \ref{fig:lifetime_2Dcomp}.
+See text for discussion.}  \label{fig:lifetime_radprof}}
+\end{figure}
+
+\clearpage
 
@@
 Next, we analyze the change in speckle lifetime with guide star brightness, telescope diameter, wind speed, and seeing.
-Note that power per Fourier mode scales as $D^{-2}$.
 We held relative spatial frequency constant with $D$, such that the resultant speckle occurs at the same relative projected separation in $\lambda/D$ units.
-This means that the input temporal PSD has a different shape as well as different power.
 This compares the performance of coronagraphic AO systems given the improvement in spatial resolution with $D$.
+\textbf{Note that the input power per Fourier mode scales as $D^{-2}$, that is it will be lower given larger $D$.
+This means that the input temporal PSD has a different shape due to relative spatial frequency as well as different power due to both the wind speed and diameter.
+The input noise PSD will also be lower for each mode, but by $D^{-1}$ since the signal to noise ratio goes as S/N $\propto \sqrt{D}$.
+The balance betwen these inputs can be expected to affect the performance of the optimized controllers, resulting in variation in speckle lifetimes.}
 
@@
 The SI controller, in general, shows a decrease in lifetime with wind speed $\mathcal{V}$ and an increase with diameter $D$ for both guide stars.
-When analyzed as the crossing time $D/\mathcal{V}$, the SI controller scales roughly proportional to the crossing-time up to $D/\mathcal{V}$$\sim$$2$.
+When analyzed as the crossing time $D/\mathcal{V}$, the SI controller scales roughly proportional to the crossing-time up to $D/\mathcal{V}$ $\sim$ $2$.
 For longer crossing times the trend becomes flat.
@@
 
-The LP controller shows much weaker dependence on the same parameters.
-Except at the lowest wind speeds, the speckle lifetime is $\sim$$20$ ms on the 5th-magnitude guide star.
-On the 8th-magnitude star photon noise makes predictive control less effective, and there is a proportional scaling up to $D/\mathcal{V}$$\sim$$3$.
-
-In most cases, improving seeing (larger $r_0$), reduces speckle lifetime.
+\textbf{The LP controller shows much weaker dependence on the same parameters.
+Except at the lowest wind speeds, the speckle lifetime is $\sim$$20$ ms on the 5th-magnitude guide star with little variation.
+Here the photon noise is well below the input turbulence PSD, and the LP can effectively whiten the disturbance to the photon noise floor.
+The result is that the output PSD has the same shape, giving the same speckle lifetime regardless of the parameters.
+On the 8th-magnitude star photon noise makes predictive control less effective, causing some variation with the parameters.
+There is a proportional scaling up to $D/\mathcal{V}$$\sim$$3$.}
+
+\textbf{For the SI controller, improving seeing (larger $r_0$) reduces speckle lifetime.
 As the turbulence power is reduced, the control system is better able to reject low temporal frequencies, resulting in lower speckle lifetimes.
-Interestingly, on the 8th-mag guide star and for the largest apertures the LP controller shows the opposite trend.
-The power in each Fourier mode is proportional to $1/D^2$.
-Projected actuator spacing was held constant, so the photon noise per mode is constant in these comparisons.
-Furthermore, given the highest jet stream driven wind speeds we can expect, the GSMTs will normally operate in the regime of $D/\mathcal{V} \gtrsim 1$.
-Thus the GSMTs are operating in a different regime compared to the 6.5 m, and the LP controller produces different dynamics  when photon noise becomes important on the 8th-mag star.
-
-\section{The Potential of Reconstruction Post-Processing}
+As with $D$ and $\mathcal{V}$, the LP controller shows little variation in lifetimes vs.
+$r_0$.
+Here again, the predictive controller whitens the noise to the same floor set by photon noise, which produces an output PSD of roughly the same shape resulting in similar speckle lifetimes.
+The lifetime increases somewhat vs $r_0$ on the larger diameters as the turbulence power becomes relatively lower compared to photon noise, resulting in reduced effectiveness of the predictor such that more correlations remain in the output.
+}
+
+\section{The Potential of \textbf{PSF} Reconstruction Post-Processing}
 \label{sec:reconstruction}
+\textbf{The reconstruction of the post-AO PSF using AO system telemetry has been studied for some time, primarily focused on the long-exposure PSF, e.g.
+\citet{1997JOSAA..14.3057V} and see the recent review by \citep{2020SPIE11448E..0AB}.
+PSF reconstruction in the context of high-contrast imaging is mainly concerned with short exposures and typically has been based on image data alone \citep{2006ApJ...641..556M,2007ApJ...660..770L,2012ApJ...755L..28S}, however the potential for telemetry-based PSF estimation in high-contrast imaging has been recognized \citep{2013ApJ...767...21F,2013ApJ...767..100C}.
+Here we use our treatment of the intensity dynamics behind an AO-fed coronagraph to analyze the performance of PSF estimation and subtraction, particularly motivated by the use of WFS telemetry to improve on the sensitivity limit set by atmospheric speckle lifetimes.}
+
 \textbf{A closed-loop wavefront control system measures the coefficients of the controlled modes in the residual wavefront at each loop time step.
-In this analysis we have treated the Fourier basis, with modal coefficients $h_{mn}^{p}(t)$.
-The residual phase is reconstructed as in Equation \ref{eqn:phi_expansion}.
-The control system could use any modal basis, and the phase reconstructed in equivalent fashion.
-More realistic coronagraphs can be modeled, and the wavefront propagated through the system model.
-}.
-
-\textbf{Here we continue analyzing an ExAO system controlling in the Fourier basis, which will produce measured coefficients $\Delta h_{mn}^{p}(t)$ which we assume are recorded for later analysis.
-Additionally we assume, for now, that focal plane intensity is recorded with a detector reading out at the same rate as and synchronized to the WFS, such that measurements $I_{mn}(t)$ are recorded at each pixel.
-The measured focal plane intensity may include signal from a planet in addition to scattered starlight.
-Our goal is to use the measured modal coefficients to estimate the contribution from scattered starlight, so that any planet signal can be measured.}
-
-\textbf{To accomplish this goal, at some later time we use the measured residual coefficients to reconstruct an estimate of the focal plane intensity due to scattered starlight alone.
+The WFS produces measurements of the residual phase with modal coefficients $\Delta h_{mn}^{p}(t)$ which we assume are recorded.
+Additionally we assume that focal plane intensity is recorded with a detector reading out at the same rate as and synchronized to the WFS, such that measurements $I_{mn}(t)$ are recorded at each pixel.
+The measured focal plane intensity may include signal from a planet in addition to scattered starlight.}
+
+\textbf{At some later time we use the measured residual coefficients to reconstruct an estimate of the focal plane intensity due to scattered starlight alone.
 To begin, we reconstruct the residual phase based on Equation \ref{eqn:phi_expansion} }
 \begin{equation}
-\Delta\Phi(\vec{q},t) = \frac{2\pi}{\lambda} \sum\limits_{mn} \left[ \Delta h_{mn}^{+}(t) M_{mn}^{+}(\vec{q}) + \Delta h_{mn}^{-}(t) M_{mn}^{-}(\vec{q})\right]
+\Delta\Phi(\vec{q},t) = \frac{2\pi}{\lambda} \sum\limits_{mn} \left[ \Delta h_{mn}^{+}(t) M_{mn}^{+}(\vec{q}) + \Delta h_{mn}^{-}(t) M_{mn}^{-}(\vec{q})\right].
 \label{eqn:phi_residual}
@@
 \end{equation}
-\textbf{where $G\{\cdot\}$ is the system propagation operator, which includes the effect of the coronagraph.
+\textbf{where $\mathcal{A}(\vec{q})$ defines the pupil and $G\{\cdot\}$ is the system propagation operator, which could include information such as the effect of the coronagraph, non-common path aberrations (not sensed by the WFS which measures the $\Delta h_{mn}^{p}(t)$), and Fresnel propagation.
 See \citet{2017MNRAS.467L.105H} for an example of the construction of such a model.
-In the case of the perfect coronagraph used in this work, the propagation operator reduces to  Equation \ref{eqn:cross_terms}.
-The result is an estimated intensity time series at position $\vec{r}_{mn}$ }
+In the case of the perfect coronagraph and ideal optical system used in this work, the propagation operator reduces to Equation \ref{eqn:cross_terms}.
+The result of applying $G\{\cdot\}$ to the reconstructed wavefront is an estimated intensity time series at position $\vec{r}_{mn}$ }
 \begin{equation}
@@
 \begin{equation}
-I_{mn,pp}(t) = I_{mn}(t) - \sum_{\eta=0}^{\eta=N-1} \hat{a}_\eta P_\eta. 
+I_{mn,pp}(t) = I_{mn}(t) - \sum_{\eta=0}^{\eta=N-1} \hat{a}_\eta P_\eta\left(\frac{2}{T}t\right). 
 \end{equation}
@@
 In the figure, the horizontal line corresponds to the photon-noise limited exposure time on an 8th magnitude star.
-If the the first $\sim$5 low-order Legendre polynomials can be determined from, say, WFS-based reconstruction, the reduction in speckle lifetime from removing the low-order temporal modes dramatically improves sensitivity by orders of magnitude.
+If the first $\sim$5 low-order Legendre polynomials can be determined from, say, WFS-based reconstruction, the reduction in speckle lifetime from removing the low-order temporal modes improves sensitivity \textbf{significantly}.
 If the modes are reconstructed to better than $\sim$1\% accuracy in the $a_\eta$, the photon-noise limit in the focal plane can be reached.
@@
 
-A key assumption underlying the mathematics of the PSD analysis is that the speckle intensity is wide-sense stationary (WSS), which means that the mean is constant and that the autocorrelation does not depend on time.
-In this context, WSS essentially implies that the atmosphere parameters, namely $r_0$ and the layer $\vec{\mathcal{V}}$, are constant.
-This is of course not strictly true, but can be expected to be approximately true over short periods of time --- at least long enough to estimate a PSD.
-We also expect AO control to damp changes, keeping the output speckle formation process within a more or less well-defined range.
-What this ultimately means is that we must consider the speckle lifetime as itself a time-variable parameter, but expect there to be a representative value over an observation that characterizes long exposure sensitivity.
+\textbf{A key assumption underlying the mathematics of the PSD analysis is that the speckle intensity is wide-sense stationary (WSS), which means that the mean is constant and that the autocorrelation does not depend on time.
+As discussed above, this assumption implies that $r_0$ and the layer $\vec{\mathcal{V}}$ are constant such that the statistics of the AO control output process are WSS.
+In the simulations presented these assumptions are true by construction, and so the comparison to our semi-analytic predictions is valid.
+On-sky input turbulence parameters will vary, and so the process will not be WSS over long periods.
+However, these variations occur over finite times.
+For our results to be useful the time-scales of these variations would need to be long enough to estimate a PSD which meaningfully captures the intensity dynamics.
+What this ultimately means is that we must consider the speckle lifetime as itself a time-variable parameter, but we can expect there to be a representative value over an observation that characterizes long exposure sensitivity.}
 
@@
 The temporal decorrelation of speckle noise is one of the main benefits of PSF-subtraction post-processing \citep{2006ApJ...641..556M,2008ApJ...673..647M}, even though algorithms have typically focused on exploiting spatial correlations.
-Recently, studies have begun to consider the time-domain in post-processing explicitly \citep{2021A&A...646A..24S}, and several algorithms show promise in exploiting short exposure focal-plane data \citep{2019PASP..131k4506W, 2021arXiv210306898S} and WFS telemetry (Rodack et al (submitted); Frazin and Rodack (submitted)) to estimate speckle intensity.
+Recently, studies have begun to consider the time-domain in post-processing explicitly \citep{2021A&A...646A..24S}, and several algorithms show promise in exploiting short exposure focal-plane data \citep{2019PASP..131k4506W, 2021arXiv210306898S} and WFS telemetry \textbf{\citep[][submitted]{2021arXiv210506589R}\citep[][submitted]{2021arXiv210506590F}} to estimate speckle intensity.
 Knowledge of the intensity PSD allows for such algorithms to be analyzed efficiently.
 We demonstrate this by considering the prospect of telemetry-based reconstruction of the speckle time-series.
-Nominally such an algorithm would make use of the main WFS images, but other information sources such as accelerometers, low-order WFS telemetry, and environmental sensors could be used.
-While we did not present a specific algorithm, we used the temporal intensity PSDs to predict the sensitivity gain possible.
+\textbf{The algorithm we outlined made} use of the main WFS images, but other information sources such as accelerometers, low-order WFS telemetry, and environmental sensors could be used.
+\textbf{The key challenge is in the construction of the system propagation model.}  We used the temporal intensity PSDs to predict the sensitivity gain possible \textbf{if this can be done}.
 If the first few low-order temporal modes, here described in terms of the Legendre polynomials, can be determined and subtracted from the intensity in the focal plane, large improvements in exposure time are possible.

### 2021-05-24 response.txt
update

@@
 
-And to the caption of Figure 8 we added: "The large difference in median lifetime between simulation and calculation for the LP is due to the wedge shaped region where the simulation predicts higher lifetimes (see Figure \ref{fig:lifetime_2Dcomp}.
-See text for discussion."
+And to the caption of Figure 8 we added: "The large difference in median lifetime between simulation and calculation for the
+LP on the bright star is due to the wedge shaped region where the simulation predicts higher lifetimes (see Figure 7, and text for discussion)"
 
@@
 
-"{For the SI controller, improving seeing (larger $r_0$) reduces speckle lifetime.
+"For the SI controller, improving seeing (larger $r_0$) reduces speckle lifetime.
 As the turbulence power is reduced, the control system is better able to reject low temporal frequencies, resulting in lower speckle lifetimes.

### 2021-05-24 specklelives.tex
update

@@
 Speckle lifetime is shorter under predictive control compared to the simple integrator.
-\textbf{The large difference in median lifetime between simulation and calculation for the LP is due to the wedge shaped region where the simulation predicts higher lifetimes (see Figure \ref{fig:lifetime_2Dcomp}.
-See text for discussion.}  \label{fig:lifetime_radprof}}
+\textbf{The large difference in median lifetime between simulation and calculation for the LP on the bright star is due to the wedge shaped region where the simulation predicts higher lifetimes (see Figure \ref{fig:lifetime_2Dcomp}, and text for discussion).}  \label{fig:lifetime_radprof}}
 \end{figure}

### 2021-05-25 specklelives.tex
updated

@@
 Suppression of speckle noise is the key goal of wavefront control and adaptive optics (AO), coronagraphy, and a host of post-processing techniques.
-The temporal correlation of speckles reduces sensitivity as they average much more slowly than photon noise.
 Speckles average at a rate set by the statistical speckle lifetime, and speckle-limited integration time in long exposures is directly proportional to this lifetime.
-As progress continues in post-AO coronagraphic wavefront control, residual atmospheric speckles will become the limiting noise source in high-contrast imaging, so a complete understanding of their statistical behavior is crucial to optimizing high-contrast imaging instruments.
+As progress continues in \textbf{post-coronagraph} wavefront control, residual atmospheric speckles will become the limiting noise source in high-contrast imaging, so a complete understanding of their statistical behavior is crucial to optimizing high-contrast imaging instruments.
 Here we present a novel power spectral density (PSD) method for calculating the lifetime, and develop a semi-analytic method for predicting intensity PSDs behind a coronagraph.
@@
 \begin{equation}
-\sigma_\mathrm{mean}^2 =  \frac{\sigma_o^2}{T/\tau} \tag{\ref{eqn:varmean_def} restated}
-\end{equation}
-as long as $T \gg \tau$.
+\sigma_\mathrm{mean}^2 =  \frac{\sigma_o^2}{t/\tau} \tag{\ref{eqn:varmean_def_intro} restated}
+\end{equation}
+as long as $t \gg \tau$.
 
@@
 The WFS was spatially filtered in the Fourier domain to minimize aliasing.
-The perfect coronagraph was simulated \textbf{\citep{2006A&A...447..397C}}.
+\textbf{The perfect coronagraph was simulated} \textbf{\citep{2006A&A...447..397C}}.
 The wavefront was spatially filtered before the coronagraph as well, to minimize Gibbs ringing on the dark hole edge due to the digitized pupil.

### 2021-05-25 specklelives.tex
updated

@@
 \citet{1997JOSAA..14.3057V} and see the recent review by \citep{2020SPIE11448E..0AB}.
-PSF reconstruction in the context of high-contrast imaging is mainly concerned with short exposures and typically has been based on image data alone \citep{2006ApJ...641..556M,2007ApJ...660..770L,2012ApJ...755L..28S}, however the potential for telemetry-based PSF estimation in high-contrast imaging has been recognized \citep{2013ApJ...767...21F,2013ApJ...767..100C}.
-Here we use our treatment of the intensity dynamics behind an AO-fed coronagraph to analyze the performance of PSF estimation and subtraction, particularly motivated by the use of WFS telemetry to improve on the sensitivity limit set by atmospheric speckle lifetimes.}
+PSF reconstruction in the context of high-contrast imaging is typically based on image data alone \citep{2006ApJ...641..556M,2007ApJ...660..770L,2012ApJ...755L..28S}, however the potential for telemetry-based PSF estimation in high-contrast imaging has been recognized \citep{2013ApJ...767...21F,2013ApJ...767..100C}.
+Here we use our treatment of the intensity dynamics behind an AO-fed coronagraph to analyze the performance of PSF estimation and subtraction.
+We are particularly motivated by the use of WFS telemetry to improve on the sensitivity limit set by atmospheric speckle lifetimes, however this analysis could be applied to conventional post-processing such as the KLIP algorithm \citep{2012ApJ...755L..28S}.}
 

### 2021-05-26 specklelives.tex
Updates from Overleaf

@@
 \textbf{The reconstruction of the post-AO PSF using AO system telemetry has been studied for some time, primarily focused on the long-exposure PSF, e.g.
-\citet{1997JOSAA..14.3057V} and see the recent review by \citep{2020SPIE11448E..0AB}.
-PSF reconstruction in the context of high-contrast imaging is typically based on image data alone \citep{2006ApJ...641..556M,2007ApJ...660..770L,2012ApJ...755L..28S}, however the potential for telemetry-based PSF estimation in high-contrast imaging has been recognized \citep{2013ApJ...767...21F,2013ApJ...767..100C}.
+\citet{1997JOSAA..14.3057V} and see the recent review by \citet{2020SPIE11448E..0AB}.
+PSF reconstruction in the context of high-contrast imaging is typically based on image data alone \citep{2006ApJ...641..556M,2007ApJ...660..770L}, however the potential for telemetry-based PSF estimation in high-contrast imaging has been recognized \citep{2013ApJ...767...21F,2013ApJ...767..100C}.
 Here we use our treatment of the intensity dynamics behind an AO-fed coronagraph to analyze the performance of PSF estimation and subtraction.
@@
 \begin{equation}
-\hat{a}_\eta = \int_{-1}^1 \hat{I}_{mn}(\frac{T}{2}\xi) P_\eta(\xi) d\xi. 
+\hat{a}_\eta = \int_{-1}^1 \hat{I}_{mn}\left(\frac{T}{2}\xi\right) P_\eta(\xi) d\xi. 
 \end{equation}

### 2021-05-26 response.txt
updated

@@
 In particular we have significantly expanded the discussion of post-processing to include an outline of a workable algorithm.
+All changes are in bold text in the attached pdf.
 
@@
 Response: We have added a brief introduction to AO PSF reconstruction, and provided an outline of the basic algorithm needed to perform reconstruction
-for high-contrast imaging.  The new text read:
+for high-contrast imaging.  The new text reads:
 
-[TODO: add text once we're all happy with it]
+The reconstruction of the post-AO PSF using AO system telemetry has been studied
+for some time, primarily focused on the long-exposure PSF, e.g. Veran et al. (1997)
+and see the recent review by Beltramo-Martin et al. (2020). PSF reconstruction in the
+context of high-contrast imaging is typically based on image data alone (Marois et al.
+2006; Lafrenière et al. 2007), however the potential for telemetry-based PSF estimation
+in high-contrast imaging has been recognized (Frazin 2013; Codona & Kenworthy 2013).
+Here we use our treatment of the intensity dynamics behind an AO-fed coronagraph
+to analyze the performance of PSF estimation and subtraction. We are particularly
+motivated by the use of WFS telemetry to improve on the sensitivity limit set by
+atmospheric speckle lifetimes, however this analysis could be applied to conventional
+post-processing such as the KLIP algorithm (Soummer et al. 2012).
+
+A closed-loop wavefront control system measures the coefficients of the controlled
+modes in the residual wavefront at each loop time step. The WFS produces mea-
+surements of the residual phase with modal coefficients ∆hpmn(t) which we assume are
+recorded. Additionally we assume that focal plane intensity is recorded with a detector
+reading out at the same rate as and synchronized to the WFS, such that measurements
+Imn (t) are recorded at each pixel. The measured focal plane intensity may include signal
+from a planet in addition to scattered starlight.
+
+At some later time we use the measured residual coefficients to reconstruct an estimate
+of the focal plane intensity due to scattered starlight alone. To begin, we reconstruct
+the residual phase based on Equation 28
+
+\Delta\Phi(\vec{q},t) = \frac{2\pi}{\lambda} \sum\limits_{mn} \left[ \Delta h_{mn}^{+}(t) M_{mn}^{+}(\vec{q}) + \Delta h_{mn}^{-}(t) M_{mn}^{-}(\vec{q})\right].
+
+We then propagate the residual phase to the focal plane
+
+\hat{I}(\vec{r},t) = G \left\{\mathcal{A}(\vec{q}) \mbox{e}^{i\Delta\Phi(\vec{q},t)} \right\}
+
+where A(~q) defines the pupil and G{·} is the system propagation operator, which could
+include information such as the effect of the coronagraph, non-common path aberrations
+(not sensed by the WFS which measures the ∆hpmn (t)), and Fresnel propagation. See
+Herscovici-Schiller et al. (2017) for an example of the construction of such a model.
+In the case of the perfect coronagraph and ideal optical system used in this work,
+the propagation operator reduces to Equation 29. The result of applying G{·} to the
+reconstructed wavefront is an estimated intensity time series at position ~rmn
+
+\hat{I}_{mn}(t) = \hat{I}(\vec{r}_{mn},t)
+
+In Section 3 we developed the expansion of the intensity time-series in terms of orthonormal Legen-
+dre polynomials Pη . For a segment of the reconstructed intensity of length T, we measure
+the coefficients of the Legendre polynomials
+
+\hat{a}_\eta = \int_{-1}^1 \hat{I}_{mn}\left(\frac{T}{2}\xi\right) P_\eta(\xi) d\xi.
+
+We then subtract the first $N$ Legendre modes of the reconstructed time-series from the measured focal plane intensity time-series, giving the ``post-processed'' intensity
+
+I_{mn,pp}(t) = I_{mn}(t) - \sum_{\eta=0}^{\eta=N-1} \hat{a}_\eta P_\eta\left(\frac{2}{T}t\right).
 
@@
 ----------------------------------------------
+
+Other changes:
+Abstract: For brevity deleted the sentence "The temporal correlation of speckles reduces sensitivity as they average much more slowly than photon noise."
+
+Notation: changed the Legendre polynomial index from n to \eta to avoid confusion with the spatial frequency index
+
+A few minor typos were corrected.

### 2021-05-26 specklelives.tex
merged
updated

@@
 This means that the input temporal PSD has a different shape due to relative spatial frequency as well as different power due to both the wind speed and diameter.
-The input noise PSD will also be lower for each mode, but by $D^{-1}$ since the signal to noise ratio goes as S/N $\propto \sqrt{D}$.
+The input noise PSD will also be lower for each mode.
 The balance betwen these inputs can be expected to affect the performance of the optimized controllers, resulting in variation in speckle lifetimes.}
@@
 The fact that only a few low-order temporal modes need to be reconstructed has the important consequence that any such post-processing algorithm does not need to accurately determine the intensity at every time interval, and does not need to be perfect.
-Accuracy in the mean, slope, and other low-order terms of a few percent is all that is needed to make these gains.
+Accuracy in the mean, slope, and other low-order terms of a few percent is all that is needed to \textbf{lower the speckle lifetime and therefore increase sensitivity}.
 

### 2021-05-26 response.txt
updated

@@
 
-Response: We ave droped "1st order" and added a citation.  Te sentence now reads:
+Response: We ave dropped "1st order" and added a citation.  Te sentence now reads:
 "The perfect coronagraph was simulated (Cavarroc et al. 2006)"
@@
 This means that the input temporal PSD has a different shape due to relative spatial frequency as well as different power due both to the wind speed and diameter.
-The input noise PSD will also be lower for each mode, but by $D^{-1}$ since the signal to noise ratio goes as S/N $\propto \sqrt{D}$.
+The input noise PSD will also be lower for each mode.
 The balance betwen these inputs can be expected to affect the performance of the optimized controllers, resulting in variation in speckle lifetimes."
@@
 
-Response: This point was not made explicity.  We added the following paragraph at the end of Section 5.4:
+Response: This point was not made explicitly.  We added the following paragraph at the end of Section 5.4:
 
-"For comparison, the ``crossing-time'' speckle lifetime of \citet{2005SPIE.5903..170M} would be $\tau=0.3D/\mathcal{V} = 224$ ms for the turbulence averaged windspeed used in these calculations, with the caveat that the constant of propotionality is expected to vary with coronagraph.
+"For comparison, the ``crossing-time'' speckle lifetime of \citet{2005SPIE.5903..170M} would be $\tau=0.3D/\mathcal{V} = 224$ ms for the turbulence averaged windspeed used in these calculations, with the caveat that the constant of proportionality is expected to vary with coronagraph.
 We find lifetimes much shorter than this, up to a factor of 10 shorter for the LP controller.
